### 2014-09-11 ms.tex
initial commit

%\documentstyle[emulateapj,apjfonts,epsfig]{article}
\documentclass[iop,apjl,letterpaper]{emulateapj}
%\documentclass[12pt,preprint]{aastex}
%\usepackage{emulateapj5,apjfonts}

\newcommand{\EXO}{\mbox{EXO 0748-676}}
\newcommand{\keV}{${\rm keV}$}
\newcommand{\Fe}{${\rm Fe} \ $}
\newcommand{\Mn}{${\rm Mn} \ $}
\newcommand{\Cr}{${\rm Cr} \ $}
\newcommand{\V}{${\rm V} \ $}
\newcommand{\Ti}{${\rm Ti} \ $}
\newcommand{\Sc}{${\rm Sc} \ $}
\newcommand{\Ca}{${\rm Ca} \ $}

\shorttitle{Neutron Star Redshift Measurements} 
\shortauthors{Bildsten, Chang and Paerels}

\begin{document}

\title{Atomic Spectral Features During Thermonuclear Flashes on 
Neutron Stars}

\author{Lars Bildsten\altaffilmark{1}, Philip Chang\altaffilmark{2},
and Frits Paerels\altaffilmark{3}}
\altaffiltext{1}{Kavli Institute for Theoretical Physics,  Kohn Hall, University
of California, Santa Barbara, CA 93106, USA; email:
bildsten@kitp.ucsb.edu}
\altaffiltext{2}
{Department of Physics, Broida Hall, University of California,
Santa Barbara, CA 93106; pchang@physics.ucsb.edu}
\altaffiltext{3}
{Columbia Astrophysics Laboratory and Department of Astronomy,
Columbia University, 538 W. 120th St., New York, NY 10027;
frits@astro.columbia.edu}

                        %%%%%%%%%%%%%%%%%%%%%%%%
                        %       abstract       %
                        %%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}

The gravitational redshift measured by Cottam, Paerels and Mendez for
the neutron star (NS) in the low-mass X-ray binary EXO~0748-676
depends on the identification of an absorption line during a type I
burst as the H$\alpha$ line from hydrogenic Fe. We show that Fe is
present above the photosphere as long as $\dot M>4\times
10^{-13}M_\odot {\rm yr^{-1}}$ during the burst. In this limit, the
total Fe column is $N_{\rm Fe}\approx 3\times 10^{19}{\rm cm^{-2}}$
for incident material of solar abundances and only depends on the
nuclear physics of the proton spallation. The Fe destruction creates
many heavy elements with $Z<26$ which may imprint photo-ionization
edges on the NS spectra during a radius expansion event or in a burst
cooling tail. Detecting these features in concert with those from Fe
would confirm a redshift measurement. We also begin to address the
radiative transfer problem, and find that a concentrated Fe layer with
$kT=1.2-1.4\ {\rm keV}$ and column $N_{\rm Fe}= 7-20 \times 10^{20}\
{\rm cm}^{-2}$ (depending on the line depth) above the hotter
continuum photosphere is required to create the H$\alpha$ line of the
observed strength. This estimate must be refined by considerations of
non-LTE effects as well as resonant line transport. Until these are
carried out, we cannot say whether the Fe column from accretion and
spallation is in conflict with the observations.  We also show that
hydrogenic Fe might remain in the photosphere due to radiative
levitation from the high burst flux.

\end{abstract}

\keywords{diffusion -- nuclear reactions -- 
stars: abundances, surface -- stars: neutron -- X-rays: binaries, bursts}

\section{Introduction}

  The dominant thermal flux during type I bursts and superbursts offer
an ideal opportunity for studying neutron star (NS) properties through
spectral measurements of their photospheres (e.g. London, Taam \&
Howard 1984, 1986; Foster, Ross \& Fabian 1987). Though never
confirmed, the reports by Waki et al. (1984), Turner \& Breedon
(1984), Nakamura, Inoue \& Tanaka (1988) and Magnier et al.  (1989) of
4 keV absorption lines during type I bursts motivated theoretical work
on the formation of the Fe Ly$\alpha$ line and edge (e.g. Foster et
al. 1987; Day, Fabian \& Ross 1992). Cottam, Paerels \& Mendez (2002;
hereafter CPM) observed the accreting NS \EXO\ using the Reflection
Grating Spectrometer (RGS) aboard the XMM-Newton satellite and claimed
detection of hydrogen-like and helium-like Fe absorption features from
the $n=2\rightarrow 3$ H$\alpha$ transitions during type I bursts.
This {\it Letter} is an initial study on the implications of the
hydrogen-like feature and predictions for heavy element abundances
near the photosphere. The helium-like feature will be left for future
work.

XMM/Newton observed \EXO\ for 335 ks during which 28 type I X-ray
bursts were seen. CPM combined the burst spectra and split them into a
high luminosity ($L$) part (from the initial rise and fall) and a low
$L$ part, which covers the fading to the baseline $L$. In the high $L$
part, the detected absorption line was identified as the H$\alpha$
transition of hydrogenic Fe. In the low $L$ spectrum, a line
identified as the He-like H$\alpha$ transition was seen. For these
identifications, CPM measured the NS's gravitational redshift at
$1+z=1.35$, giving $R=4.4GM/c^2$ from the line centroid (neglecting
the rotational corrections of Fujimoto 1985; \"Ozel \& Psaltis 2003).
These spectra have photospheric temperatures $kT_{\rm eff,
\infty}\approx 1.6 \ {\rm keV}$ and $kT_{\rm eff, \infty}< 1.4 \ {\rm
keV}$, using the measured color temperatures ($kT_C=1.8 \ {\rm keV}$
and $kT_C<1.5 \ {\rm keV}$; CPM) and correcting for the redshift and
spectral hardening from Comptonization (London et al. 1986; Madej
1997).

In \S~2, we calculate the Fe abundance when accretion is active,
showing that proton spallation of Fe (Bildsten, Salpeter \& Wasserman
1992; hereafter BSW) naturally yields $N_{\rm Fe} = 3.4 \times
10^{19}\ {\rm cm}^{-2}$ for accretion of solar metallicity
material. The spallation also produces large columns of elements with
$Z<26$, which could be observable via their photoionization edges
during the cooling phase of the burst or a radius expansion burst.  In
\S~3, we show that Stark broadening (Paerels 1997) dominates the
H$\alpha$ line equivalent width and that the measured equivalent width
(properly redshifted to the NS surface) of $W_{\rm H\alpha}=10$ eV
implies an Fe column of $N_{\rm Fe} \approx 7-20 \times 10^{20}\ {\rm
cm}^{-2}$ depending on line depth. However, these abundance
calculations are highly uncertain due to non-LTE (NLTE) effects and
resonant line transport which we believe play an important role in the
line formation.  We summarize in \S 4 and note that the radiative flux
during the burst can levitate hydrogenic Fe.

\section{Heavy Element Abundances from Accretion}

The continuum photosphere for the 1.276 keV photons of the Fe
H$\alpha$ transition is at an electron column density of $N_{\rm
e}\approx \sigma_{\rm Th}^{-1}$, where $\sigma_{\rm Th}$ is the
Thomson cross section. If \Fe is uniformly abundant at the solar value
($f_i \approx 3\times 10^{-5}$ by number; Grevesse \& Sauval 1999)
above the continuum photosphere at 1 keV then $N_{\rm Fe}\approx
f_i/\sigma_{\rm Th}\approx 5\times 10^{19}\ {\rm cm^{-2}}$. Joss
(1977) showed that convection during the burst cannot bring burning
fluid elements to the photosphere as the entropy there exceeds that in
the burning layers by large factors. Hence, we look to Fe in the
accreted material as the likely source of the photospheric
elements. BSW studied the deposition and separation of accreting
elements, and we use their results here.

Imagine first that accretion halts during the burst.  In a pure H
atmosphere, the upward pointing atmospheric electric field is
$m_pg/2$, so that the force on Fe is $F=-43m_pg$. The
counter-balancing drag force on the drifting Fe nucleus is
$F_{dr}=Z^2r_{ep}(m_p/m_e)^{1/2}\rho v/m_p$, where $r_{ep}$ is the
electron-proton coupling from BSW.  The downward drift speed is
$v\approx 1 \ {\rm cm \ s^{-1}}T_7^{3/2}/\rho$ so that it takes
$t_{\rm s}\approx 1 \ {\rm s}(y/1\ {\rm g \ cm^{-2}})/T_7^{3/2}$ for
Fe to fall to a column depth $y$.  Hence, if accretion halts, Fe would
sink beneath the photosphere in much less than the burst duration.

Given the sub-Eddington peak luminosity, accretion likely remains
active during the bursts.  Kuulkers et al. (2002) have shown this to
be the case for GX 17+2.  The persistent flux from \EXO \ when type I
bursts occur implies $\dot M> 2\times 10^{-10} M_\odot {\rm yr^{-1}}$
(Gottwald et al. 1986; CPM).  In order to obtain a lower limit on the
accretion rate per unit area we assume that the accretion covers the
whole NS, so that $\dot m> 10^3 {\rm \ g \ cm^{-2}\ s^{-1}}$.  The
measured redshift implies $R=4.4GM/c^2$ (CPM), so that the NS's
surface is inside the last stable orbit regardless of its rotation
rate (Kluzniak \& Wagoner 1985; Kluzniak, Michelson \& Wagoner
1990). The resulting accretion gap allows for the accreting beam to
strike the atmosphere with a velocity comparable to free-fall
(Kluzniak \& Wilson 1991), realizing the Coulomb stopping scenario of
Zel'dovich and Shakura (1969). The integration of a free trajectory
from the last stable orbit to the surface yields an impact angle
$\sin\theta\approx 0.1$ and kinetic energy of $\approx 200-300$
MeV/nucleon.

Since some angular momentum and energy can be lost in the accretion
gap, we presume that $\sin\theta>0.1$ and fix the incident energy at
$200$ MeV/nucleon, resulting in stopping a particle of charge $Ze$ and
mass $Am_p$ after traversing a vertical column $y_{\rm s}\approx
5.4\sin\theta (m_p/\sigma_{\rm Th})A/Z^2$. BSW showed the
repercussions of the $A/Z^2$ scaling: (i) the incident electrons stop
first, (ii) hydrogen and helium penetrate furthest to $y_{\rm s,p}$,
and (iii) all heavier elements stop prior to the H/He beam. For
example, Fe stops at $y_{\rm s,Fe}=0.08y_{\rm s,p}$ and is spalled by
the energetic protons (of current $J_p\approx \dot m/m_p$) until it
has sunk to safety (i.e.  $y>y_{\rm s,p}$). This is the mechanism
shown by BSW to destroy accreted heavy elements. Grazing incidence
reduces the distance that must be fallen to reach safety and increases
the surviving fraction of the incident CNO elements relative to that
in BSW.

 The infalling Fe could have bound electrons as it spirals into the NS
surface. However, the combination of varying Doppler shifts and the
short flight time compared to the residence time in the atmosphere
(derived below) means that the inflow is an unlikely environment for
narrow line formation. We thus focus our discussion on the thermalized
Fe in the atmosphere.  At 200 MeV per nucleon, $\approx 30$\% of the
incident Fe suffer a charge changing nuclear reaction before
traversing $\approx 1\ {\rm g \ cm^{-2}}$ and thermalizing. Accretion
is thus depositing thermal Fe into the NS atmosphere at $y_{\rm s,
Fe}=1.0\sin\theta\ {\rm g \ cm^{-2}}$ at a rate $\dot N_{\rm
Fe}\approx 2f_i J_p/3$. For the spallation cross section (i.e. that
cross section for $^{56}{\rm Fe}$ to be chipped to lower Z)
$\sigma_D\approx 600 \ {\rm mb}$ (Silberberg et al. 1998) the Fe
lifetime to spallation from the proton beam is $t_{\rm d}=(\sigma_D
J_p)^{-1}\approx 2.8\ {\rm ms}(10^3\ {\rm g \ cm^{-2} \ s^{-1}}/\dot
m) $. The Fe must sink to a depth of $y_{\rm s,p}=14\sin\theta\ {\rm g
\ cm^{-2}}$ for safety, so that when $\dot m> 2\ {\rm g \ cm^{-2} \
s^{-1}}(0.1/\sin\theta)T_7^{3/2}$ (or $\dot M>4\times 10^{-13}M_\odot
\ {\rm yr^{-1}} T_7^{3/2}$), destruction dominates settling
(i.e. $t_{\rm s}>t_{\rm d}$).

 Hence, as long as $\dot M$ during the burst exceeds 1\% of that
occurring prior to the burst, the destruction by the proton beam
completely determines the Fe residence time in the upper atmosphere.
This seems a condition easily satisfied for many bursters and greatly
simplifies the calculation of $N_{\rm Fe}$ as well as the Stark
broadening discussed in \S~3, since, in this limit, the Fe is present
only near the stopping point. Because of the complicated interplay of
the currents and forces in the upper atmosphere (see BSW), the local
abundance near the stopping point ranges from $\approx 1-40$ times the
incident abundance. The proton spallation reduces the Fe abundance as
it sinks, imprinting an abundance profile $\propto \exp(-t_{\rm
s}/t_{\rm d})$, which is an exponential decay with depth. The Fe rich
region above the Fe poor layer puts a small inverted gradient in the
mean molecular weight ($\mu$) that would trigger a rapid overturn in
an isothermal layer if $|d\ln\mu/d\ln P|>2/5$. Gradients this large
will not be reached as long as the accreting material has an abundance
less than ten times solar. Doubly diffusive instabilites might still
occur, but those timescales are likely no shorter than $ t_{\rm s}$.

Hence, for finding $N_{\rm Fe}$, we only need to balance the Fe
destruction rate (neglecting the $\alpha$ beam is less than a 20\%
effect, as those cross sections are rarely larger than twice those
induced by protons; Ferrando et al. 1988) $\dot N_{\rm Fe}=N_{\rm
Fe}\sigma_{D}J_p$, with the Fe deposition rate $\dot N_{\rm
Fe}=2f_iJ_p/3$, giving
\begin{equation}
N_{\rm Fe}={2f_i\over 3\sigma_{D}}\approx 3.4\times 10^{19} \ {\rm
cm^{-2}}\left(f_i\over 3\times 10^{-5}\right)\left(600 \ {\rm mb}\over
\sigma_{D}\right). 
\end{equation}
Due to neutron knockout reactions, about 10\% of those Fe nuclei are
$^{55}{\rm Fe}$ and 5\% are $^{54}{\rm Fe}$. The Fe column is
comparable to the uniform solar abundance calculation because
$\sigma_D\approx \sigma_{\rm Th}$.

Since many spallation reactions occur as the nuclei sink, we also
found the abundances of all lower $Z$ and $A$ nuclei. We calculated
this nuclear cascade by following the evolution of each isotope under
the constraint that production equals destruction. The cross-sections
from Silberberg et al. (1998) were used.  We found that the sum of all
nuclei with $20<Z<26$ outnumber Fe by a factor of 2.5-3.0 (see Table
1). In the absence of the BSW spallation scenario, these nuclei would
not be present.  The single most abundant heavy nucleus is $^{56}{\rm
Fe}$.

 The immediate question that comes to mind is whether these spallation
products can be detected.  They may be detected via their H$\alpha$
lines, but the calculation of their predicted EW is left to future
work.  In Table 1, we find the $kT_{1/2}$ at which $1/2$ of these
nuclei would be in the hydrogenic state (for $y=0.5\ {\rm g\ cm^{-2}}$
in Table 1), allowing for a possible photoionization edge detection
once $kT <kT_{1/2}$ (presuming LTE).  To fit all $Z$, we used the
explicit cross-section from Rybicki \& Lightman (1979) and set the
bound-free Gaunt factor to $g=0.8$, giving $\sigma_{\rm
bf}(E)=6.32\times 10^{-18}\ {\rm cm^2}(E_{\rm e}/E)^3 Z^{-2}$ where
$E_{\rm e}$ is the edge energy (shown in Table 1, value in parentheses
is the redshifted edge energy) from Bethe \& Salpeter (1957; hereafter
BS). This $\sigma_{\rm bf}$ agreed to within (3-5)\% to Verner et
al.'s (1996) analytic fits for Ar, Ca and Fe. For now, we estimate the
impact of such an edge on the spectrum by considering the integral
over the optical depth above the photoionization edge, $\tau_{\rm
e}=N\sigma_{\rm bf}$, assuming an effectively cold atmosphere (i.e. no
reemission above the edge) $EW_{\rm e}=\int (1-\exp(-N\sigma_{\rm
bf}(E)))dE$. For $\tau_{\rm e}\ll 1$, the $EW_{\rm e}$ integral is
simply expanded to obtain $EW_{\rm e}=175\ {\rm eV}(N/10^{19}\ {\rm
cm^{-2}})$ integrated from $E=E_{\rm e}$ to $1.3E_{\rm e}$ which is
independent of $Z$.

Table 1 shows $EW_{\rm e}$ assuming the element is present at the
level implied by the spallation scenario and finding the amount of it
in the Hydrogenic ground state from Saha equilibrium for $kT=1.2 \
{\rm keV}$.  As the burst cools or during a radius expansion burst,
these edges could become prominent, and the entry of $\tau_{\rm e,max}
$ in Table 1 presumes all of the spallation created elements have at
least one bound electron. It is intriguing to note that these edge
energies and $EW_{\rm e}$ are close to those previously reported by
Waki et. al. (1984), Turner \& Breedon (1984), Nakamura et al. (1988),
Magnier et al.  (1989) and Kuulkers et al. (2002).

\section{Estimated Iron Abundances from Observed Equivalent Widths}

We now estimate the Fe column needed to give the H$\alpha$ line
observed by CPM.  Motivated by our accretion scenario, we initially
presume that this line forms in a thin isothermal layer of Fe in LTE
at temperature $T$ above the hotter continuum photosphere.  The
observed redshift implies a surface gravity of $g\approx 3\times
10^{14}\ {\rm cm\ s}^{-2}\,(1.4M_\odot/M)$.  Though rapid rotation
($\approx 100\,{\rm Hz}$) might explain the observed line width
($\approx 40\,{\rm eV}$), the line's equivalent width depends most
strongly on 
the intrinsic broadening mechanisms. The scale for thermal Doppler
broadening is $\Delta E_D/E_0 = \sqrt{kT/Am_pc^2}\approx 1.5 \times
10^{-4} (kT/1\ {\rm keV})^{1/2}$.  The natural line width, $\Gamma
\approx 5\times 10^{13}\ {\rm s}^{-1}$ (BS), is given by the
spontaneous decay rate for the $3 \rightarrow 1,2$ transitions, which in
units of the thermal Doppler shift is $a = \hbar\Gamma/2\Delta E_D
\approx 0.3$.

We now show that Stark broadening is larger by an order of magnitude
than either of these effects for the H$\alpha$ transition.  Stark
broadening results from the shift of line energies due to the electric
field from neighboring ions, $e/r_0^2$, where $r_0 = (4\pi
n_e/3)^{-1/3}$ is the mean ion spacing (Mihalas 1978). The energy
perturbation from the linear Stark effect is $\Delta E_{\rm S} =e^2
\Delta x/r_0^2$, where $\Delta x$ is the size of the orbital.  Using
the linear Stark effect is justified since its energy shift at the
densities and temperatures considered exceeds that of the degeneracy
breaking Lamb shift for the relativistic H-like ion (BS). For the
$n=3$ state of Fe at a Thomson depth unity, we find ${\Delta E_{\rm
S}}/E_0 = 1.5 \times 10^{-3}$, an order of magnitude larger than
thermal Doppler.

Since the majority of the intrinsic line width is attributed to Stark
broadening, $N_{\rm Fe}$  must be found from the measured EW
given by the curve of growth (Mihalas 1978),
\begin{equation}
W = 2 A_0 \Delta E \int_0^{\infty} \frac {\eta(\alpha)} {1 +
\eta(\alpha)} d\alpha,
\end{equation}
where $W$ is the EW in energy units, $A_0$ is the normalized line
depth at line center, $\Delta E$ is the intrinsic energy width (in our
case the Stark shift), $\eta$ is the ratio of the line opacity to the
continuum opacity and $\alpha = E/\Delta E$. We use the electric field
distribution for a uniform ion sea (e.g. the Holtsmark distribution,
$W_H$; Mihalas 1978), giving $\eta(\alpha) = \eta_0 W_H(\alpha)$ where
$\eta_0 = {N_{\rm Fe, n=2} \sigma_{\rm line}}/{N_e \sigma_{\rm Th}}$
is the ratio of line optical depth to continuum optical depth at line
center. Depending on density either thermal Doppler or Stark
broadening dominates, so we convolve the Holtsmark distribution with
the Voigt profile to cover the complete range of densities.  We also
ignore Debye screening effects on the electric field distribution
since the deviation from the Holtsmark distribution is small for the
typical $T$ and $\rho$.

 Because rotational broadening acts to lower the apparent line depth
and is likely to be important, the observed value of $A_0\approx 0.3 $
is a lower limit. Indeed, the true line depth could be significantly
larger than 0.3. However, such a large value of $A_0$ is difficult to
understand for realistic temperature gradients and presuming LTE. For
a gray atmosphere (a reasonable approximation for the temperature
profile of a type I burst atmosphere) $A_0$ attains a maximal value of
0.2 when the Fe resides at very low column depths.  The resolution of
this dilemma may be the role of NLTE effects and resonant line
transport in the H$\alpha$ line. Resonant line transport could
generate such an absorption feature with a large line depth without
invoking large temperature contrasts as in LTE (Jefferies 1968). NLTE
effects clearly play a role in determining abundances from the EW
measurements.  For instance, an immediate measure of the role of NLTE
effects is that the radiative ionization rate for the Fe ground state
exceeds the collisional ionization rate (London et al. 1986),
complicating the ionization balance for Fe.

 To resolve the role of NLTE effects, the radiative transfer equation
must be solved self-consistently for both the resonant transport of
the H$\alpha$ and Ly$\alpha$ lines with detailed balanced of the Fe
ionization and excitation in the strong radiation field of a Type I
burst. For now, we simply present results for $A_0 \approx 0.3$, which
is set by observations and $A_0 = 0.2$, which is characteristic of a
gray atmosphere.

We calculate the $N_{\rm Fe}$ required to reach $W_{\rm H\alpha} = 10\
{\rm eV}$ presuming all Fe resides at the column depth, $y=P/g$.  We
assume LTE and that the dominant states for Fe are either fully
ionized, or in the hydrogenic ground state or helium-like ground state
to approximate the partition function.  For radial infall, with the Fe
layer between $kT = 1.2-1.4\ {\rm keV}$ and $A_0 = 0.3\ (0.2)$,
$N_{\rm Fe} = 7\ (20) \times 10^{20}\ {\rm cm}^{-2}$.  This
temperature range is consistent with the absence of helium-like Fe in
the high $L$ phase of the burst (CPM).  Since the occupation of the
$n=1$ state exceeds that in $n=2$ by $\sim 10^{2-3}$, we also
calculate $W_{\rm Ly\alpha}$ for the implied $N_{\rm Fe}$ values.  For
this line, thermal Doppler and natural line broadening dominate Stark
broadening (Madej 1989).  When $kT = 1.2 - 1.4\ {\rm keV}$ and the Fe
resides at the pressure implied by radial accretion, $W_{\rm Ly\alpha}
< 30\ {\rm eV}$.  Foster et al. (1987) also calculated $W_{\rm
Ly\alpha}$ for a uniformly distributed Fe abundance at the solar value
and found that it would not exceed $30\ {\rm eV}$. It would require
an $A_0\approx 1$ to give an Fe column consistent with that calculated
from spallation.  However, such large temperature contrasts (and hence
such large values of $A_0$) are not possible in these atmospheres and
we need to understand the NLTE line transport problem before we
can address this apparent inconsistency.

 Day, Fabian \& Ross (1992) pointed out that a uniform presence of Fe
at the solar value would create an observable photo-ionization edge.
The Fe column require to produce the observed H$\alpha$ line would
yield enormous edges. However, NLTE effects would weaken the strength
of these photoionization features significantly (London et al. 1986).
The reported broad spectral features in the 6-9 keV range (van
Paradijs et al. 1990, Strohmayer \& Brown 2002) during type I bursts
and superbursts need to be re-analysed as possible Fe photo-ionization
edges.

\section{Summary and Conclusions}

  Active accretion in the context of a beam hitting the atmosphere
naturally gives $N_{\rm Fe}\approx 3\times 10^{19}\ {\rm cm}^{-2}$
which only depends on the physics of proton spallation of Fe. In this
scenario, heavy elements with $Z<26$ are produced in quantities
comparable to Fe and may be observed via their photoionization edges
or atomic lines (see Table 1) during the cooling phase of type I
bursts or superbursts. Our initial LTE calculations have found that an
Fe column of $\approx 7-20 \times 10^{20}\ {\rm cm}^{-2}$ is needed to
produce the measured $W_{\rm H\alpha}$ of CPM. However, until a more
complete calculation of the resonant line transport for the H$\alpha$
and Ly$\alpha$ line has been undertaken in this strong radiation
environment, we cannot say whether this factor of twenty discrepency
is certain.

 Throughout this paper, we have presumed that Fe is present because of
active accretion during the burst. However, the radiative force on
hydrogenic Fe in the ground state is $F_{\rm rad} = \int {\sigma_{E}
F_{E}}/{c}\ dE,$ where $F_{E}$ is the flux per photon energy at the
Ly$\alpha$ transition ($E=6.9\ {\rm keV}$). Since $F_{E}$ is nearly
constant across the line, $F_{\rm rad}= \pi h e^2 f_{1\rightarrow 2}
F_E/m_e c^2$ in the optically thin limit (Michaud 1970), where
$f_{1\rightarrow 2}=0.42$ is the oscillator strength from BS. If $F_E$
is given by a black-body, then at $kT_{\rm eff} = 1\ {\rm keV}$, the
radiative acceleration on Fe is $a_{\rm rad} \approx 10^{15} {\rm cm\
s}^{-2}$, a factor of three larger than $g$. The radiative force will
be even larger for realistic type I burst spectra, which are harder
than a blackbody at the observed $T_{\rm eff}$ (London et al. 1986;
Pavlov, et al. 1991).  The Fe diffusion time is less than the burst
duration, so there is adequate time for the accelerated Fe to reach a
new diffusive equilibrium. The calculation of the resulting Fe column,
which could well saturate the line and lower the radiative force, is
beyond the scope of this {\it Letter}, as it requires a simultaneous
solution of the radiative transfer and the Fe diffusion equations,
like that done for hot white dwarfs (e.g. Chayer, Fontaine \& Wesemael
1995).

\acknowledgments

 We thank A. Piro for calculating the infall energy and incident
angles and N. Barghouty for providing the spallation cross-section
code of Silberberg et al. (1998). We thank the referee, Marten van
Kerkwijk, for critical comments and Ira Wasserman for many insightful
conversation and for pointing out the critical role NLTE effects can
play in line formation. This work was supported by the National
Science Foundation under grants PHY99-07949 and AST02-05956 and NASA
through grant NAG 5-8658.  L. B. is a Cottrell Scholar of the Research
Corporation.

\begin{references}

\noindent
Bethe, H.~A. \& Salpeter, E.~E. 1957 \textit{Quantum Mechanics of One-
  and Two-Electron Atoms} (Berlin: Springer-Verlag) (BS)

\noindent
Bildsten, L., Salpeter, E.~E. \& Wasserman, I. 1992, \apj, 384, 143
(BSW)

\noindent
Chayer, P., Fontaine, G., Wesemael, F. 1995, ApJS, 99, 189

\noindent
Cottam, J., Paerels, F. \& Mendez, M. 2002, \nat, 420, 51 (CPM)

\noindent
Day, C. S. R., Fabian, A. C. \& Ross, R. R. 1992, MNRAS, 257, 471

\noindent 
Ferrando, P., et al. 1988, Phys Rev. C, 37, 1490

\noindent
Foster, A. J., Ross, R. R., \& Fabian, A. C. 1987, MNRAS, 228, 259

\noindent
Fujimoto, M. Y. 1985, ApJ, 293, L19

\noindent
Gottwald, M., Haberl, F., Parmar, A. N., \& White, N. E. 1986, \apj,
308, 213

\noindent
Grevesse, N. \& Sauval, A. J. 1999, AA, 347, 348

\noindent
Jefferies, J. T. 1968 \textit{Spectral Line Formation} (Waltham:
Blaisdell)

\noindent 
Joss, P. C. 1977, Nature, 270, 310

\noindent 
Kluzniak, W. \& Wagoner, R. V. 1985, ApJ, 297, 548

\noindent 
Kluzniak, W., Michelson, P. \& Wagoner, R. V. 1990, ApJ, 358, 538

\noindent 
Kluzniak, W. \& Wilson, J. R. 1991, ApJ, 372, L87

\noindent
Kuulkers, E., Homan, J., van der Klis, M., Lewin, W. H. G., \& Mendez,
M. 2002, A\&A, 382, 947

\noindent
London, R.~A., Taam, R.~E., \& Howard, W. M. 1984, 287, L27

\noindent
London, R.~A., Taam, R.~E., \& Howard, W. M. 1986, 306, 170

\noindent
Madej, J. 1989, A\&A, 209, 226

\noindent 
Madej, J. 1991, \apj, 376, 161

\noindent 
Madej, J. 1997, A\&A, 320, 177

\noindent
Magnier, E., Lewin, W. H. G., van Paradijs, J., Tan, J., Penninx, W. \& Damen, E. 1989, MNRAS, 237, 729

\noindent
Michaud, G. 1970, \apj, 160, 641

\noindent
Mihalas, D. 1978, \textit{Stellar Atmospheres}
(San Francisco: W. H. Freeman)

\noindent
Nakamura, N., Inoue, H. \& Tanaka, Y. 1988, PASJ, 40, 209

\noindent
\"Ozel, F., \& Psaltis, D. 2003, \apj, 582, L31

\noindent
Paerels, F. K. 1997, \apj,  476, 47

\noindent
Pavlov, G.~G., Shibanov, Y.~A., \& Zavlin, V.~E. 1991, \mnras, 253, 193

\noindent
Rybicki, G. B. \& Lightman, A. P. 1979 \textit{Radiative Processes in
Astrophysics} (New York: Wiley)

\noindent 
Silberberg, R., Tsao, C. H. \& Barghouty, A. F. 1998, Ap. J., 501, 911

\noindent
Strohmayer, T. E. \& Brown, E. F. 2002, \apj, 566, 1045

\noindent
Turner, M. J. L. \& Breedon, L. M. 1984, MNRAS, 208, 29p

\noindent
van Paradijs, J., Dotani, T., Tanaka, Y., \& Tsuru, T. 1990, PASJ, 42,
633

\noindent 
Verner, D. A., Ferland, G. J., Korista, K. T., Yakovlev, D. G. 1996,
ApJ, 465, 487

\noindent
Waki, I. et al. 1984, PASJ, 36, 819

\noindent
Zel'dovich, Ya. \& Shakura, N. 1969, Soviet Astron, 13, 175

\end{references}

\begin{deluxetable}{c c c c c c}
  \tablecolumns{7} \tablewidth{0pt} \tablecaption{Heavy Element Abundances
  and Atomic Features}

  \tablehead{
    \colhead{} & \colhead{$E_{\rm e}$} &
    \colhead{$\tau_{\rm e,max}$} & \colhead{$N_i/N_{\rm Fe}$} &
    \colhead{$EW_{\rm e}$} & \colhead{$kT_{1/2}$}\\  
    \colhead{} & \colhead{[keV(redsh)]} &\colhead{}
    & \colhead{} & \colhead{[eV]} & \colhead{[keV]}}
  \startdata 
  Fe & 9.279(6.87) & 0.32 & 1.00 & 277 & 1.12\\ 
  Mn & 8.573(6.35) & 0.18 & 0.36 & 66 & 1.05\\ 
  Cr & 7.897(5.85) & 0.26 & 0.48 & 56 & 0.99\\ 
  V  & 7.248(5.37) & 0.26 & 0.43 & 32 & 0.93\\ 
  Ti & 6.627(4.91) & 0.31 & 0.48 & 22 & 0.86\\ 
  Sc & 6.035(4.47) & 0.41 & 0.58 & 17 & 0.81\\ 
  Ca & 5.471(4.05) & 0.47 & 0.59 & 11 & 0.75\\ 
  K  & 4.934(3.65) & 0.48 & 0.55 & 7 & 0.69 \\ 
  Ar & 4.427(3.28) & 0.53 & 0.54 & 4 & 0.65 \\ 
  \enddata
\end{deluxetable}
\end{document}

### 2014-09-16 ms.tex
added new figures and change .tex

@@
 
-The gravitational redshift measured by Cottam, Paerels and Mendez for
-the neutron star (NS) in the low-mass X-ray binary EXO~0748-676
-depends on the identification of an absorption line during a type I
-burst as the H$\alpha$ line from hydrogenic Fe. We show that Fe is
-present above the photosphere as long as $\dot M>4\times
-10^{-13}M_\odot {\rm yr^{-1}}$ during the burst. In this limit, the
-total Fe column is $N_{\rm Fe}\approx 3\times 10^{19}{\rm cm^{-2}}$
-for incident material of solar abundances and only depends on the
-nuclear physics of the proton spallation. The Fe destruction creates
-many heavy elements with $Z<26$ which may imprint photo-ionization
-edges on the NS spectra during a radius expansion event or in a burst
-cooling tail. Detecting these features in concert with those from Fe
-would confirm a redshift measurement. We also begin to address the
-radiative transfer problem, and find that a concentrated Fe layer with
-$kT=1.2-1.4\ {\rm keV}$ and column $N_{\rm Fe}= 7-20 \times 10^{20}\
-{\rm cm}^{-2}$ (depending on the line depth) above the hotter
-continuum photosphere is required to create the H$\alpha$ line of the
-observed strength. This estimate must be refined by considerations of
-non-LTE effects as well as resonant line transport. Until these are
-carried out, we cannot say whether the Fe column from accretion and
-spallation is in conflict with the observations.  We also show that
-hydrogenic Fe might remain in the photosphere due to radiative
-levitation from the high burst flux.
+write this
 
@@
 
-  The dominant thermal flux during type I bursts and superbursts offer
-an ideal opportunity for studying neutron star (NS) properties through
-spectral measurements of their photospheres (e.g. London, Taam \&
-Howard 1984, 1986; Foster, Ross \& Fabian 1987). Though never
-confirmed, the reports by Waki et al. (1984), Turner \& Breedon
-(1984), Nakamura, Inoue \& Tanaka (1988) and Magnier et al.  (1989) of
-4 keV absorption lines during type I bursts motivated theoretical work
-on the formation of the Fe Ly$\alpha$ line and edge (e.g. Foster et
-al. 1987; Day, Fabian \& Ross 1992). Cottam, Paerels \& Mendez (2002;
-hereafter CPM) observed the accreting NS \EXO\ using the Reflection
-Grating Spectrometer (RGS) aboard the XMM-Newton satellite and claimed
-detection of hydrogen-like and helium-like Fe absorption features from
-the $n=2\rightarrow 3$ H$\alpha$ transitions during type I bursts.
-This {\it Letter} is an initial study on the implications of the
-hydrogen-like feature and predictions for heavy element abundances
-near the photosphere. The helium-like feature will be left for future
-work.
+\section{Analytic Expectations}
+\section{Comparison with Numerical Results}
 
-XMM/Newton observed \EXO\ for 335 ks during which 28 type I X-ray
-bursts were seen. CPM combined the burst spectra and split them into a
-high luminosity ($L$) part (from the initial rise and fall) and a low
-$L$ part, which covers the fading to the baseline $L$. In the high $L$
-part, the detected absorption line was identified as the H$\alpha$
-transition of hydrogenic Fe. In the low $L$ spectrum, a line
-identified as the He-like H$\alpha$ transition was seen. For these
-identifications, CPM measured the NS's gravitational redshift at
-$1+z=1.35$, giving $R=4.4GM/c^2$ from the line centroid (neglecting
-the rotational corrections of Fujimoto 1985; \"Ozel \& Psaltis 2003).
-These spectra have photospheric temperatures $kT_{\rm eff,
-\infty}\approx 1.6 \ {\rm keV}$ and $kT_{\rm eff, \infty}< 1.4 \ {\rm
-keV}$, using the measured color temperatures ($kT_C=1.8 \ {\rm keV}$
-and $kT_C<1.5 \ {\rm keV}$; CPM) and correcting for the redshift and
-spectral hardening from Comptonization (London et al. 1986; Madej
-1997).
+What methods did we use?
 
-In \S~2, we calculate the Fe abundance when accretion is active,
-showing that proton spallation of Fe (Bildsten, Salpeter \& Wasserman
-1992; hereafter BSW) naturally yields $N_{\rm Fe} = 3.4 \times
-10^{19}\ {\rm cm}^{-2}$ for accretion of solar metallicity
-material. The spallation also produces large columns of elements with
-$Z<26$, which could be observable via their photoionization edges
-during the cooling phase of the burst or a radius expansion burst.  In
-\S~3, we show that Stark broadening (Paerels 1997) dominates the
-H$\alpha$ line equivalent width and that the measured equivalent width
-(properly redshifted to the NS surface) of $W_{\rm H\alpha}=10$ eV
-implies an Fe column of $N_{\rm Fe} \approx 7-20 \times 10^{20}\ {\rm
-cm}^{-2}$ depending on line depth. However, these abundance
-calculations are highly uncertain due to non-LTE (NLTE) effects and
-resonant line transport which we believe play an important role in the
-line formation.  We summarize in \S 4 and note that the radiative flux
-during the burst can levitate hydrogenic Fe.
+What plots should we show?
+-density vs r
+-velocities vs r
+-angular momentum vs r
+-mass vs r
 
-\section{Heavy Element Abundances from Accretion}
+\begin{figure}
+\plotone{mass.png}
+\end{figure}
 
-The continuum photosphere for the 1.276 keV photons of the Fe
-H$\alpha$ transition is at an electron column density of $N_{\rm
-e}\approx \sigma_{\rm Th}^{-1}$, where $\sigma_{\rm Th}$ is the
-Thomson cross section. If \Fe is uniformly abundant at the solar value
-($f_i \approx 3\times 10^{-5}$ by number; Grevesse \& Sauval 1999)
-above the continuum photosphere at 1 keV then $N_{\rm Fe}\approx
-f_i/\sigma_{\rm Th}\approx 5\times 10^{19}\ {\rm cm^{-2}}$. Joss
-(1977) showed that convection during the burst cannot bring burning
-fluid elements to the photosphere as the entropy there exceeds that in
-the burning layers by large factors. Hence, we look to Fe in the
-accreted material as the likely source of the photospheric
-elements. BSW studied the deposition and separation of accreting
-elements, and we use their results here.
+\begin{figure}
+\plotone{Track_0132_000_rho.png}
+\end{figure}
 
-Imagine first that accretion halts during the burst.  In a pure H
-atmosphere, the upward pointing atmospheric electric field is
-$m_pg/2$, so that the force on Fe is $F=-43m_pg$. The
-counter-balancing drag force on the drifting Fe nucleus is
-$F_{dr}=Z^2r_{ep}(m_p/m_e)^{1/2}\rho v/m_p$, where $r_{ep}$ is the
-electron-proton coupling from BSW.  The downward drift speed is
-$v\approx 1 \ {\rm cm \ s^{-1}}T_7^{3/2}/\rho$ so that it takes
-$t_{\rm s}\approx 1 \ {\rm s}(y/1\ {\rm g \ cm^{-2}})/T_7^{3/2}$ for
-Fe to fall to a column depth $y$.  Hence, if accretion halts, Fe would
-sink beneath the photosphere in much less than the burst duration.
-
-Given the sub-Eddington peak luminosity, accretion likely remains
-active during the bursts.  Kuulkers et al. (2002) have shown this to
-be the case for GX 17+2.  The persistent flux from \EXO \ when type I
-bursts occur implies $\dot M> 2\times 10^{-10} M_\odot {\rm yr^{-1}}$
-(Gottwald et al. 1986; CPM).  In order to obtain a lower limit on the
-accretion rate per unit area we assume that the accretion covers the
-whole NS, so that $\dot m> 10^3 {\rm \ g \ cm^{-2}\ s^{-1}}$.  The
-measured redshift implies $R=4.4GM/c^2$ (CPM), so that the NS's
-surface is inside the last stable orbit regardless of its rotation
-rate (Kluzniak \& Wagoner 1985; Kluzniak, Michelson \& Wagoner
-1990). The resulting accretion gap allows for the accreting beam to
-strike the atmosphere with a velocity comparable to free-fall
-(Kluzniak \& Wilson 1991), realizing the Coulomb stopping scenario of
-Zel'dovich and Shakura (1969). The integration of a free trajectory
-from the last stable orbit to the surface yields an impact angle
-$\sin\theta\approx 0.1$ and kinetic energy of $\approx 200-300$
-MeV/nucleon.
-
-Since some angular momentum and energy can be lost in the accretion
-gap, we presume that $\sin\theta>0.1$ and fix the incident energy at
-$200$ MeV/nucleon, resulting in stopping a particle of charge $Ze$ and
-mass $Am_p$ after traversing a vertical column $y_{\rm s}\approx
-5.4\sin\theta (m_p/\sigma_{\rm Th})A/Z^2$. BSW showed the
-repercussions of the $A/Z^2$ scaling: (i) the incident electrons stop
-first, (ii) hydrogen and helium penetrate furthest to $y_{\rm s,p}$,
-and (iii) all heavier elements stop prior to the H/He beam. For
-example, Fe stops at $y_{\rm s,Fe}=0.08y_{\rm s,p}$ and is spalled by
-the energetic protons (of current $J_p\approx \dot m/m_p$) until it
-has sunk to safety (i.e.  $y>y_{\rm s,p}$). This is the mechanism
-shown by BSW to destroy accreted heavy elements. Grazing incidence
-reduces the distance that must be fallen to reach safety and increases
-the surviving fraction of the incident CNO elements relative to that
-in BSW.
-
- The infalling Fe could have bound electrons as it spirals into the NS
-surface. However, the combination of varying Doppler shifts and the
-short flight time compared to the residence time in the atmosphere
-(derived below) means that the inflow is an unlikely environment for
-narrow line formation. We thus focus our discussion on the thermalized
-Fe in the atmosphere.  At 200 MeV per nucleon, $\approx 30$\% of the
-incident Fe suffer a charge changing nuclear reaction before
-traversing $\approx 1\ {\rm g \ cm^{-2}}$ and thermalizing. Accretion
-is thus depositing thermal Fe into the NS atmosphere at $y_{\rm s,
-Fe}=1.0\sin\theta\ {\rm g \ cm^{-2}}$ at a rate $\dot N_{\rm
-Fe}\approx 2f_i J_p/3$. For the spallation cross section (i.e. that
-cross section for $^{56}{\rm Fe}$ to be chipped to lower Z)
-$\sigma_D\approx 600 \ {\rm mb}$ (Silberberg et al. 1998) the Fe
-lifetime to spallation from the proton beam is $t_{\rm d}=(\sigma_D
-J_p)^{-1}\approx 2.8\ {\rm ms}(10^3\ {\rm g \ cm^{-2} \ s^{-1}}/\dot
-m) $. The Fe must sink to a depth of $y_{\rm s,p}=14\sin\theta\ {\rm g
-\ cm^{-2}}$ for safety, so that when $\dot m> 2\ {\rm g \ cm^{-2} \
-s^{-1}}(0.1/\sin\theta)T_7^{3/2}$ (or $\dot M>4\times 10^{-13}M_\odot
-\ {\rm yr^{-1}} T_7^{3/2}$), destruction dominates settling
-(i.e. $t_{\rm s}>t_{\rm d}$).
-
- Hence, as long as $\dot M$ during the burst exceeds 1\% of that
-occurring prior to the burst, the destruction by the proton beam
-completely determines the Fe residence time in the upper atmosphere.
-This seems a condition easily satisfied for many bursters and greatly
-simplifies the calculation of $N_{\rm Fe}$ as well as the Stark
-broadening discussed in \S~3, since, in this limit, the Fe is present
-only near the stopping point. Because of the complicated interplay of
-the currents and forces in the upper atmosphere (see BSW), the local
-abundance near the stopping point ranges from $\approx 1-40$ times the
-incident abundance. The proton spallation reduces the Fe abundance as
-it sinks, imprinting an abundance profile $\propto \exp(-t_{\rm
-s}/t_{\rm d})$, which is an exponential decay with depth. The Fe rich
-region above the Fe poor layer puts a small inverted gradient in the
-mean molecular weight ($\mu$) that would trigger a rapid overturn in
-an isothermal layer if $|d\ln\mu/d\ln P|>2/5$. Gradients this large
-will not be reached as long as the accreting material has an abundance
-less than ten times solar. Doubly diffusive instabilites might still
-occur, but those timescales are likely no shorter than $ t_{\rm s}$.
-
-Hence, for finding $N_{\rm Fe}$, we only need to balance the Fe
-destruction rate (neglecting the $\alpha$ beam is less than a 20\%
-effect, as those cross sections are rarely larger than twice those
-induced by protons; Ferrando et al. 1988) $\dot N_{\rm Fe}=N_{\rm
-Fe}\sigma_{D}J_p$, with the Fe deposition rate $\dot N_{\rm
-Fe}=2f_iJ_p/3$, giving
-\begin{equation}
-N_{\rm Fe}={2f_i\over 3\sigma_{D}}\approx 3.4\times 10^{19} \ {\rm
-cm^{-2}}\left(f_i\over 3\times 10^{-5}\right)\left(600 \ {\rm mb}\over
-\sigma_{D}\right). 
-\end{equation}
-Due to neutron knockout reactions, about 10\% of those Fe nuclei are
-$^{55}{\rm Fe}$ and 5\% are $^{54}{\rm Fe}$. The Fe column is
-comparable to the uniform solar abundance calculation because
-$\sigma_D\approx \sigma_{\rm Th}$.
-
-Since many spallation reactions occur as the nuclei sink, we also
-found the abundances of all lower $Z$ and $A$ nuclei. We calculated
-this nuclear cascade by following the evolution of each isotope under
-the constraint that production equals destruction. The cross-sections
-from Silberberg et al. (1998) were used.  We found that the sum of all
-nuclei with $20<Z<26$ outnumber Fe by a factor of 2.5-3.0 (see Table
-1). In the absence of the BSW spallation scenario, these nuclei would
-not be present.  The single most abundant heavy nucleus is $^{56}{\rm
-Fe}$.
-
- The immediate question that comes to mind is whether these spallation
-products can be detected.  They may be detected via their H$\alpha$
-lines, but the calculation of their predicted EW is left to future
-work.  In Table 1, we find the $kT_{1/2}$ at which $1/2$ of these
-nuclei would be in the hydrogenic state (for $y=0.5\ {\rm g\ cm^{-2}}$
-in Table 1), allowing for a possible photoionization edge detection
-once $kT <kT_{1/2}$ (presuming LTE).  To fit all $Z$, we used the
-explicit cross-section from Rybicki \& Lightman (1979) and set the
-bound-free Gaunt factor to $g=0.8$, giving $\sigma_{\rm
-bf}(E)=6.32\times 10^{-18}\ {\rm cm^2}(E_{\rm e}/E)^3 Z^{-2}$ where
-$E_{\rm e}$ is the edge energy (shown in Table 1, value in parentheses
-is the redshifted edge energy) from Bethe \& Salpeter (1957; hereafter
-BS). This $\sigma_{\rm bf}$ agreed to within (3-5)\% to Verner et
-al.'s (1996) analytic fits for Ar, Ca and Fe. For now, we estimate the
-impact of such an edge on the spectrum by considering the integral
-over the optical depth above the photoionization edge, $\tau_{\rm
-e}=N\sigma_{\rm bf}$, assuming an effectively cold atmosphere (i.e. no
-reemission above the edge) $EW_{\rm e}=\int (1-\exp(-N\sigma_{\rm
-bf}(E)))dE$. For $\tau_{\rm e}\ll 1$, the $EW_{\rm e}$ integral is
-simply expanded to obtain $EW_{\rm e}=175\ {\rm eV}(N/10^{19}\ {\rm
-cm^{-2}})$ integrated from $E=E_{\rm e}$ to $1.3E_{\rm e}$ which is
-independent of $Z$.
-
-Table 1 shows $EW_{\rm e}$ assuming the element is present at the
-level implied by the spallation scenario and finding the amount of it
-in the Hydrogenic ground state from Saha equilibrium for $kT=1.2 \
-{\rm keV}$.  As the burst cools or during a radius expansion burst,
-these edges could become prominent, and the entry of $\tau_{\rm e,max}
-$ in Table 1 presumes all of the spallation created elements have at
-least one bound electron. It is intriguing to note that these edge
-energies and $EW_{\rm e}$ are close to those previously reported by
-Waki et. al. (1984), Turner \& Breedon (1984), Nakamura et al. (1988),
-Magnier et al.  (1989) and Kuulkers et al. (2002).
-
-\section{Estimated Iron Abundances from Observed Equivalent Widths}
-
-We now estimate the Fe column needed to give the H$\alpha$ line
-observed by CPM.  Motivated by our accretion scenario, we initially
-presume that this line forms in a thin isothermal layer of Fe in LTE
-at temperature $T$ above the hotter continuum photosphere.  The
-observed redshift implies a surface gravity of $g\approx 3\times
-10^{14}\ {\rm cm\ s}^{-2}\,(1.4M_\odot/M)$.  Though rapid rotation
-($\approx 100\,{\rm Hz}$) might explain the observed line width
-($\approx 40\,{\rm eV}$), the line's equivalent width depends most
-strongly on 
-the intrinsic broadening mechanisms. The scale for thermal Doppler
-broadening is $\Delta E_D/E_0 = \sqrt{kT/Am_pc^2}\approx 1.5 \times
-10^{-4} (kT/1\ {\rm keV})^{1/2}$.  The natural line width, $\Gamma
-\approx 5\times 10^{13}\ {\rm s}^{-1}$ (BS), is given by the
-spontaneous decay rate for the $3 \rightarrow 1,2$ transitions, which in
-units of the thermal Doppler shift is $a = \hbar\Gamma/2\Delta E_D
-\approx 0.3$.
-
-We now show that Stark broadening is larger by an order of magnitude
-than either of these effects for the H$\alpha$ transition.  Stark
-broadening results from the shift of line energies due to the electric
-field from neighboring ions, $e/r_0^2$, where $r_0 = (4\pi
-n_e/3)^{-1/3}$ is the mean ion spacing (Mihalas 1978). The energy
-perturbation from the linear Stark effect is $\Delta E_{\rm S} =e^2
-\Delta x/r_0^2$, where $\Delta x$ is the size of the orbital.  Using
-the linear Stark effect is justified since its energy shift at the
-densities and temperatures considered exceeds that of the degeneracy
-breaking Lamb shift for the relativistic H-like ion (BS). For the
-$n=3$ state of Fe at a Thomson depth unity, we find ${\Delta E_{\rm
-S}}/E_0 = 1.5 \times 10^{-3}$, an order of magnitude larger than
-thermal Doppler.
-
-Since the majority of the intrinsic line width is attributed to Stark
-broadening, $N_{\rm Fe}$  must be found from the measured EW
-given by the curve of growth (Mihalas 1978),
-\begin{equation}
-W = 2 A_0 \Delta E \int_0^{\infty} \frac {\eta(\alpha)} {1 +
-\eta(\alpha)} d\alpha,
-\end{equation}
-where $W$ is the EW in energy units, $A_0$ is the normalized line
-depth at line center, $\Delta E$ is the intrinsic energy width (in our
-case the Stark shift), $\eta$ is the ratio of the line opacity to the
-continuum opacity and $\alpha = E/\Delta E$. We use the electric field
-distribution for a uniform ion sea (e.g. the Holtsmark distribution,
-$W_H$; Mihalas 1978), giving $\eta(\alpha) = \eta_0 W_H(\alpha)$ where
-$\eta_0 = {N_{\rm Fe, n=2} \sigma_{\rm line}}/{N_e \sigma_{\rm Th}}$
-is the ratio of line optical depth to continuum optical depth at line
-center. Depending on density either thermal Doppler or Stark
-broadening dominates, so we convolve the Holtsmark distribution with
-the Voigt profile to cover the complete range of densities.  We also
-ignore Debye screening effects on the electric field distribution
-since the deviation from the Holtsmark distribution is small for the
-typical $T$ and $\rho$.
-
- Because rotational broadening acts to lower the apparent line depth
-and is likely to be important, the observed value of $A_0\approx 0.3 $
-is a lower limit. Indeed, the true line depth could be significantly
-larger than 0.3. However, such a large value of $A_0$ is difficult to
-understand for realistic temperature gradients and presuming LTE. For
-a gray atmosphere (a reasonable approximation for the temperature
-profile of a type I burst atmosphere) $A_0$ attains a maximal value of
-0.2 when the Fe resides at very low column depths.  The resolution of
-this dilemma may be the role of NLTE effects and resonant line
-transport in the H$\alpha$ line. Resonant line transport could
-generate such an absorption feature with a large line depth without
-invoking large temperature contrasts as in LTE (Jefferies 1968). NLTE
-effects clearly play a role in determining abundances from the EW
-measurements.  For instance, an immediate measure of the role of NLTE
-effects is that the radiative ionization rate for the Fe ground state
-exceeds the collisional ionization rate (London et al. 1986),
-complicating the ionization balance for Fe.
-
- To resolve the role of NLTE effects, the radiative transfer equation
-must be solved self-consistently for both the resonant transport of
-the H$\alpha$ and Ly$\alpha$ lines with detailed balanced of the Fe
-ionization and excitation in the strong radiation field of a Type I
-burst. For now, we simply present results for $A_0 \approx 0.3$, which
-is set by observations and $A_0 = 0.2$, which is characteristic of a
-gray atmosphere.
-
-We calculate the $N_{\rm Fe}$ required to reach $W_{\rm H\alpha} = 10\
-{\rm eV}$ presuming all Fe resides at the column depth, $y=P/g$.  We
-assume LTE and that the dominant states for Fe are either fully
-ionized, or in the hydrogenic ground state or helium-like ground state
-to approximate the partition function.  For radial infall, with the Fe
-layer between $kT = 1.2-1.4\ {\rm keV}$ and $A_0 = 0.3\ (0.2)$,
-$N_{\rm Fe} = 7\ (20) \times 10^{20}\ {\rm cm}^{-2}$.  This
-temperature range is consistent with the absence of helium-like Fe in
-the high $L$ phase of the burst (CPM).  Since the occupation of the
-$n=1$ state exceeds that in $n=2$ by $\sim 10^{2-3}$, we also
-calculate $W_{\rm Ly\alpha}$ for the implied $N_{\rm Fe}$ values.  For
-this line, thermal Doppler and natural line broadening dominate Stark
-broadening (Madej 1989).  When $kT = 1.2 - 1.4\ {\rm keV}$ and the Fe
-resides at the pressure implied by radial accretion, $W_{\rm Ly\alpha}
-< 30\ {\rm eV}$.  Foster et al. (1987) also calculated $W_{\rm
-Ly\alpha}$ for a uniformly distributed Fe abundance at the solar value
-and found that it would not exceed $30\ {\rm eV}$. It would require
-an $A_0\approx 1$ to give an Fe column consistent with that calculated
-from spallation.  However, such large temperature contrasts (and hence
-such large values of $A_0$) are not possible in these atmospheres and
-we need to understand the NLTE line transport problem before we
-can address this apparent inconsistency.
-
- Day, Fabian \& Ross (1992) pointed out that a uniform presence of Fe
-at the solar value would create an observable photo-ionization edge.
-The Fe column require to produce the observed H$\alpha$ line would
-yield enormous edges. However, NLTE effects would weaken the strength
-of these photoionization features significantly (London et al. 1986).
-The reported broad spectral features in the 6-9 keV range (van
-Paradijs et al. 1990, Strohmayer \& Brown 2002) during type I bursts
-and superbursts need to be re-analysed as possible Fe photo-ionization
-edges.
-
-\section{Summary and Conclusions}
-
-  Active accretion in the context of a beam hitting the atmosphere
-naturally gives $N_{\rm Fe}\approx 3\times 10^{19}\ {\rm cm}^{-2}$
-which only depends on the physics of proton spallation of Fe. In this
-scenario, heavy elements with $Z<26$ are produced in quantities
-comparable to Fe and may be observed via their photoionization edges
-or atomic lines (see Table 1) during the cooling phase of type I
-bursts or superbursts. Our initial LTE calculations have found that an
-Fe column of $\approx 7-20 \times 10^{20}\ {\rm cm}^{-2}$ is needed to
-produce the measured $W_{\rm H\alpha}$ of CPM. However, until a more
-complete calculation of the resonant line transport for the H$\alpha$
-and Ly$\alpha$ line has been undertaken in this strong radiation
-environment, we cannot say whether this factor of twenty discrepency
-is certain.
-
- Throughout this paper, we have presumed that Fe is present because of
-active accretion during the burst. However, the radiative force on
-hydrogenic Fe in the ground state is $F_{\rm rad} = \int {\sigma_{E}
-F_{E}}/{c}\ dE,$ where $F_{E}$ is the flux per photon energy at the
-Ly$\alpha$ transition ($E=6.9\ {\rm keV}$). Since $F_{E}$ is nearly
-constant across the line, $F_{\rm rad}= \pi h e^2 f_{1\rightarrow 2}
-F_E/m_e c^2$ in the optically thin limit (Michaud 1970), where
-$f_{1\rightarrow 2}=0.42$ is the oscillator strength from BS. If $F_E$
-is given by a black-body, then at $kT_{\rm eff} = 1\ {\rm keV}$, the
-radiative acceleration on Fe is $a_{\rm rad} \approx 10^{15} {\rm cm\
-s}^{-2}$, a factor of three larger than $g$. The radiative force will
-be even larger for realistic type I burst spectra, which are harder
-than a blackbody at the observed $T_{\rm eff}$ (London et al. 1986;
-Pavlov, et al. 1991).  The Fe diffusion time is less than the burst
-duration, so there is adequate time for the accelerated Fe to reach a
-new diffusive equilibrium. The calculation of the resulting Fe column,
-which could well saturate the line and lower the radiative force, is
-beyond the scope of this {\it Letter}, as it requires a simultaneous
-solution of the radiative transfer and the Fe diffusion equations,
-like that done for hot white dwarfs (e.g. Chayer, Fontaine \& Wesemael
-1995).
+\section{Discussion and Conclusions}
 
@@
 
- We thank A. Piro for calculating the infall energy and incident
-angles and N. Barghouty for providing the spallation cross-section
-code of Silberberg et al. (1998). We thank the referee, Marten van
-Kerkwijk, for critical comments and Ira Wasserman for many insightful
-conversation and for pointing out the critical role NLTE effects can
-play in line formation. This work was supported by the National
-Science Foundation under grants PHY99-07949 and AST02-05956 and NASA
-through grant NAG 5-8658.  L. B. is a Cottrell Scholar of the Research
-Corporation.
+Write
 
@@
 
-\noindent
-Chayer, P., Fontaine, G., Wesemael, F. 1995, ApJS, 99, 189
-
-\noindent
-Cottam, J., Paerels, F. \& Mendez, M. 2002, \nat, 420, 51 (CPM)
-
-\noindent
-Day, C. S. R., Fabian, A. C. \& Ross, R. R. 1992, MNRAS, 257, 471
-
-\noindent 
-Ferrando, P., et al. 1988, Phys Rev. C, 37, 1490
-
-\noindent
-Foster, A. J., Ross, R. R., \& Fabian, A. C. 1987, MNRAS, 228, 259
-
-\noindent
-Fujimoto, M. Y. 1985, ApJ, 293, L19
-
-\noindent
-Gottwald, M., Haberl, F., Parmar, A. N., \& White, N. E. 1986, \apj,
-308, 213
-
-\noindent
-Grevesse, N. \& Sauval, A. J. 1999, AA, 347, 348
-
-\noindent
-Jefferies, J. T. 1968 \textit{Spectral Line Formation} (Waltham:
-Blaisdell)
-
-\noindent 
-Joss, P. C. 1977, Nature, 270, 310
-
-\noindent 
-Kluzniak, W. \& Wagoner, R. V. 1985, ApJ, 297, 548
-
-\noindent 
-Kluzniak, W., Michelson, P. \& Wagoner, R. V. 1990, ApJ, 358, 538
-
-\noindent 
-Kluzniak, W. \& Wilson, J. R. 1991, ApJ, 372, L87
-
-\noindent
-Kuulkers, E., Homan, J., van der Klis, M., Lewin, W. H. G., \& Mendez,
-M. 2002, A\&A, 382, 947
-
-\noindent
-London, R.~A., Taam, R.~E., \& Howard, W. M. 1984, 287, L27
-
-\noindent
-London, R.~A., Taam, R.~E., \& Howard, W. M. 1986, 306, 170
-
-\noindent
-Madej, J. 1989, A\&A, 209, 226
-
-\noindent 
-Madej, J. 1991, \apj, 376, 161
-
-\noindent 
-Madej, J. 1997, A\&A, 320, 177
-
-\noindent
-Magnier, E., Lewin, W. H. G., van Paradijs, J., Tan, J., Penninx, W. \& Damen, E. 1989, MNRAS, 237, 729
-
-\noindent
-Michaud, G. 1970, \apj, 160, 641
-
-\noindent
-Mihalas, D. 1978, \textit{Stellar Atmospheres}
-(San Francisco: W. H. Freeman)
-
-\noindent
-Nakamura, N., Inoue, H. \& Tanaka, Y. 1988, PASJ, 40, 209
-
-\noindent
-\"Ozel, F., \& Psaltis, D. 2003, \apj, 582, L31
-
-\noindent
-Paerels, F. K. 1997, \apj,  476, 47
-
-\noindent
-Pavlov, G.~G., Shibanov, Y.~A., \& Zavlin, V.~E. 1991, \mnras, 253, 193
-
-\noindent
-Rybicki, G. B. \& Lightman, A. P. 1979 \textit{Radiative Processes in
-Astrophysics} (New York: Wiley)
-
-\noindent 
-Silberberg, R., Tsao, C. H. \& Barghouty, A. F. 1998, Ap. J., 501, 911
-
-\noindent
-Strohmayer, T. E. \& Brown, E. F. 2002, \apj, 566, 1045
-
-\noindent
-Turner, M. J. L. \& Breedon, L. M. 1984, MNRAS, 208, 29p
-
-\noindent
-van Paradijs, J., Dotani, T., Tanaka, Y., \& Tsuru, T. 1990, PASJ, 42,
-633
-
-\noindent 
-Verner, D. A., Ferland, G. J., Korista, K. T., Yakovlev, D. G. 1996,
-ApJ, 465, 487
-
-\noindent
-Waki, I. et al. 1984, PASJ, 36, 819
-
-\noindent
-Zel'dovich, Ya. \& Shakura, N. 1969, Soviet Astron, 13, 175
-
 \end{references}
 
-\begin{deluxetable}{c c c c c c}
-  \tablecolumns{7} \tablewidth{0pt} \tablecaption{Heavy Element Abundances
-  and Atomic Features}
-
-  \tablehead{
-    \colhead{} & \colhead{$E_{\rm e}$} &
-    \colhead{$\tau_{\rm e,max}$} & \colhead{$N_i/N_{\rm Fe}$} &
-    \colhead{$EW_{\rm e}$} & \colhead{$kT_{1/2}$}\\  
-    \colhead{} & \colhead{[keV(redsh)]} &\colhead{}
-    & \colhead{} & \colhead{[eV]} & \colhead{[keV]}}
-  \startdata 
-  Fe & 9.279(6.87) & 0.32 & 1.00 & 277 & 1.12\\ 
-  Mn & 8.573(6.35) & 0.18 & 0.36 & 66 & 1.05\\ 
-  Cr & 7.897(5.85) & 0.26 & 0.48 & 56 & 0.99\\ 
-  V  & 7.248(5.37) & 0.26 & 0.43 & 32 & 0.93\\ 
-  Ti & 6.627(4.91) & 0.31 & 0.48 & 22 & 0.86\\ 
-  Sc & 6.035(4.47) & 0.41 & 0.58 & 17 & 0.81\\ 
-  Ca & 5.471(4.05) & 0.47 & 0.59 & 11 & 0.75\\ 
-  K  & 4.934(3.65) & 0.48 & 0.55 & 7 & 0.69 \\ 
-  Ar & 4.427(3.28) & 0.53 & 0.54 & 4 & 0.65 \\ 
-  \enddata
-\end{deluxetable}
 \end{document}

### 2014-09-18 ms.tex
added more figures and add description of methods

@@
 %\documentstyle[emulateapj,apjfonts,epsfig]{article}
-\documentclass[iop,apjl,letterpaper]{emulateapj}
+\documentclass{emulateapj}
 %\documentclass[12pt,preprint]{aastex}
@@
 
-\title{Atomic Spectral Features During Thermonuclear Flashes on 
-Neutron Stars}
+\title{NEED TITLE}
 
@@
 
-What methods did we use?
+We use the adaptive mesh refinement code FLASH ver. 4.0.1 (Fryxell et al.
+2000
+; Dubey
+et al.
+2008
+) to model isothermal, self-gravitating, hydrodynamic turbulence on isothermal gas with three-dimensional (3D),
+periodic grids and 10 levels of refinement on a root grid of $128^3$, giving an effective resolution of $128K^3$.  
+Self-gravity is computed with a multi-grid Poisson solver (see Ricker
+2008), coupled with a fast-Fourier transform solution on the root grid.
+
+To initialize our simulations, we drive turbulence by applying a large scale ($1 \le k \le 2$) solenoidal 
+acceleration field as a momentum and energy source term.  We apply this field in the absence of gravity and sink particle formation for 3 dynamical times until a statistical steady state is reached.
+{\bf WHY JUST SOLENOIDAL}
+
+This fully developed turbulent state is the initial condition to which we add self-gravity and sink particle formation for
+our star formation experiments.
+Sink particles are formed as described in Lee et al.
+(2014).
+{\bf INCLUDE EQUATION}.
+Note that this is different from the sink particle prescription Federrath et al (???) when additional checks need to be performed.
+
+The gravitational force is computed differently than the gas.
+Sink particle-sink particle forces are computed via direct N-body calculation.
+While sink particle-gas and gas-sink particle is computed via the multi-grid Poisson solver.
+As a result of these additional computations, two large scale gravity solutions must be found per timestep as oppose to one.
+This allows to avoid the computationally expensive task of computing gas-sink particle forces via direct summation.
 
@@
 -mass vs r
+
+\begin{figure}
+\plotone{frame0241.png}
+\end{figure}
+
+\begin{figure*}
+\plottwo{movie_disk_frame_0259.png}{movie_disk_frame_0257.png}
+\plottwo{movie_disk_frame_0256.png}{movie_disk_frame_0272.png}
+\end{figure*}
 

### 2014-09-18 ms.tex
added analytic expectations

@@
 \documentclass{emulateapj}
+\usepackage{amsmath, mathtools}
+
 %\documentclass[12pt,preprint]{aastex}
@@
 \newcommand{\Ca}{${\rm Ca} \ $}
+\newcommand{\be}{\begin{equation}}
+\newcommand{\ee}{\end{equation}}
+\newcommand       \etaeff       {\eta}
+\newcommand       \tff          {\tau_{\rm ff}}
+\newcommand       \tdyn         {\tau_{\rm dyn}}
 
@@
 \section{Analytic Expectations}
-\section{Comparison with Numerical Results}
+
+Recently, Murray \& Chang (2014) considered the spherical collapse of isothermal turbulently supported gas accounting for adiabatic heating from compression of the turbulence.
+In particular,  {\bf SHOW EQUATIONS}
+Their basic results are that the run of density asymptotes to 
+\be
+\rho(r,t)=
+\begin{dcases}
+\rho(r_0)\left({r\over r_0}\right)^{-3/2}, & r<r_*\\
+\rho(R,t)\left({r\over R}\right)^{-k_\rho}, \ k_\rho\approx1.6-1.8 & r>r_*.
+\end{dcases}
+\ee
+Since $k_\rho$ is fairly close to $1.5$ at all radii, taking $r_0=R$
+is a fair approximation. 
+The infall velocity 
+%
+\be
+u_r(r,t)=
+\begin{dcases}
+-\Gamma\sqrt{GM_*(t)\over r}, \sim r^{-1/2} & r<r_*\\
+-\Gamma\sqrt{GM(r,t)\over r} \sim r^{0.2} & r>r_*,
+\end{dcases}
+\ee
+%
+where $\Gamma \approx 0.7$ at small radii, and $\Gamma\approx 1.0$ at
+large radii.
+
+The turbulent velocity 
+%
+\be
+v_T(r,t)=
+\begin{dcases}
+{1\over 2\etaeff}\Gamma\sqrt{GM_*(t)\over r}, \sim r^{-1/2} & r<r_*\\
+{1.2\over \etaeff}\Gamma\sqrt{GM(r,t)\over r} \sim r^{0.2} & r>r_*,
+\end{dcases}
+\ee
+%
+
+The stellar mass increases quadratically with time
+%
+\be  %$
+M_*(t)=\phi M_{\rm cl}\left({t-t_*\over \tff}\right)^2.
+\ee  %$
+%
+
+The mass accretion rate 
+%
+\be
+\dot M(r,t)=
+\begin{dcases}
+4\pi R^2\rho(R)u_r(r,t), \sim t\,r^{0} & r<r_*\\
+4\pi R^2\rho(R)u_r(r,t) \sim t^0\,r^{0.2} & r>r_*.
+\end{dcases}
+\ee
+%
+
+\section{Numerical Results}
 
@@
 -mass vs r
+-mdot vs r
 

### 2014-09-23 ms.tex
added further description in numerical results, captions and graph for rho

@@
 
-\author{Lars Bildsten\altaffilmark{1}, Philip Chang\altaffilmark{2},
-and Frits Paerels\altaffilmark{3}}
-\altaffiltext{1}{Kavli Institute for Theoretical Physics,  Kohn Hall, University
-of California, Santa Barbara, CA 93106, USA; email:
-bildsten@kitp.ucsb.edu}
-\altaffiltext{2}
-{Department of Physics, Broida Hall, University of California,
-Santa Barbara, CA 93106; pchang@physics.ucsb.edu}
-\altaffiltext{3}
-{Columbia Astrophysics Laboratory and Department of Astronomy,
-Columbia University, 538 W. 120th St., New York, NY 10027;
-frits@astro.columbia.edu}
+\author{some people}
 
@@
 
+{\bf include discussion of mesh refinement}
+
+{\bf include discussion of star formation.}
+
+To initialize our simulation, we begin by applying solenoidal stirring forces to the uniform, periodic simulation volume at the root grid resolution of $128^3$ until statistical equilibrium is reached (approximately three crossing times).
+After this equilibrium is established, we then apply self gravity, adaptive mesh refinement, and sink particle creation.
+In Figure \ref{fig:entire projection} we show a projection along the x-axis of the entire simulation volume that has up to 10 levels of refinement, giving an effective resolution of $128\,{\rm K}^3$.
+Regions that are highly refined are the densest regions which are smoother than the low-density more pixelated regions.
+
+The high density regions appear to be organized along filaments.
+These filaments span most of the simulation size, and have a width of 0.5 to 1 pc.
+Moreover, these filaments appear to flow into large clumps.
+This is in line with previous work including Lee et al.
+2014.
+
+The simulated volume has a large number of refined regions.
+We show four of them in Figure \ref{fig:snapshot} as slices perpendicular to the angular momentum vector.
+We determine the angular momentum vector by taking a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
+
+These figures suggest that the accretion can occur by streams or by spherical flows.
+However, the degree by which these modes of accretion dominate, is not clear.
+At small radii (similar to $10^{-2}$ pc), a protostellar disk appears in all four snapshots.
+The radial size of these disks varies.
+These disks also show a coherent sense of rotation about a common axis, e.g., no counter-rotating disks.
+In addition, some of the snapshots show multiple high density regions.
+This could be a result of accretion of high density clumps or fragmentation in the disk or streams.
+
+These simulations display much more complicated dynamics than is captured by the simple analytic theory of Murray and Chang 2014.
+In spite of this, we compare radially averaged profiles around these high density points to their analytic theory.
+In fig
+
+However, to make a comparison of the average behavior of high density star forming regions, we average over a number
+
 What plots should we show?
@@
 \plotone{frame0241.png}
+\caption{Projection along the x-axis of the entire simulation volume.
+The root grid is $128^3$ with up to 10 levels of refinement, giving an effective resolution of $128\,{\rm K}^3$.
+This snapshot is taken at ??.
+\label{fig:entire projection}}
 \end{figure}
@@
 \plottwo{movie_disk_frame_0256.png}{movie_disk_frame_0272.png}
+\caption{Slices along the angular momentum axis of four zoomed in simulation.
+To determine the angular momentum axis, we take a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
+We then take a slice in the plane normal to the angular momentum to plot the regions around these dense points.
+\label{fig:snapshots}}
 \end{figure*}
@@
 \begin{figure}
+\plotone{Track_0256_000_rho.png}
+\caption{Density as a function of radial position from maximum density point.
+The density reaches a maximum of $10^6$ times mean density at $r \approx 10^{-3}$ pc and decline to approximately mean density at $r\approx 3$ pc.
+The three straight solid lines are $r^{-3/2}$, $r^{-2}$, and $r^{-5/2}$ power laws.
+}
+\end{figure}
+
+\begin{figure}
 \plotone{Track_0132_000_rho.png}
+\caption{Density as a function of radial position from maximum density point.
+The density reaches a maximum of $10^7$ times mean density at $r \approx 10^{-3}$ pc and decline to approximately mean density at $r\approx 3$ pc, a }
 \end{figure}

### 2014-09-25 ms.tex
Added figures, continued discussion of numerical results

@@
 Regions that are highly refined are the densest regions which are smoother than the low-density more pixelated regions.
+\begin{figure}
+\plotone{frame0241.png}
+\caption{Projection along the x-axis of the entire simulation volume.
+The root grid is $128^3$ with up to 10 levels of refinement, giving an effective resolution of $128\,{\rm K}^3$.
+This snapshot is taken at ??.
+\label{fig:entire projection}}
+\end{figure}
 
@@
 
+\begin{figure*}
+\plottwo{movie_disk_frame_0259.png}{movie_disk_frame_0257.png}
+\plottwo{movie_disk_frame_0256.png}{movie_disk_frame_0272.png}
+\caption{Slices along the angular momentum axis of four zoomed in simulation.
+To determine the angular momentum axis, we take a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
+We then take a slice in the plane normal to the angular momentum to plot the regions around these dense points.
+\label{fig:snapshots}}
+\end{figure*}
+
 These figures suggest that the accretion can occur by streams or by spherical flows.
@@
 In spite of this, we compare radially averaged profiles around these high density points to their analytic theory.
-In fig
+In the left panel of Figure \ref{fig:good_rho_example}, we show the radial density profile around one such high density point.
+As this figure shows, the radial density profile follows a $r^{-3/2}$ power law in agreement with analytic expections.
+This power-law breaks down at small radii, but this is not unexpected as this radii corresponds to the presents of a rotationally support disk.
+Such good agreement is not universal for all high density points as shown in the right panel of Figure \ref{fig:bad_rho_example}.
+Here the radial density profile follows a $r^{-2}$ power law.
+\begin{figure*}
+\plottwo{Track_0256_000_rho.png}{Track_0132_000_rho.png}
+\caption{Density as a function of radial position from maximum density point.
+The density reaches a maximum of $10^6$ times mean density at $r \approx 10^{-3}$ pc and decline to approximately mean density at $r\approx 3$ pc.
+The three straight solid lines are $r^{-3/2}$, $r^{-2}$, and $r^{-5/2}$ power laws.
+\label{fig:good_rho_example}}
+\end{figure*}
 
@@
 What plots should we show?
--density vs r
 -velocities vs r
@@
 
-\begin{figure}
-\plotone{frame0241.png}
-\caption{Projection along the x-axis of the entire simulation volume.
-The root grid is $128^3$ with up to 10 levels of refinement, giving an effective resolution of $128\,{\rm K}^3$.
-This snapshot is taken at ??.
-\label{fig:entire projection}}
-\end{figure}
+In Figure \ref{fig:vr_example}, we plot $v_r$ (blue line) and $v_{\rm rms}$ (green line) as a function of $r$.
+In the left panel, the radial velocity curve is sporadic, but the general trend is that $v_r$ declines in toward $0.1$ pc, but then rises inward of this point.
+Plotted in red is a $r^{-1/2}$ power law and this inward rise appears to follow this general trend.
+The turbulent velocity is smoother, but it follows the general trend of the $v_r$ curve in that it falls slowly inward toward $]\approx 0.1 $ pc before rising inward of that, roughly tracking the $r^{-1/2}$ power law.
+These general trends are in rough agreement with our analytic expectations.
+
+{\bf DISCUSS RIGHT PANEL}
 
 \begin{figure*}
-\plottwo{movie_disk_frame_0259.png}{movie_disk_frame_0257.png}
-\plottwo{movie_disk_frame_0256.png}{movie_disk_frame_0272.png}
-\caption{Slices along the angular momentum axis of four zoomed in simulation.
-To determine the angular momentum axis, we take a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
-We then take a slice in the plane normal to the angular momentum to plot the regions around these dense points.
-\label{fig:snapshots}}
+\plottwo{Track_0256_000_vr.png}{Track_0132_000_vr.png}
+\caption{Radial velocity, $v_r$ (blue line), and turbulent velocity, $v_{\rm rms}$ (green line) as a function of radial position from maximum density point.
+A $r^{-1/2}$ power law is shown as a red line.
+\label{fig:vr_example}}
 \end{figure*}
 
-\begin{figure}
-\plotone{mass.png}
-\end{figure}
+A hint of why these different behaviors emerge from looking at the $M(r)$.
+In Figure \ref{fig:mass_example}, we plot the mass as a function of radius.
+Regions where $M(r)$ flattens implies that the mass in this region can be treated using a point mass approximation.
+Here it is apparent that the regions that can be treated a point mass corresponds to regions in where the density follows a $r^{-3/2}$ power law and the radial and turbulent velocities follow a $r^{-1/2}$ power law.
 
-\begin{figure}
-\plotone{Track_0256_000_rho.png}
-\caption{Density as a function of radial position from maximum density point.
-The density reaches a maximum of $10^6$ times mean density at $r \approx 10^{-3}$ pc and decline to approximately mean density at $r\approx 3$ pc.
-The three straight solid lines are $r^{-3/2}$, $r^{-2}$, and $r^{-5/2}$ power laws.
-}
-\end{figure}
-
-\begin{figure}
-\plotone{Track_0132_000_rho.png}
-\caption{Density as a function of radial position from maximum density point.
-The density reaches a maximum of $10^7$ times mean density at $r \approx 10^{-3}$ pc and decline to approximately mean density at $r\approx 3$ pc, a }
-\end{figure}
+\begin{figure*}
+\plottwo{Track_0256_000_mass.png}{Track_0132_000_mass.png}
+\caption{WRITE CAPTION: Radial velocity, $v_r$ (blue line), and turbulent velocity, $v_{\rm rms}$ (green line) as a function of radial position from maximum density point.
+A $r^{-1/2}$ power law is shown as a red line.
+\label{fig:mass_example}}
+\end{figure*}
 

### 2014-09-30 ms.tex
Added ang mom graphs, mass caption, and discussion on vr/vrms

@@
 \caption{Density as a function of radial position from maximum density point.
-The density reaches a maximum of $10^6$ times mean density at $r \approx 10^{-3}$ pc and decline to approximately mean density at $r\approx 3$ pc.
+The density reaches a maximum of $10^6$ times mean density at $r \approx 10^{-3}$ pc and declines to approximately mean density at $r\approx 3$ pc.
 The three straight solid lines are $r^{-3/2}$, $r^{-2}$, and $r^{-5/2}$ power laws.
-\label{fig:good_rho_example}}
+\label{fig:good_rho_example}
+{bf\ This discusses left panel only, should we move the discussion outside the caption?}}
 \end{figure*}
@@
 These general trends are in rough agreement with our analytic expectations.
+In the right panel, the radial velocity curve is smoother and increases inwards towards $0.01$ pc, tracking the $r^{-1/2}$ power law fairly closely.
+This radial velocity begins to fall inward of $0.01$ pc, possibly due to rotational support from the angular momentum of a disk.
+The turbulent velocity in the right panel seems to roughly follow an inverse relationship with the radial velocity.
+As with the left panel, however, the turbulent velocity is significantly less volatile than the radial velocity and varies only slightly with $r$.
 
@@
 \plottwo{Track_0256_000_mass.png}{Track_0132_000_mass.png}
-\caption{WRITE CAPTION: Radial velocity, $v_r$ (blue line), and turbulent velocity, $v_{\rm rms}$ (green line) as a function of radial position from maximum density point.
-A $r^{-1/2}$ power law is shown as a red line.
-\label{fig:mass_example}}
+\caption{WRITE CAPTION: Mass, $MSun$, as a function of radial position from a local maximum density point.     \label{fig:mass_example}}
 \end{figure*}

### 2014-09-30 ms.tex
included angular momentum

@@
 
-\section{Numerical Results}
+\section{Detailed Simulations of Turbulent Collapse}
 
@@
 This fully developed turbulent state is the initial condition to which we add self-gravity and sink particle formation for
-our star formation experiments.
-Sink particles are formed as described in Lee et al.
-(2014).
-{\bf INCLUDE EQUATION}.
-Note that this is different from the sink particle prescription Federrath et al (???) when additional checks need to be performed.
+our star formation experiments. Sink particles are formed as described in Lee et al. (2014) and described briefly below.
 
@@
 
-{\bf include discussion of mesh refinement}
+We have also implemented a new algorithm for mesh refinement in these simulations.
+As gas collapse under self gravity, certain regions rapidly increase in density.
+These regions are refine when the Truelove criterion ($\lambda \le 4 \Delta x$) is met.
+This corresponds to a condition on the density
+\begin{equation}
+refinement criteria
+\end{equation}
+which when met causes the local grid to be refined by a factor of 2 provide that the maximum refinement level is not reached.
 
-{\bf include discussion of star formation.}
+When the Truelove criterion is met at highest refinement level, the excess mass in a cell is transferred either to a newly created sink particle or to a sink particle whose accretion radius includes the cell.
+This behavior is the same as in Lee et al.
+(2014), albeit at a much higher resolution.
+We should also note that like Lee et al.
+(2014), our sink particle creation prescription is different from the prescription of Federrath et al (???) where additional checks need to be performed.
+
+\section{Numerical Results}
 
@@
 
+\subsection{Examples of Collapse}
+
 These simulations display much more complicated dynamics than is captured by the simple analytic theory of Murray and Chang 2014.
@@
 
-{\bf DISCUSS RIGHT PANEL}
-
 \begin{figure*}
@@
 Here it is apparent that the regions that can be treated a point mass corresponds to regions in where the density follows a $r^{-3/2}$ power law and the radial and turbulent velocities follow a $r^{-1/2}$ power law.
+This is due to mass concentration at smaller radii as evident by the sudden drop in $M(r)$ at $r \approx 10^{-2}$ pc.
 
@@
 \end{figure*}
+
+\begin{figure*}
+\plottwo{Track_0256_000_l.png}{Track_0132_000_l.png}
+\caption{WRITE CAPTION: Mass, $MSun$, as a function of radial position from a local maximum density point.     \label{fig:angular_momentum_example}}
+\end{figure*}
+
+\subsection{Average Profiles}
 

### 2014-10-28 Turbulent_collapse.tex
added disks

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End:

%\documentclass[iop,apj,letterpaper]{emulateapj}
\documentclass[iop,apj,numberedappendix]{emulateapj}

%\documentclass[12pt,manuscript]{aastex}
\usepackage[left=1in,top=1in,bottom=1in,right=1in]{geometry}
\usepackage{natbib}
%\usepackage{lscape}
\usepackage{url}
\usepackage{amsmath, mathtools}
% for corrections
\usepackage[normalem]{ulem}
\usepackage{color}
\bibliographystyle{apj}

\newcommand       \phil[1]      {{\color{blue} #1}}

\newcommand       \be		{\begin{equation}}
\newcommand       \ee		{\end{equation}}
\newcommand       \bea          {\begin{eqnarray}}
\newcommand       \eea          {\end{eqnarray}}

\newcommand       \kpc		{\,{\rm kpc \,}}
\newcommand       \pc		{\,{\rm pc \,}}
\newcommand       \erg		{\,{\rm erg \,}}
\newcommand       \kb		{\,{\rm k_b \,}}
\newcommand       \yr		{\,{\rm yr \,}}
\newcommand       \Myr		{\,{\rm Myr \,}}
\newcommand       \cm		{\,{\rm cm \,}}
\newcommand       \s		{\,{\rm s \,}}
\newcommand       \g		{\,{\rm g \,}}
\newcommand       \kms          {\,{\rm km \,\, s}^{-1}}
\newcommand       \eff          {\epsilon_{\rm ff}}
%\newcommand       \etaeff       {\eta_{\rm eff}}
\newcommand       \etaeff       {\eta}
\newcommand       \tff          {\tau_{\rm ff}}
\newcommand       \tdyn         {\tau_{\rm dyn}}
\newcommand       \tDyn         {\tau_{\rm DYN}}
\newcommand       \tDYN         {\tau_{\rm DYN}}
%\newcommand       \R           {R}
\newcommand       \R           {r_0}

\newcommand       \trho 	{{\tau_{\rho}}}
\newcommand       \tvr  	{{\tau_{u_r}}}
\newcommand       \tvt  	{{\tau_{v_T}}}
\newcommand       \vt		{v_{T}}
\newcommand       \mybeta	{\,{\beta}}
\newcommand       \xbv  	{{\left(x-\mybeta v\right)}}

\voffset 0.25in

\begin{document}

\title{Star formation in self-gravitating turbulent fluids}
\author{Norman Murray\altaffilmark{1,2} \& Philip Chang\altaffilmark{3},
  } 
\altaffiltext{1}{Canadian Institute for Theoretical Astrophysics, 60
  St.George Street, University of Toronto, Toronto ON M5S 3H8, Canada;
  murray@cita.utoronto.ca} 
\altaffiltext{2}{Canada Research Chair in Astrophysics}
\altaffiltext{3}{Department of Physics, University of
  Wisconsin-Milwaukee, 1900 E. Kenwood Blvd., Milwauke, WI 53211, USA;
  chang65@uwm.edu}

\begin{abstract}
We present a model of star formation in self-gravitating turbulent
gas. We treat the turbulent velocity $v_T$ as a dynamical variable,
and assume that it is adiabatically heated by the collapse. The theory
predicts the run of density, infall velocity, and turbulent velocity,
and the rate of star formation in compact massive gas clouds. The
turbulent pressure is dynamically important at all radii, a result of
the adiabatic heating. The system evolves toward a coherent spatial
structure with a fixed run of density, $\rho(r,t)\to\rho(r)$; mass
flows through this structure onto the central star or star cluster.
We define the sphere of influence of the accreted matter by
$m_*=M_g(r_*)$, where $m_*$ is the stellar plus disk mass in the
nascent star cluster and $M_g(r)$ is the gas mass inside radius
$r$. The density is given by a broken power law with a slope $-1.5$
inside $r_*$ and $\sim -1.6$ to $-1.8$ outside $r_*$. Both $v_T$ and
the infall velocity $|u_r|$ {\em decrease} with decreasing $r$ for
$r>r_*$; $v_T(r)\sim r^p$, the size-linewidth relation, with
$p\approx0.2-0.3$, explaining the observation that Larson's Law is
altered in massive star forming regions. The infall velocity is
generally smaller than the turbulent velocity at $r>r_*$. For $r<r_*$,
the infall and turbulent velocities are again similar, and both {\em
  increase} with decreasing $r$ as $r^{-1/2}$, with a magnitude about
half of the free-fall velocity. The accreted (stellar) mass grows
super-linearly with time, $\dot M_*=\phi M_{\rm cl}(t/\tff)^2$, with
$\phi$ a dimensionless number somewhat less than unity, $M_{\rm cl}$
the clump mass and $\tff$ the free-fall time of the clump. We suggest
that small values of $p$ can be used as a tracer of convergent
collapsing flows.
\end{abstract}

\section{Introduction}
\label{sec:intro}

The importance of massive stars in a human context became apparent 
when \citet{1957RvMP...29..547B} showed that such stars were
responsible for the production and distribution of most of the heavy
elements that composed the Earth, and which form the building blocks
of life.

More recently it has emerged that massive stars are important
on a less parochial scale. Star formation on galactic scales is
observed to be slow, in the sense that the time to deplete the supply
of gas from which stars are made is some fifty times the dynamical
time. Quantitatively,
%
\be \label{eqn: Schmidt Kennicutt} %$
\dot M_*=\eta {M_g\over \tDyn},
\ee  %$
%
where $\eta\approx0.017$ is a dimensionless constant, and
$\tDyn=R/v_c$ is the dynamical time of a galactic disk of radius $R$
with circular velocity $v_c$
\citep{1998ApJ...498..541K,2008AJ....136.2782L}. Massive stars provide
the feedback that limits the rate of star formation in galaxies. The
feedback comes in the form of radiation pressure, gas pressures in
ionized regions, blast waves from supernovae, and cosmic rays. The
feedback also drives gas out of galaxies, setting the stellar
mass-halo mass relationship for halo masses
$M_h\lesssim10^{13}M_\odot$, e.g.,
\citet{2013arXiv1311.2073H,2014arXiv1404.2613A,2014MNRAS.442.1545C}. Understanding
how massive stars form is clearly important to our understanding of
the evolution of the universe, and many of its components.

Stars more massive than sixty solar masses ($M*>60M_\odot$) live only
$4\Myr$. Since we see stars with masses in excess of $60M_\odot$ outside
their natal dense gas clumps, the early accretion rate onto such a
star must exceed $\approx3\times10^{-5}M_\odot\yr^{-1}$, where we have
assumed that the most massive stars are visible for $\gtrsim1\Myr$, hence
they can accrete for no more than $3\Myr$. In the Milky Way, these
massive stars are usually found in compact $1-10\pc$ radius clusters
containing up to $10^5M_\odot$ of stars.  The accretion rates for
clusters must exceed a few hundredths of a solar mass per year, $\dot
M_*\gtrsim 3\times10^{-2}M_\odot\yr^{-1}$. For example, Cygnus OB2
contains $M_*\gtrsim 4-10\times10^4M_\odot$ inside a half light radius
$r_h=6.4\pc$ \citep{2000A&A...360..539K}, yielding $\dot M_*\approx
0.02M_\odot\yr^{-1}$. The free-fall time of the natal cluster is
%
\bea  %$
\tff&\equiv&\sqrt{3\pi\over 32G\rho}\nonumber\\
&=&1\left({3.5\times10^{4}M_\odot\over M_*}\right)^{-1/2}
\left({R_{1/2}\over6.4\pc}\right)^{3/2}
\left({\epsilon\over0.5}\right)\Myr,
\eea  %$
%
where all quantities are estimated at the half light radius, and
$\epsilon\equiv M_*/M_g$ is the fraction of natal gas in a dense
cluster-forming region that is converted into stars. The star
formation rate per free fall time, the small-scale analog of $\eta$ in
eqn (\ref{eqn: Schmidt Kennicutt}), is
%
\bea  %$
\eff&\equiv&{\dot M_*\over \tff M_g}
={\tff\over 3\Myr}\epsilon\nonumber\\
&\approx&0.16
\left({R_{1/2}\over 6.4\pc}\right)^{3/2}
\left({M_{*,1/2}\over 3.5\times10^4M_\odot}\right)^{1/2}
\left({\epsilon\over 0.5}\right)^{1/2}.
\eea  %$
%

This estimate of $\eff$ is consistent with many estimates in the
literature, for example, \citet{2005ApJ...635L.173W}, \citet{2010ApJ...724..687L},
\citet{2010ApJ...723.1019H}, and
\citet{2011ApJ...729..133M}, although there is some disagreement, e.g.,
\citep{2012ApJ...745...69K}.

The fact that $\eff>>\eta$ is one of the reasons that astronomers
believe that stellar feedback is necessary to explain the slow global
rate of star formation; if $\eff=\eta$, no feedback would be
required. A second motivation for feedback is that simulations on the
scale of galaxies, and on the scale of star forming clumps ($r\sim
1-10\pc$) both find very rapid star formation if no feedback is included.

In this paper we are interested in the physics behind rapid star
formation rather than the physics of feedback. The problem arises from
the high accretion rates, discussed above, required to make massive
stars. The early model of \citet{1977ApJ...214..488S} gave an estimate
of the star formation rate. His model assumed that stars formed from
hydrostatic cores supported by thermal gas pressure. The accretion
rate in his model was independent of time, given by $\dot
M_*=m_0c_s^3/G$, where $c_s\approx0.2\kms$ is the sound speed in
molecular gas, and $m_0=0.975$. Thus the maximum accretion rate was
\phil{\sout{$\dot M_*\approx2\times10^6M_\odot\yr^{-1}$}$\dot M_*\approx2\times10^{-6}M_\odot\yr^{-1}$}, too small by a factor of
ten to explain the origin of massive stars.

The difficulty with the accretion rate was overcome by
\citet{1992ApJ...396..631M}, who noted that cores in massive star
forming regions had linewidths that exceeded the thermal line width by
factors of several or more. They pointed out that the accretion rate
in Shu's model was limited by the signal speed of the expanding
collapse wave, then took advantage of the larger signal speed
available if they assumed that the pressure in the fluid was given by
%
\be \label{eqn: turbulent pressure}%$
P(r,t)=\rho(r,t) \left[c_s^2+ v_T^2(r)\right],
\ee %$
%
where $\rho(r,t)$ is the local density, $c_s$ is the sound speed, and
$v_T(r)$ is the turbulent velocity a distance $r$ from the center of
the collapse.

The model was further developed by \citet{1997ApJ...476..750M} and
\citet{2003ApJ...585..850M}. These models kept some of the features of
Shu's model, including the assumption of a hydrostatic initial core,
and the static nature of the turbulent velocity in the equation of
state (\ref{eqn: turbulent pressure}). In particular, they assumed
that $v_T$ was not affected by the collapse, and that it was
independent of time. They did not address the question of how a
hydrostatic core was assembled.

Observations in large GMCs, and low mass cores, find 
%
\be \label{eqn: turbulent} %$
v_T(r)=v_T(R)\left({r\over R}\right)^p
\ee  %$
%
with $p\approx 0.5$ \citep{1981MNRAS.194..809L, 1988ApJ...329..392M};
as we will discuss further below, in massive star forming regions
$v_T$ behaves differently, with $p\approx0.2$ and much larger
turbulent velocities at a given separation $r$ than in low mass star
forming regions. The turbulent core models either explicitly assume at
turbulent velocity of the form (\ref{eqn: turbulent}), or assumed
implicitly that it follows a power law in radius so that $P(r)\sim
\rho^\gamma(r)$.

In a companion paper \citep{2014arXiv1406.4148L} (Paper I) we present
three dimensional hydro and MHD simulations which recover the result
$p\approx 0.5$ in the bulk of the gas. However, we also show that in
small regions around density peaks, which we identify as cores,
$p\approx 0.2$.

In this paper we assume that massive stars form in regions of
turbulent flows in which the gas is converging, so that there is never
any time at which the region is in hydrostatic equilibrium. The infall
therefore extends over a much larger region than that envisioned in the
hydrostatic turbulent core models.

A more profound difference arises from the fact that we model
the turbulent velocity $v_T(r,t)$ as a dynamical variable, changing in
response to the evolution of the system. This requires that we
introduce a third dynamical equation, the energy equation. However,
we are unable to solve the equations that result. This is not
surprising, because numerical solutions of the full set, carried out
in three dimensions, show that the motion becomes turbulent.

Suitably chastened, we then make two simplifications. First, we assume
spherical symmetry, reducing the problem from three spatial dimensions
to one. Second, we introduce a closure for the energy equation, that
proposed by \citet{2012ApJ...750L..31R}.

We import the notion of the sphere of influence of the star from
galactic dynamics, and define the radius $r_*(t)$ at which the mass of
gas just equals the (time dependent) mass of the star. With our two
assumptions, and the concept of the sphere of influence, we are able
to find self-similar solutions that describe the collapse of a
self-gravitating turbulent gas. We calculate the run of density
$\rho(r,t)$, of infall velocity $u_r(r,t)$, and of the turbulent
velocity $v_T(r,t)$. We find the striking result that for $r<r_*(t)$,
the density approaches a fixed function of radius, independent of
time, i.e., $\rho(r,t)\to \rho_*(r)\sim r^{-3/2}$. For $r>r_*$,
$\rho(r,t)\sim r^{-k_\rho}$, with $k_\rho\approx 1.6-1.8$.

We show that the acceleration due to the pressure gradient always
tracks the acceleration of gravity. For $r<r_*(t)$, the result is that
both the infall velocity and the turbulent velocity increase with
decreasing radius, as $r^{1/2}$. But for $r>r_*(t)$, the pressure
force exceeds that due to gravity. As a result, both $v_T(r,t)$ and
$u_r(r,t)$ increase slowly with increasing radius, $v_T(r,t)\sim r^p$,
with $p\approx 0.2$. This appears to explain the deviations from
Larson's size-linewidth relation seen in massive star forming regions.

It follows from the fact that the density is independent of time for
$r<r_*$, and from $u_r(r,t)\sim\sqrt{GM_*/r}$\phil{$\propto t$}, that the mass accretion
rate increases linearly with time, so that
%
\be  %$
M_*=\phi M_{cl}\left({t-t_*\over \tff}\right)^2,
\ee  %$
%
where $M_{cl}$ is the mass of the star forming clump, and $\tff$ is
the free fall time of the clump.

This paper is organized as follows.  In \S \ref{sec: collapse} we
present a spherically symmetric one dimensional model for the collapse
of a turbulently supported gas cloud consisting of the continuity and
momentum equations, and a simple closure for the energy equation. We
describe approximate analytic solutions to these equations, and
comment on the effects of thermal and magnetic support.  In \S
\ref{sec: numerics} we solve the equations numerically. In \S
\ref{sec:discussion} we briefly compare observations of infall in the
massive star forming regions, with special attention given to
G10.6-0.4. We also discuss previous analytic and numerical work.  We
wrap up with our conclusions in the final section.

\section{SPHERICAL GRAVITATIONAL COLLAPSE MODELS}
\label{sec: collapse}
The governing equations are the continuity equation
%
\be  %$
{\partial\rho\over\partial t}+{\bf \nabla}\cdot (\rho{\bf u}) =0,
\ee  %$
%
the momentum equation
%
\be  %$
{\partial \rho{\bf u}\over \partial t}+ \nabla\cdot \rho{\bf u u}
= -{\bf \nabla} P + \rho{\bf g},
\ee  %$
%
and the energy equation
%
\bea \label{eqn: energy} %$
{\partial\over \partial t}\left[\rho\left({1\over2}u^2
+ {\cal E}\right) \right]
&+& \nabla\cdot\left[
\rho\left(
{1\over2}u^2 + {\cal E}
\right)\cdot {\bf u}
+ {\bf u}P \right]\nonumber\\
&\qquad=& \rho {\bf g}\cdot{\bf u}.
\eea  %$
%
In these equations $\rho$ is the density, ${\bf u}$ is the fluid
velocity, $P$ is the gas pressure, ${\bf g}$ is the acceleration due
to gravity, and ${\cal E}$ is the internal energy of the gas.

%\subsection{A simplified energy equation}
In the companion paper \phil{(Paper I),} we solve these equations via a three
dimensional simulation. In this paper, we \phil{\sout{ want to}} simplify our model
\phil{\sout{so as to be able}} to gain some physical insight into the collapse
process. To do so we will simplify the energy equation. We start by
noting that simulations of both supersonic and subsonic turbulence
have generally found that the properties of the turbulent flow are
universal, in the sense that certain relations are found under a wide
variety of conditions. In particular, simulations find velocity power
spectra that have approximate power law forms, e.g., $P_v(k)\sim
k^{-5/3}$ for subsonic tubulence or $P_v(k)\sim k^{-2}$ for supersonic
turbulence. Correspondingly, $v_T(r)\sim r^{1/3}$ or $v_T(r)\sim
r^{1/2}$.

It is tempting to conclude that the energy equation 
can be replaced by a simple closure relation specifying the
turbulent velocity, namely $v_T(r,t)=v_T(r)$, where $v_T(r)$ is given
by Larson's law. This is what has been done in most previous analytic
work on turbulent collapse.

This temptation should be resisted.  
\phil{Over an eddy turnover time $r/v_T(r)$, $v_T$ will decay due to
  dissipation.  At the same time, \citet{2012ApJ...750L..31R} showed
  that compression of turbulent gas increases $v_T$. Turbulent gas
  that falls in will both be compressed, increasing $v_T$, while $v_T$
  will also decay due to dissipation.  }
{\phil Therefore, we} use a closure scheme suggested by
\citet{2012ApJ...750L..31R} \phil{that captures both processes}. In the absence of collapse, undriven
turbulence is seen to decay on an eddy turnover time:
%
\be  \label{eqn: decay}%$
{1\over2}{dv_T^2(r)\over dt} = -\eta {v_T^3\over r},
\ee  %$
%
with $\eta $ a dimensionless constant of order unity. In many
simulations of star formation, the turbulence is driven at large
scales; this energy then cascades down to smaller scales, maintaining
$v_T(r,t)\approx v_T(r)$, in the absence of self-gravity.

Collapse alters the turbulence. Ignoring for the moment the cascade of
energy from large scales to small that is responsible for
(\ref{eqn: decay}), a reduction in the radius from $R$ to $r$ results in a
scaling of the velocity $v_T(r) = v_T(R)\times (R/r)$, i.e.,
%
\be  %$
\left({dv_T\over dt}\right)_{\rm AH}=-v_T{u_r\over r},
\ee  %$
%
where $u_r=dr/dt$ is negative for a collapse. Thus the scaling relation suggests
that a collapse will tend to increase the turbulent
velocity. Robertson \& Goldreich refer to this as ``adiabatic
heating'', and show that such heating occurs in cosmological
simulations in which the scale factor decreases with time, mimicking a
gravitational collapse.

Combining this driving term with the turbulent damping,
%
\be \label{eqn: simple energy} %$
{d v_T\over dt} = 
-\left(1+\etaeff {v_T\over u_r}\right){v_Tu_r\over r}.
\ee  %$
%
We use this equation to replace the energy Equation (\ref{eqn: energy}).

\subsection{Fundamental Equations}
In \phil{\sout{the companion paper} Paper I} we show that on parsec and smaller scales, the
collapse is, very roughly, spherically symmetric\phil{\sout{,}. Motivated by this and our desire to simplify our model \sout{ so},} we employ spherical
coordinates. The continuity and momentum equations become
%
\be \label{eqn: continuity}%$
{\partial\rho\over \partial t}+{1\over r^2}{\partial\over\partial r}
\left(r^2u\rho\right) = 0,
\ee %$
%
and
%
\be \label{eqn: momentum}%$
{\partial u_r\over \partial t}+u_r{\partial u_r\over \partial r}
+{1\over \rho}{\partial \rho (v_T^2(r,t)+c_s^2)\over \partial r}+{GM(r,t)\over r^2}=0,
\ee %$
%
while the energy equation is
%
\be \label{eqn: 1D energy}%$
{\partial v_T\over \partial t}+ u_r{\partial v_T\over\partial r} +
\left(1+\etaeff {v_T\over u_r}\right){v_Tu_r\over r}=0.
\ee %$
%

In these equations $u_r(r,t)$ is the mean radial velocity of the
fluid, excluding the turbulent velocity, which is denoted by
$v_T(r,t)$. In the rest of this section only, we will neglect the sound
speed $c_s$ to simplify the presentation. \phil{In any case, we find that its inclusion has little effect on the results. \sout{It has little effect on the results.}}

The mass $M(r,t)=M(0,t)+M_g(r,t)$ where
%
\be \label{eqn: mass}%$
M_g(r,t)= 4\pi\int_{0^+}^r r^2\rho(r,t)dr
\ee %$
%
is the gas mass.

We will employ $M_*(t)$ rather than $M(0,t)$, with the
understanding that $M_*$ may refer to the mass of a single star or to
a star cluster. For later use we define the radius or sphere of
influence of the nascent star or star cluster by
%
\be  \label{eqn: rstar}%$
M_*(t) = 4\pi\int_{0^+}^{r_*(t)} r^2 \rho(r,t) dr.
\ee  %$
%
Thus $r_*(t)$ is the radius at which the enclosed gas mass equals the
mass in stars; this makes sense if the distribution of young stars is more
centrally concentrated than the gas, which is true on large enough
scales. For a pure power law density distribution $\rho(r)\sim r^{-k_\rho}$,
%
\be %$
r_*(t)=R\left[{M_*(t)\over M_{cl}}\right]^{1/(3-k_\rho)},
\ee %$
%
where $R$ is a fiducial radius, and $M_{cl}$ is the mass of gas
inside that radius. We have in mind the thickness of the ribbon-like
features found in hydrodynamic simulations, e.g., those in paper I,
for $R$.

%We point out that $M_*(t)<<M_g(R,t)$ except possibly at very late
%times; it follows that $M_g(r_*(t),t)<< M_g(R,t)$.

\subsection{Solution}
We will concentrate on a solution from the time when a star first
forms.  We define scaled variables
%
\phil{ SHOULD $r_0 \rightarrow R$}
\be %$
x\equiv {r\over r_0}
\ee %$
and
%
\be %$
y\equiv{t\over\tDyn},
\ee %$
%
where $r_0$ is a fiducial radius, and $\tDyn=R/u_r(R)$. 
%

We define
%
\be  %$
p\equiv{r\over v_T}{\partial v_T\over \partial r},
\ p'\equiv{r\over u_r}{\partial u_r\over \partial r},\ {\rm and}
\ k_\rho\equiv-{r\over \rho}{\partial \rho\over \partial r},
\ee  %$
%
anticipating that each will vary only slowly with radius. In other
words, we look for solutions in which the density, infall velocity,
and turbulent velocity are all approximately power laws in both $r$
and $t$,
%
\be %$
\rho(r,t)=\rho_0x^{-k_\rho} y^\alpha,
\ee %$
%. 
%
\be %$
u_r(r,t)=u_0 x^{p'} y^\beta,
\ee %$
%
and
%
\be %$
v_T(r,t)=v_0 x^p y^\gamma.
\ee %$
%
where we have defined $\rho_0\equiv\rho(r_0,\tDyn)$, with similar
expressions for $u_0$ and $v_0$.

For an infall solution, we expect that the magnitude of the infall
velocity increases with time; mass accumulates in the central object,
which then exerts a gravitational force on the surrounding material
that increases with time. Our energy closure then ensures that the
turbulent velocity will increase (at a fixed radius) with increasing
time, so that
%
\be %$
\beta>0,\quad\gamma>0.
\ee %$
%

We define $\tilde\rho(t)\equiv\rho_0 y^\alpha$, so that
$\rho(r,t)=\tilde\rho(t) x^{-k_\rho}$.

\subsection{Solution at Small Radius, $r<r_*(t)$} \label{sec: inside}
The continuity equation becomes
% 
\be  \label{eqn: continuity below rstar}%$
{\partial\tilde\rho\over\partial t} x^{-k_\rho}+{u_0v_0\over
  r_0}x^{p'-k_\rho-1}y^{\alpha+\beta}
\left[2+p'-k_\rho\right]=0.
\ee  %$
%
As $r\to0$, the second term grows more rapidly than the first term,
since $p'<1$; in fact we will show that $p'=-1/2$. Thus at small radii
($r<<r_*)$ \phil{ satisfying this equation demands that both terms go to zero independently.  For second term, this implies}
%
\be \label{eqn: rho u}%$
k_\rho=2+p'.
\ee %$
%
\phil{Similarly, for the first term, we find \sout{
It follows that at small radii }}
% 
\be  \label{eqn: density attractor}%$
{\partial \rho(r,t)\over \partial t}=0.
\ee  %$
%
This is a very striking result: self-similar gravitational collapse
(in which the density and infall velocity follow power laws in radius)
evolves to an attractor, in which the density does not vary with
time. Note that the argument has made no reference to the type of
pressure support, i.e., thermal, turbulent, or magnetic.

\subsubsection{Using the energy closure to relate $u_r$ to $v_T$}\label{sec:energy closure}
Using the power law ansatz, Equation (\ref{eqn: 1D energy}) becomes
% 
\bea  \label{eqn: coupling}%$
\gamma&{v_0\over\tDyn}&y^{\gamma-1}x^p+{u_0v_0\over r_0} 
y^{\beta+\gamma}x^{p+p'-1}\nonumber\\
\qquad&\times&
\left[p+1+ \left({v_T(r,t)\over u_r(r,t)}\right)\eta\right]=0.
\eea  %$
%

\phil{\sout{As with the continuity equation, the term involving spatial
derivatives grows more rapidly as $r\to0$ than the term involving the
time derivative. Thus for $r<<r_*$ we find} We can then show}
% 
\be \label{eqn: eliminate} %$
v_T(r,t)=-u_r(r,t){1+p\over \eta}-{\gamma\over\etaeff}{r\over t}.
\ee  %$
%

\phil{At small $r$ and large $t$, the result (\ref{eqn: eliminate}) implies that $v_T \sim u_r$ and \sout{
The result (\ref{eqn: eliminate})}} is the analog of the result in
\citet{2012ApJ...750L..31R} that $v_T\sim a^{\tilde\beta-1}$, where $a$ is
the scale factor in their cosmological simulations and $\tilde\beta$ is
given by their Equation (9). We note that this result implies
%
\be \label{eqn: energy exponent}%$
p=p',\quad \beta=\gamma, \quad {\rm for}\ r<r_*.
\ee %$
%

\subsubsection{The Momentum Equation at Small Radii}
We start by noting that
%
\be %$
M(r,t)=M_*(t) + {4\pi\over 3-k_\rho}r_0^3\rho_0x^{3-k_\rho},
\ee %$
%
where the second term on the right hand side does not depend on
time\phil{, a result due to equation (\ref{eqn: density attractor})}.

The momentum equation is
%
\bea  \label{eqn: puzzle}%$
&\Bigg(&{u_0\over \tDyn}y^{\beta-1}x^{p'}+{4\pi G\rho_0r_0\over
  3-k_\rho}
x^{1-k_\rho}\Bigg)+\nonumber\\
&&\quad\Bigg[
{u_0^2\over r_0}p'y^{2\beta}x^{2p'-1}+{v_0^2\over
  r_0}(2p-k_\rho)y^{2\gamma}x^{2p-1}+\nonumber\\
&&\quad{GM_*(t)\over r_0^2}x^{-2}
\Bigg]
=0.
\eea  %$
%
\phil{We have written equation (\ref{eqn: puzzle}) anticipating the
  result.  Namely, we need to balance the gravitational term from the
  gas (second terms in parenthesis) and the central star (third term
  in square brackets) against the acceleration term (first term in
  parenthesis), the inertial term (first term in square brackets), or
  the pressure term (second term in square brackets). As $r\to 0$, the
  stellar gravity term dominates and we choose to balance it against
  both the inertial and pressure terms given the result of
  \S~\ref{sec:energy closure}. This gives
\sout{The second term (in square brackets) increases much more
rapidly than the first term (in curved brackets) as $r\to 0$, so we must have}}
%
\be \label{eqn: u_r small r}%$
u_r=-\Gamma
\sqrt{GM_*(t)\over r},
\ee %$
%
where we have used Equation (\ref{eqn: eliminate}), and defined
%
\be \label{eqn: Gamma}%$
\Gamma\equiv\left[(k_\rho-2p)\left({1+p\over\etaeff}\right)^2-p'\right]^{-1/2}.
\ee %$
%

Combining Equation (\ref{eqn: u_r small r})
with Equation (\ref{eqn: rho u}) shows that 
%
\be \label{eqn: vanishing r predictions} %$
p'=p=-{1\over2},\quad k_\rho={3/2},\quad (r<r_*).
\ee %$
%
Equation (\ref{eqn: u_r small r}) also shows that $M_*(t)\sim
t^{2\beta}$.

The first term in Equation (\ref{eqn: puzzle}) must separately equal to
zero, leading to
%
\be %$
{\partial u_r\over \partial t}=-{4\pi G \rho(r_0)r_0\over 3-k_\rho}x^{1-k_\rho}.
\ee %$
%
This shows that the infall velocity increases linearly with time
($\beta=1$) after
the central object forms. Integrating,
%
\be \label{eqn: second velocity}%$
u_r(r,t)=u_r(r,0)-{4\pi G \rho(r_0)r_0\over 3-k_\rho}x^{-1/2}t.
\ee %$
%

The two expression (\ref{eqn: u_r small r}) and (\ref{eqn: second
  velocity}) must agree at late times, so that:
%
\be %$
M_*(t)=\Gamma^{-2}
\left({4\pi\rho(r_0)\over 3-k_\rho}\right)^2Gr_0^3 t^2.
\ee %$
%
Anticipating a result from the next section, that $k_\rho\approx1.6$,
and ignoring for simplicity the difference between this value and
$k_\rho=1.5$, this can be written in terms of the clump mass
%
\be \label{eqn: mass vs time}%$
M_*(t)=\left({\pi\over 2\Gamma}\right)^2
M_{cl}\left({t-t_*\over\tff}\right)^2,
\ee %$
%
where we now explicitly display $t_*$, the time at which the central density diverges, or that
at which the central star forms.

It follows that
%
\be \label{eqn: r_star}%$
r_*(t)=\left[{\pi\over2\Gamma}\right]^{4/3}
\left({t-t_*\over\tff}\right)^{4/3}R.
\ee %$
%

It is instructive to compare the acceleration terms in the momentum
equation, normalizing to $g\approx G M_*(t)/r^2$.  The acceleration
due to the pressure gradient is
%
\be  \label{eqn: small r pressure}%$
{1\over g}\left({1\over\rho}{\partial P\over \partial r}\right)
\approx(2p-k_\rho)\left({v_T\over u_r}\right)^2\Gamma^2
\approx-{5\over 5+4\etaeff^2},
\ee  %$
%
while the net acceleration of the gas is approximately
%
\be  %$
{1\over g}{du_r\over dt}\approx p'\Gamma^2\approx-{4\etaeff^2\over 5+4\etaeff^2}.
\ee  %$
%
\phil{\sout{
For $\etaeff\approx1$, the acceleration due to the pressure gradient
and the inward acceleration are roughly equal, with each equal to
about half the acceleration due to gravity. In other words, for
$r<r_*$, gravity dominates the dynamics, but only by a factor of two;
the infall velocity is substantially smaller than the free fall
velocity.
}
For $\etaeff = 2/3$, the ratio between the acceleration due to the pressure gradient and gravity is 
$(\partial P/\partial r)/\rho g \approx 0.74$ (eq.[\ref{eqn: small r pressure}]), i.e., they nearly balance each other.  
Hence, the net acceleration of the gas is only roughly one quarter that of gravity. In other words, for
$r<r_*$, gravity dominates the dynamics, but it is nearly balanced by turbulent pressure and 
the infall velocity is substantially smaller than the free fall
velocity.}
%See p43 in the Feb 2014 notes

In the next section we show that for $r>r_*$, both $p'$ and $p$ are greater
than zero; the turbulent pressure decelerates the infall, and the
turbulent velocity tracks the infall velocity, as prescribed by
Equation (\ref{eqn: simple energy}). As the flow crosses $r_*$, the
acceleration due to gravity, which we will show scales as $r^{-1/2}$
for $r>r_*$, transitions to scaling as $r^{-2}$ for $r<r_*$.  The
infall velocity responds to this increased gravity rapidly, i.e. in a
moderately small fraction of the local dynamical time; the turbulent
velocity follows, but with a substantial lag (of order the local
dynamical time), eventually reaching $p=p'$, as enforced by our energy
closure, Equation(\ref{eqn: simple energy}). The acceleration due to the
pressure gradient tracks that due to gravity at all radii. However for
$r>r_*$ the pressure term dominates the gravity term, while for
$r<r_*$, the gravity term is roughly twice the pressure term, as we
have just seen.

%% Outside the sphere of influence, but still at small radii
%% $r_*<<r<<R$, combining eqns. (\ref{eqn: small r momentum}) and (\ref{eqn: mass}) 
%% % 
%% \be  \label{eqn: p from momentum}%$
%% 2p=2-k_\rho.
%% \ee  %$
%% %
%% Combining this with Equation (\ref{eqn: attractor}), $k_\rho=2$ and
%% $p=0$. The infall velocity approaches a constant, non-zero
%% value. Thus, while the run of density looks like that of a
%% a singular isothermal sphere in hydrostatic equilibrium, the
%% solution is not hydrostatic. This solution only appears if there are
%% many decades in radius between the outer scale of the turbulence and
%% the minimum radius, which will be set by physics not included in our
%% model, e.g., an angular momentum barrier in collapse which is not
%%perfectly spherically symmetric.

%%For large $\etaeff$, the
%%infall velocity approaches the free-fall value, while $v_T\to 0$. For
%%more physically plausible values $\etaeff\lesssim 1$, the infall
%%velocity is smaller than the free-fall velocity, and $v_T\approx -u_r/2\etaeff$.

\subsubsection{Summary at $r<<R$}

Gathering the results, we have shown that the turbulent velocity
tracks the infall velocity, that $k_\rho=2+p'$ (from the continuity
equation), and $p'=-1/2$ for $r<r_*(t)$; it follows that $k_\rho=-3/2$ for
$r<r_*(t)$.

The mass accretion rate for $r<r_*(t)$ is roughly independent
of radius,
% 
\bea  \label{eqn: flat mdot}%$
{dM_g(r,t)\over dt}&=&4\pi r^2 \rho(r,t)u_r(r,t)\nonumber\\
&\approx& 4\pi \R^2\rho_f(\R)u_r(\R,t),
\eea  %$
%
where we used the fact that $2-k_\rho+p'=0$ for $r<r_*$.

Although the density $\rho(r,t)=\rho(r)$, i.e., it is independent of
time, the accretion rate does vary with time since the infall velocity
$u_r(t,r)$ does. In other words, the accretion rate does not vary with
radius (for $r<r_*$) at a fixed time, but it does increase linearly
with time at a fixed radius.

\subsection{Solution for $r_*<r\lesssim R$}
It is useful to define times scales for the density, radial velocity,
and turbulent velocity to change by a factor of order unity:
%
\be  %$
\tau_\rho\equiv\left({1\over \rho}{\partial \rho\over \partial t}\right)^{-1},
\ee  %$
%
%
\be  \label{eqn: tau u_r}%$
\tau_{u_r}\equiv\left({1\over u_r}{\partial u_r\over \partial t}\right)^{-1},
\ee  %$
%
and
%
\be  %$
\tau_{v_T}\equiv\left({1\over v_T}{\partial v_T\over \partial t}\right)^{-1}.
\ee  %$
%
Well away from $r=r_*$ and $r=R$, we expect that these should be of
order the local dynamical time
%
\be  %$
\tdyn\equiv{r\over v_r}.
\ee  %$
%

The momentum equation can be written as
%
\be  %$
{\tdyn\over\tau_{u_r}}+p'+\left({v_T\over u_r}\right)^2(2p-k_\rho)
+\left({v_T\over u_r}\right)^2{GM/r\over v_T^2}=0.
\ee  %$
%
Since the turbulent velocity tracks the infall velocity, $v_T/u_r$
varies slowly with $r$; for $r_*<< r<<R$, we expect $\tdyn/\tau_{u_r}$
to vary slowly with $r$ as well. Then
% 
% 
\be  \label{eqn: large r velocity}%$
v_T(r,t)=\Gamma'\sqrt{GM(r,t)\over r},
\ee  %$
%
where 
%
\be  \label{eqn: Gprime}%$
\Gamma'\equiv \left|{v_T\over u_r}\right|\Gamma
=\left[
k_\rho-2p-
\left(p'+{\tdyn\over \tvr}\right)
\left({u_r\over v_T}\right)^2
\right]^{-1/2}
\ee  %$
%
is a slowly varying function of $r$ for $r>r_*$. Note that
$\Gamma^{\prime 2}\approx \alpha_{\rm vir}$, the virial parameter.

Because $\Gamma'$ is a slowly varying function of $r$, it follows from
Equation (\ref{eqn: large r velocity}) that 
% 
\be  \label{eqn: large r momentum} %$
2p=2-k_\rho, \quad r>r_*.
\ee  %$
%

The continuity equation can be written as
%
\be  %$
{\tdyn\over \tau_\rho}+2+p'-k_\rho=0.
\ee  %$
%
Combining this with Equation (\ref{eqn: large r momentum}), we
find
%
\be \label{eqn: large r p prime} %$
p'=\left|{\tdyn\over\trho}\right| - 2p, \quad r>r_*.
\ee  %$
%

In the appendix we argue that if the outer boundary condition on the
density is $\rho(R,t)=\bar\rho$, we can approximate
%
\be  \label{eqn: trho}%$
\left|{\tdyn\over\trho}\right|\approx
\left|{\ln r/r_*\over\ln R/r_*}-1\right|.
\ee  %$
%

In the Appendix we show that 
%
\be  %$
p(r) = {1\over2}k_\rho -{1\over2}\left({u_r\over  v_T}\right)^2
\left[{\tdyn\over \tvr}+p'+\Gamma^{-2}(0)\right],
\ee  %$
and estimate $p\approx0.1-0.2$.

From Equation (\ref{eqn: large r momentum}), we find
%
\be  %$
k_\rho \approx1.6-1.8,
\ee  %$
%
again depending on the initial conditions and on $\etaeff$.

Unlike the case at $r<r_*$, here the dynamics are controlled by the
pressure gradient, and not by gravity. The ratio
%
\be \label{eqn: pressure gradient ratio} %$
{1\over g}\left({1\over\rho}{\partial P\over \partial r}\right)=-(k_\rho-2p)\Gamma'^2\approx-3
\ee  %$
\phil{
\be \label{eqn: pressure gradient ratio} %$
{1\over g}\left({1\over\rho}{\partial P\over \partial r}\right)=-(k_\rho-2p)\Gamma'^2\approx-3 \,{\rm to}\, -4
\ee  %$
}
for $v_T\approx1.6 \sqrt{GM(R)/R}$, the fiducial value used in the
numerical simulations below\phil{, and $p \approx 0 - 0.2$, motivated by the results of the Appendix}.
%
%% while the ratio of the inertial term to that of gravity is
%% %
%% \be  %$
%% {d u_r\over dt}\Big/ g=
%% \left(
%% {u_r \over v_T}
%% \right)^2
%% \left(
%% {\tdyn\over \tau_{u_r}}+p'
%% \right)
%% \Gamma'^2.
%% \ee  %$
%

The force due to the steep pressure gradient (or more precisely the
density gradient, since $k_\rho >> |2p|$) is larger than the force of
gravity.  This outward-directed force causes the infall velocity to
decrease inward, for $r_*<r<<R$. This is quantified by combining
Equation (\ref{eqn: large r p prime}) with Equation (\ref{eqn: trho}). At
$r_*\lesssim r<<R$, we expect $p'\approx1-2p>0$, so that $|u_r|$
decreases with decreasing radius.

This result, that $|u_r|$ decreases inward for $r>r_*$, is in contrast to
self-similar solutions with an assumed fixed turbulent velocity. In
such theories, the infall velocity is either zero outside the radius
of the expanding collapse wave \citep{1977ApJ...214..488S,2003ApJ...585..850M} or has
a magnitude that {\em increases} inward even at large radii
\citep{2004ApJ...615..813F}.

In the next section, we use direct numerical integration of the 1D
equations to show that $0.1\lesssim p\lesssim0.25$, depending on the
outer boundary conditions and on $\etaeff$, similar to the result of
3D simulations in Paper I.

This value of the turbulent exponent is much smaller than the exponent
in non-self-gravitating supersonic turbulence, $p=1/2$, and smaller
than the Kolmogorov value $p=1/3$ for subsonic turbulence. Recall that
the Kolmogorov value follows from conservation of energy in the
cascade from large scales to small scales. In the case of supersonic
collapse, the exponent is smaller than the value one would derive
under the assumption that the kinetic energy in the cascade is
conserved. Thus, the kinetic energy on small scales is larger than
that produced by an energy conserving cascade.

The reason is simple: a fraction of the ordered inflow kinetic energy,
and a fraction of the potential energy released in the collapse, is
being converted into turbulent kinetic energy. In our simple theory,
this fraction is set by the closure relation, Equation (\ref{eqn: simple
  energy}), which captures the behavior seen in the numerical
simulations of \citet{2012ApJ...750L..31R}. This extra energy input
boosts the turbulent velocity relative to that in an energy conserving
cascade.

Inside the sphere of influence $r_*$ the conversion of gravitational
potential energy into turbulent energy is even more dramatic, as we
saw above ($p=-1/2$), so that the turbulent velocity {\em increases} with
decreasing $r$.

Simulations of non-self-gravitating turbulence consistently show
$p\approx0.5$, as do observations of GMCs
\citep{1981MNRAS.194..809L,2004ApJ...615L..45H} and low mass star
formation regions \citep{1983ApJ...270..105M, 1992A&A...257..715F} in
the Milky Way. We suggest that measurements of the scaling of
turbulent velocity can be used as a sign post for collapse, a point we
return to below.

As in the case $r<r_*$, we have found that the dynamics converges to a
run of density that varies only on the large scale dynamical time. In
this case (for $r>r_*$) $k_\rho$ is slightly larger than the value
$k_\rho=1.5$ found for $r<r_*$.

Unlike the small $r$ case, however, we also find that the run of
velocity (either $u_r(r)$ or $v_T(r)$) evolves rapidly to a solution
that varies only on the dynamical time evaluated at the outer scale
$R$, i.e., for $r_*<r<R$ the velocity increases only very slowly with
time. This variation, on the global dynamical time, is simply a result of
the change in the mass contained inside $r=R$. For $r_*<<r<<R$ the
turbulent velocity is well described by a power law (with
$p\approx0.2$) while the infall velocity shows substantial deviations
from a simple power law, as given by eqns. (\ref{eqn: large r p
  prime}) and (\ref{eqn: trho}).

It follows that the mass accretion rate for $r>r_*$ varies fairly rapidly
with radius,
%
\be  \label{eqn: larger r mdot}%$
{dM_g(r,t)\over dt}=4\pi r^2\rho(R)u_r(R,t)\left({r\over R}\right)^{2-k_\rho+p'},
\ee  %$
%
where $2-k_\rho+p'\approx 0.2-0.3$. We have allowed
for the possibility that the infall velocity at $R$ increases with
time, as the growing mass of the clump will tend to accelerate the
infall, but this increase is on the (large scale) dynamical time, and
hence proceeds rather leisurely; essentially, the mass accretion rate
is independent of time for $r>r_*$, but not for $r<r_*$!

%% In fact we can give a simple expression for the mass accretion rate
%% onto the central object, in terms of the mass of the clump (recall
%% that the mass is dominated by the gas mass). We begin by estimating
%% %
%% \be  %$
%% \rho(r)=\rho(R)\left({r\over R}\right)^{-\bar k_\rho},
%% \ee  %$
%% %
%% where now $\bar k_\rho$ corresponds to a fit over the entire range of $r$;
%% since the values for $k_\rho$ at $r<r_*$ and at $r>r_*$ are not too
%% dissimilar, we simply take $\bar k_\rho=1.5$. Then
%% %
%% \be  %$
%% {dM_*\over dt}=4\pi r_{in}\rho(r_{in})u_r(r_{in},t)
%% \ee  %$
%% %
%% becomes
%% %
%% \be  %$
%% {dM_*\over dt}=4\pi \rho(R) \Gamma \sqrt{GM_{cl}\over R}
%% \left({M_*\over M_{cl}}\right)^{1/2} ,
%% \ee  %$
%% %
%% where
%% %
%% \be  %$
%% M_{cl}\approx {4\pi\over 3}R^3\rho(R){3\over 3-\bar k_\rho}.
%% \ee  %$
%% %
%% We remind the reader that $\Gamma=\sqrt{8\etaeff^2/(5 +
%%   4\etaeff^2)}$ is of order unity. Then
%% %
%% \be  \label{eqn: mass vs time}%$
%% M_*(t)=\pi^2{(3-\bar k_\rho)^2\over 2}\Gamma^2 
%% M_{cl}\left({t-t_*\over  \tff(R)}\right)^2.
%% %\nonumber\\&\approx& M_{cl}\left({t-t_*\over  \tff(R)}\right)^2.
%% \ee  %$
%
In a more realistic model, including the effects of thermal gas
pressure, magnetic pressure, and magnetic tension, as well as three
dimensional effects, this mass accretion rate will be reduced. We
discuss these effects briefly in the next subsection.

\subsection{The Effects of Thermal and Magnetic Pressure, Rotation, and
  Magnetic Tension} There are several factors that will reduce the
mass accretion rate in a more realistic model. First, we do not expect
that gas will fall onto the star forming region from every direction,
as we have assumed so far. In the three dimensional simulations in
Paper I, we find that the infall covers a solid angle
$\Omega_{in}<4\pi$ steradians, which we account for by a factor
$f_\Omega\equiv\Omega_{in}/(4\pi)$. In our MHD simulations, the
fraction of the sky (as seen from the central object) over which
accretion occurs is yet smaller, as the magnetic field line tension
inhibits accretion across field lines, described by a factor
$f_{\Omega_B}\le1$. Finally, the pressure associated with the magnetic
field energy density provides an outward pressure gradient much like
the turbulent pressure.

Rotational support will tend to slow the rate of accretion. However,
observations suggest that the rotational velocity $v_c$ in star
forming cores is much smaller than the turbulent velocity, e.g.,
\citet{2011A&A...525A.151B,2014ApJ...785...42P}, who find
$v_c/v_T\lesssim0.3$ at $r\approx 0.05\pc$. This
appears to hold down to scales of order $1000{\rm AU}$ or
$5\times10^{-3}\pc$, below which there is evidence for rotating disks,
e.g., \citet{2000ApJ...537..283V}.

We can include the effects of thermal pressure, and, in an approximate
way, some of the effects of magnetic pressure and tension. All three
tend to reduce the mass accretion rate, although for the massive star
formation regions we consider here, with large virial velocities and
high Mach numbers, the effects of thermal pressure are small.

The effects of magnetic pressure are potentially very significant. We
expect that the strong turbulence generated in the converging flows we
consider will tend to magnify the ambient magnetic field, possibly
reaching equipartition, i.e., a magnetic pressure equal to $\rho
v_T^2$, or a substantial fraction thereof. This effect can be
incorporated in a very rough way by assuming a pressure of the form
%
\be \label{eqn: magnetic pressure} %$
P(r)=\rho(r)\left[c_s^2+\phi_Bv_T^2\right],
\ee %$
%
where $\phi_B=1+B^2/(8\pi\rho v_T^2)$. This expression does not
entirely capture the effect of the magnetic pressure, since it implies
that the magnetic field decays when the turbulent velocity does,
contrary to what we expect. This might be important if the magnetic
pressure becomes large enough to reduce the infall velocity of the
flow; reducing $u_r$ will reduce $v_T$, and hence in (\ref{eqn:
  magnetic pressure}), the pressure associated with $B$. In fact, once
the turbulence generates the magnetic field, the decay time of the
magnetic field need not be related to the decay time of the
turbulence.

We denote the fraction by which the magnetic pressure decreases the
accretion rate by $f_{B^2}$; we show below, using our simple model,
that for a plasma $\beta\approx1$, $f_{B^2}\approx 1/2$.

Accounting for all these effects, the expression for the stellar mass becomes
%
\bea  %$
M_*(t)&=&f_\Omega f_{\Omega_B} f_{B^2}\left({\pi\over
  4\Gamma}\right)^2
M_{\rm  cl}\left({t-t_*\over\tff}\right)^2\\
&=&\phi M_{\rm  cl}\left({t-t_*\over\tff}\right)^2.
\eea  %$
%

\section{Numerical results}\label{sec: numerics}
We numerically integrate the full set of equations
(\ref{eqn: continuity}), (\ref{eqn: momentum}), and (\ref{eqn: 1D
  energy}), supplemented by (\ref{eqn: mass}) \phil{using a simple first order upwind code}.

Our initial conditions are that the density is constant,
$\rho(r,t=0)=\bar\rho$, while the radial and turbulent velocities follow
a power law,
% 
\be  %$
u_r(r,t=0)=-|u_r(R)|\left({r\over R}\right)^{p\prime}
\ee  %$
%
and
% 
\be  %$
v_T(r,t=0)=v_T(R)\left({r\over R}\right)^p,
\ee  %$
%
with fiducial values $\bar\rho=3.0\times10^{-22}\g\s$,
\phil{\sout{$|u_r(R)|=6.0\kms$, $v_T(R)=-3.0\kms$} $v_T(R) = 6.0\kms$, $u_r(R)=-3.0\kms$} and $p=p'=1$.  The default value
for the eddy decay time-scale is $\etaeff=2/3$. We have explored
variations around these fiducial values, and report on some of the
results below. The boundary radius $R=3-10\pc$ for most of our
runs. The initial mass is $M(R,0)=4\pi R^3\bar\rho/3\approx
1.8\times10^4(R/10\pc)^3M_\odot$, and the corresponding Keplerian velocity is
%
\be  %$
v_k(R,0)=\sqrt{GM(R,0)/R}\approx2.8\kms.
\ee  %$
%

Our outer boundary conditions specify inflow, with
\phil{\sout{$u_r(R,t)=-|u_r(R)|$,}} $v_T(R,t)=v_T(R)$\phil{\sout{,}} and
$\rho(R,t)=\bar\rho$. \phil{For $u_r(R,t)$, we set it to be a power law extrapolation of $u_r(R-dr, t)$, where $dr$ is the grid scale.} The
inner boundary conditions are $u_r(r_{in},t)=u_r(r_{in}+dr,t)$,
$v_T(r_{in},t)=v_T(r_{in}+dr,t)$, and \phil{\sout{$\rho(r_{in},t)=\rho(r_{in}+dr,t$)} $(r_{in})^2\rho(r_{in}, t) = (r_{in}+dr)^2\rho(r_{in}+dr)$}, i.e., we insist that the mass flowing into the innermost shell
$\dot M(r,t)$,
which is deposited onto the central star, be the same as the mass flow
rate into the surrounding shell.

We have integrated the equations both with and without including the
effects of a finite sound speed. The results are virtually
identical to those when the sound speed is neglected.

Figure \ref{fig_rho_v_vs_r_080} shows the run of density and infall
velocity for fifteen snapshots between $t=0\tau_{\rm DYN}$ and
\phil{\sout{$t=0.247\tau_{\rm DYN}$}$t\approx 0.253\tau_{\rm DYN}$}, where $\tau_{\rm DYN}\equiv |R/u_r(R)|$. The
first four snapshots show a core in the density distribution,
$\rho(r)=const.$, with the constant increasing rapidly from
$3\times10^{-22}\g\cm^{-3}$ in the initial snapshot to
\phil{\sout{$10^{-17}\g\cm^{-3}$}$10^{-16}\g\cm^{-3}$} in the fourth snapshot. By the seventh snapshot, the
density distribution has evolved into a (slightly broken) power
law. The break in the power law occurs at $r=r_*(t)$; for $r<r_*(t)$,
$\rho(r)\sim r^{1.5}$, while for $r>r_*(t)$, $\rho(r)\sim r^{k_\rho}$,
with $k_\rho\approx1.7$.

The infall velocity $-u_r(r,t)$ also evolves into a broken power
law. For $r<r_*(t)$, the exponent $p'=-0.5$, while for $r>r_*(t)$,
$p'\approx 0.2$\phil{\sout{; at very large radius $u_r$ flattens to meet our
outer boundary value, so that $p'$
decreases, reaching negative values at $r\approx R$.}, before flattening at $r \approx R$.}

\begin{figure}
\epsscale{1.2}
\plotone{fig1}
\caption{\label{fig_rho_v_vs_r_080}The run of density $\rho(r)$ (upper
  panel), infall velocity $-v_r(r)$ (lower panel, solid line), and
  turbulent velocity $v_T(r)$ (lower panel, dashed line) shown for 15
  different times, ranging from the initial conditions to shortly
  after the formation of a star, which occurs at
  \phil{\sout{$t_*/\tdyn\approx0.246$} $t_*/\tdyn\approx0.253$}. The density increases from its initial value
  $\bar\rho=3\times10^{-22}\g\cm^{-3}$ to a maximum
  \phil{\sout{$\rho\approx10^{-14}\g\cm^{-3}$}$\rho\approx10^{-13}\g\cm^{-3}$} at the inner edge of the integration
  region. Starting from the time just before the star forms, the
  density distribution is very nearly independent of time. }
\end{figure}

Figure \ref{fig_pressure_gravity_vs_r} shows the ratio
% 
\be  %$
{(\nabla\rho v_T^2)/\rho \over {GM(r,t)/r^2}}
\ee  %$
%
of the pressure gradient term to the gravity term in the
momentum equation. We distinguish between times previous to the
formation of the central object (shown by dotted lines) and times
after the central object forms (shown by solid lines). \phil{Note that while we have plotted a series of snapshots, these curves do not correspond to the curves in Figure \ref{fig_rho_v_vs_r_080}, i.e., these curves do not have the same times as the previous figure}

At early times and small radii, the density is nearly constant, so
$(1/\rho)dP/dr\sim dv_T^2/dr>0$, and the pressure gradient provides an
inward force, i.e,., it acts in the same sense as gravity. Thus the
pressure gradient force and the force due to gravity both act to move
gas toward the origin. However, at large radii, the dotted curves drop
down, eventually becoming negative at large radii, i.e.,
the pressure gradient term is directed outward, opposing the force due
to gravity. Near the outer boundary, this outward pressure gradient
term is larger in magnitude than the acceleration of gravity.

At the first time plotted, the radius at which the pressure gradient
changes sign is located at \phil{\sout{$r\approx 1.5\pc$}$r\approx 0.8\pc$}; well inside this radius,
the density is uniform, $\rho(r)=const.$ At later times, the
transition to inward acceleration occurs at \phil{\sout{$r\approx 0.1\pc$,
$1\times10^{-2}\pc$, and $7\times10^{-4}\pc$}$r\approx 2\times 10^{-3}\pc$}. Once again, at a fixed
time, but well inside this transition radius, the density is uniform.
%
\begin{figure}
\plotone{fig2}
\caption{\label{fig_pressure_gravity_vs_r}The ratio of the pressure
  gradient term to the gravity term in the momentum equation,
  $(\partial \rho v_T^2/\partial r)/(\rho GM(r,t)/r^2)$. The dotted
  lines correspond to times before the central star forms, while the
  solid lines correspond to times after the star forms. At large
  radius the outward acceleration due to the pressure gradient exceeds
  the inward acceleration due to gravity (the dotted and solid lines
  are below the dot-dashed horizontal line at -1); this is why the
  infall velocity decreases inward for $r>r_*(t)$, as seen in figure
  \ref{fig_rho_v_vs_r_080}. Inside $r_*(t)$, the acceleration due to
  the mass of the star exerts a gravitational force on the surrounding
  gas, but the star provides no outward pressure gradient; the small
  amount of gas between $r=r_{in}$ and $r_*$ cannot provide a
  sufficiently strong pressure gradient, and so the magnitude of the
  infall velocity increases inward. However, even after the star
  forms, the acceleration provided by the pressure gradient is always
  a substantial fraction of the total gravitational acceleration---the
  solid lines are only slightly above the dot-dashed line for
  $r<r_*$. The horizontal dashed line is the prediction of
  Equation (\ref{eqn: small r pressure}), at $-5/(5+4\etaeff^2)$, where in
  this case $\etaeff=2/3$. For $r>r_*$, the ratio is roughly \phil{\sout{$-3$}$-4$},
  consistent with the prediction of Equation (\ref{eqn: pressure gradient
    ratio}).}
\end{figure}

After the central object forms the situation is very different. To
begin with, the pressure gradient is directed outward at all radii
(aside from some numerical artifacts near the outer boundary). However
while the pressure gradient force is directed outward, it is smaller
in magnitude than the force due to gravity at radii $r<r_*(t)$. Thus
the infall velocity of a (Lagrangian) mass shell increases inward. The
ratio of the two forces given by Equation (\ref{eqn: small r
  pressure}) is constant, and is depicted in Figure 2 by the
horizontal dashed line. The numerical result at late times (after the
star forms) tracks this prediction for $r<r_*$. As a result, the
infall velocity $u_r\sim r^{-1/2}$, but it is smaller than the
free-fall velocity by a factor $\Gamma/\sqrt{2}\approx 1/2$. This is
different than the prediction of the turbulent core model, in which
the pressure term becomes negligible as $r\to0$, so that the infall
velocity approaches the free-fall value.

In contrast, for $r>r_*(t)$ the pressure gradient term is larger in
magnitude than the gravity term, so that the infall velocity of a mass
shell {\em decreases} inward.  Note that the inertial term
$u_r\partial u_r/\partial r$ is much smaller than either the pressure
gradient term or the gravity term, or their difference, so that the
infall velocity decreases in an Eulerian sense as well, as seen in
Figure \ref{fig_rho_v_vs_r_080}.

Figure (\ref{fig_Gamma_vs_r_150}) shows $\Gamma$ and $\Gamma'$ \phil{for $t/t_{\rm DYN}\approx 0.27$}. Note
that $\Gamma$\phil{and $\Gamma'$} is nearly constant for $r<r_*$, while $\Gamma'$ is
roughly constant for $r>r_*$. The prediction of Equation (\ref{eqn:
  Gamma}) using the values for the exponents at
$r=0$ is shown in the figure as the horizontal dotted line; it agrees
\phil{\sout{fairly}} well with the numerical results.

\begin{figure}
\epsscale{1.2}
\plotone{fig3}
\caption{\label{fig_Gamma_vs_r_150}The run of $\Gamma(r)$ (solid
  line), and $\Gamma'\equiv (v_T/|u_r|)\Gamma(r)$ (dashed line). The prediction of
  eqns. (\ref{eqn: Gamma}) at $r=0$,
  $\Gamma=\sqrt{8\etaeff^2/(5+4\etaeff^2)}$, for $r<r_*$ is shown as
  the horizontal dotted line.}
\end{figure}

Figure (\ref{fig_p_vs_r_101}) shows the exponent $p$ in the
size-linewidth relation, $v_T(r)\sim r^p$. The exponent is positive
for $r>r_*$, but much smaller than the value $p=1/2$ seen in
simulations of non-self-gravitating turbulence. For $r<r_*$ the value
of $p$ quickly plunges to $-0.5$\phil{, adopting the value of a Keplerian profile. However, the normalization is below that of the Keplerian free-fall velocity as we discussed above}. The predictions of the analytic
theory, eqns. (\ref{eqn: p small r}) for $r<r_*$ and (\ref{eqn:
  p for r big}) for $r>r_*$ are also
shown, and agree well with the numerical results except in the
immediate vicinity of $r_*$.

\begin{figure}
\epsscale{1.3}
\plotone{fig4}
\caption{\label{fig_p_vs_r_101}The run of $p(r)$, the exponent in the
  size-linewidth relation (solid line),
  and the prediction of Equation (\ref{eqn: p small r}) for $r<r_*$ (dashed line),
  and of Equation (\ref{eqn: p for r big}) for $r_*<r$ (dotted line). }
\end{figure} %April 29 long

Figure (\ref{fig_p_prime_vs_r_101}) shows the result for $p'$; the
prediction $p'=-1/2$ for $r<r_*$ and the prediction of
Equation (\ref{eqn: large r p prime}) are shown as the dashed and dotted
lines, respectively.  
\phil{Here we note again that the domination of the mass of the
  central star inside $r<r_*$ drives the radial velocity to assume a
  Keplerian profile, though its normalization does not reach free-fall values.}

Figure (\ref{fig_k_rho_vs_r_101}) plots the density exponent $k_\rho$ 
\phil{(solid line) and the prediction of Equation (\ref{eqn: vanishing r predictions}) for $r<r_*$ (dashed line),
  and of Equation (\ref{eqn: large r momentum}) for $r_*<r$ (dotted line). The prediction of $k_{\rho} = 1.5$ is recovered nicely in the numerical results. }

\begin{figure}
\epsscale{1.3}
\plotone{fig5}
\caption{\label{fig_p_prime_vs_r_101}The run of $p'(r)$, where $u_r(r)=u_r(R)(r/R)^{p\prime}$
  (solid line),
  and the prediction $p'=-1/2$ for $r<r_*$ (dashed line),
  and of Equation (\ref{eqn: large r p prime}) for $r_*<r$ (dotted line). }
\end{figure} % April 29 long

\begin{figure}
\epsscale{1.3}
\plotone{fig6}
\caption{\label{fig_k_rho_vs_r_101}The run of $k_\rho(r)$, where $\rho(r)=\rho(R)(r/R)^{-k_\rho}$
  (solid line),
  and the prediction of Equation (\ref{eqn: vanishing r predictions}) for $r<r_*$ (dashed line),
  and of Equation (\ref{eqn: large r momentum}) for $r_*<r$ (dotted line). }
\end{figure} % April 29 long

Figure (\ref{fig_mass_vs_time_April_12}) shows the accreted mass
versus $t-t_*$, with $t_*$ defined to be the time when the accreted mass
exceeds some minimum mass $M_{\rm min}$; in the figure, \phil{\sout{$M_{\rm
  min}=10^{-3}M_\odot$}$M_{\rm min}=10^{-4}M_\odot$}. Varying $M_{\rm min}$ alters the appearance of
the plot at early times but has no effect at later times. The mass
increases in proportion to $(t-t_*)^2$, and the numerical result is in \phil{\sout{ excellent}good}
agreement with the prediction of Equation (\ref{eqn: mass vs time}).

\begin{figure}
\epsscale{1.3}
\plotone{fig7}
\caption{\label{fig_mass_vs_time_April_12} The accreted mass as a
  function of time since the star forms, $t-t_*$, where
  $t_*/\tdyn\approx0.25$. The solid line is the numerical result, the
  dashed line is the prediction of Equation (\ref{eqn: mass vs
    time}). }
\end{figure}

%% Figure (\ref{fig_r_star_7200}) shows the radius $r_*(t)$ of the sphere
%% of influence of the central object as a function of $t-t_*$, together
%% with the prediction of Equation (\ref{eqn: r_star}).

%% \begin{figure}
%% \epsscale{1.3}
%% \plotone{fig8}
%% \caption{\label{fig_r_star_7200} The radius of the sphere of influence
%%   $r_*(t)$. The solid line is the numerical result, the dashed
%% line is the prediction of Equation (\ref{eqn: r_star}).  }
%% \end{figure}

%Figure (\ref{fig_magnetic}) shows the run of density and velocity when
We have calculated the run of density and velocity when
the effects of a strong magnetic field, corresponding to a plasma
$\beta=1$ relative to $\rho v_T^2$, are
included in the manner described above; in the notation of
Equation (\ref{eqn: magnetic pressure}), $\phi_B=2$. The extra pressure provided by
the magnetic field slows the infall, reducing the rate of accretion
compared to a similar run with no magnetic pressure support by a
factor of 1.9.

%% \begin{figure}
%% \epsscale{1.3}
%% \plotone{fig8}
%% \caption{\label{fig_magnetic} The run of density (top panel), and of
%%   turbulent and infall velocity (bottom panel) in a run including the
%%   effects of a magnetic
%%   field of strength $B^2(r,t)=8\pi \rho(r,t) v_T^2(r,t)$,
%%   i.e. assuming that the field is in equipartition with the turbulence
%%   at all times and at all radii.  }
%% \end{figure}

\section{DISCUSSION}
\label{sec:discussion}

\subsection{The Connection Between Self-Gravity and the
Size-Linewidth Relation}

Larson's first law, the size-linewidth relation, applies to GMCs as
well as to low mass star forming clumps. The power law index in GMC's
($v_T(r)\sim r^p$) is almost
universally observed to be $p\sim 0.4-0.5$, e.g.,
\citet{1981MNRAS.194..809L,1987ApJ...319..730S,2004ApJ...615L..45H}.
 The scaling relation extends to lower mass and smaller objects, e.g.,
\citet{1983ApJ...270..105M}, who finds $p\approx0.5$.

In contrast, massive star forming regions show $p\approx 0.2$, and
have much higher levels of turbulence at a given radius than is seen
in low mass star formation regions, e.g., \citet{1992ApJ...384..523F,
  1995ApJ...446..665C,1997ApJ...476..730P}. The turbulent velocities
in these clumps are also much larger than the extrapolation of the GMC
size-linewidth relation predicts. For example
\citet{1997ApJ...476..730P} examine $r\sim0.1-1\pc$ clumps with masses
$\sim 10^3-10^4M_\odot$ selected on the basis of H$_2$O maser
emission. They find turbulent velocities $v_T\approx 2-10\kms$ (FWHM),
nearly a factor of ten above Larson's relation for $r\approx
0.3\pc$---see their Figure 5.

We have shown quantitatively how the
conversion of gravitational potential energy into turbulent motion via
adiabatic heating, coupled with the back pressure produced by that
turbulence, can explain $p\approx0.2$, for $r>r_*$, a result
illustrated in Figure \ref{fig_p_vs_r_101}.

We can extract another prediction: in our model of converging flows,
and in the simulations of \phil{Paper I}, the density
increases monotonically as $r$ decreases. However, the turbulent
velocity does not vary monotonically with $r$: at large radii
($r>r_*$) the turbulent velocity decreases as $r$ decreases, but
inside the critical radius, the turbulent velocity {\em increases}
with decreasing $r$. It follows that a plot of turbulent velocity
versus density for a single region will show that at low density the
velocity decreases with increasing density, while at high density the
velocity will increase with increasing density. In regions with
different total mass, this breakpoint will occur at different
radii. Even in regions with the same total mass, the break occurs at
different radii, depending on the age of the system, since the
critical radius depends on \phil{\sout{age} the mass of the central star or star cluster}.

\citet{1997ApJ...476..730P} examine massive clusters and find that the
linewidth increases with density. The gas densities they examine are
very high, of order $n=10^5$ to $10^7\cm^{-3}$. Their highest
luminosity systems exceed $10^6L_\odot$, which require stellar masses
of order $3\times10^3M_\odot$. They give a virial estimate of the
total mass, finding values of order $5000M_\odot$. This suggests that
they may be resolving $r_*$ in their most luminous objects.

Collapse models such as that of
\cite{1997ApJ...476..750M,2003ApJ...585..850M,2004ApJ...615..813F}
predict $u_r(r)$,
%%
%%$u_r(r)\sim r^{p'}$ with $u(r)$ equal
%%to the free-fall velocity and $p'=-1/2$ at small radii (where the
%%pressure is negligible), and $p'\le0$ at all radii. \
%%
but they do not predict the run of $v_T(r)$ at any radius---they take
it as an input. They assume that $v_T(r)\sim r^p$, with $p>0$. Thus
the linewidth will always decrease with increasing density, a
corollary of the model that is inconsistent with the observations of
\citet{1997ApJ...476..730P}, a point those authors noted.

We interpret the finding that $p\approx0.5$ in GMCs and low mass star
formation regions to mean that GMCs, as a whole, are not undergoing
rapid gravitational collapse. This is not to say that the
virial parameter $\alpha_{\rm vir}$ is larger than one; in the most massive
GMCs, $\alpha_{\rm vir}\sim1$. Conversely, we interpret the
observation that $p\approx0.2-0.3$ in compact massive clumps to mean
that they are undergoing rapid gravitational collapse \phil{and are the centers of large scale converging flows}. We note that in
the turbulent collapse model describe here the
infall velocity should be substantially below the free-fall velocity,
a point we return to below.

Papers reporting on simulations of turbulently stirred gas, in which
the turbulence is maintained at a level corresponding to $\alpha_{\rm
  vir}\approx1$, often report the slope of the velocity power
spectrum $P({\bf k})\equiv|v({\bf k})|^2\sim k^{-n_{3D}}$. This slope
is related to the index $p$ by the relation $p=(n_{3D}-3)/2$, or if
the one-dimensional power spectrum is used, $n_{1D}=n_{3D}-2$,
$p=(n_{1D}-1)/2$.  The implied values of $p=0.45$ to $0.5$ in
hydrodynamic turbulence and $p=0.45-0.6$ in MHD simulations
\citep{2007ApJ...661..972P,2006ApJ...638L..25K,2012ApJ...750...13C};
these simulations do not include gravity. However, even simulations
which do include gravity find large values of $n_{1D}$, e.g.,
\citet{2012ApJ...750...13C}, and by implication, large values of $p$.

Paper I (which included the effects of gravity) looked in the vicinity
of local density maxima, which are undergoing collapse, and showed by
direct measurement that $p\approx 0.2-0.3$, substantially smaller than
the global average value in the same turbulent box. That paper also
showed that the velocity power spectrum calculated in the immediate
vicinity of a collapsing region has more power on small scales than
the power spectrum calculated in exactly the same manner around random
points in the same simulation box. The power spectrum, when calculated
over an entire simulation box, is too crude a tool to detect the
presence of collapsing flows.

\subsection{Slow, Radially Extended Infall}\label{sec: slow infall}
The large turbulent pressure support which results from adiabatic
heating tends to reduce the magnitude of the infall velocity, so that
$|u_r(r)|\approx v_T(r)\approx-\Gamma\sqrt{GM_*/r}$ at small radii,
and $|u_r(r)|\lesssim v_T(r)\approx \sqrt{GM(r)/r}\sim r^{+0.2}$ at
large radii. The 3D simulations in Paper I show find a similar result
(see Figure 9 in that paper). The slow increase of the turbulent
velocity with increasing $r$ (for $r>r_*$) is consistent with a
modified Larson's first law, $p\approx0.2$.

\begin{figure}
\epsscale{1.3}
\plotone{fig9}
\caption{\label{fig_rho_v_data} The density, infall, and turbulent
  velocity as a function of radius. The lines give the run of density
  (top panel), and of turbulent and infall velocity (dashed and solid
  lines respectively, in the bottom panel) in the run depicted in
  Figure 1, but at the time when the central mass $M_*=89M_\odot$. The
  points in the upper panel are estimates of the density in the
  galactic massive star forming region G10.62-0.38. In order of
  increasing radius, the data is from \citet{2011A&A...525A.151B},
  \citet{2008ApJ...684.1273K}, and \citet{1986ApJ...304..501H}. The
  points in the lower panel are measurements of $v_T$ (filled symbols)
  and $-u_r$ (open symbols). In order of increasing radius, the
  measured value of $-u_r$ is represented by an open circle
  \citep{2011A&A...530A..53K}, a downward pointing open triangle
  \citep{2008ApJ...684.1273K}, and an upward pointing open triangle
  \citep{1986ApJ...304..501H} The filled square, downward filled
  triangle, and upward filled triangle are from
  \citet{2011A&A...525A.151B}, \citet{2008ApJ...684.1273K}, and
  \citet{1986ApJ...304..501H}; the observational papers do not provide
  an estimate of the errors in the turbulent velocity. The apparent
  dip in $-u_r$ around $r=0.2\pc$ is qualitatively
  consistent with the adiabatically heated turbulent collapse model.}
\end{figure}

These findings contrasts with the expanding collapse wave and related
models, which predict that $v_T$, and hence the pressure, becomes
negligible at small radii, and that the infall velocity reaches the
free-fall value $u_r=-\sqrt{2GM_*/r}$. In fact this class of
models predicts that $|u_r|$ decreases with increasing r at all radii,
or, for expanding collapse wave models, falls to zero at a finite
radius and remains there.  The turbulent velocity is
assumed to be an increasing function of $r$ at all radii.

The different predictions of the infall models should allow us to
distinguish between them. For example, \citet{2008ApJ...684.1273K} see
clear evidence of infall in line profiles of HCO$^+$ at $r=0.3\pc$ in
G10.6-0.4, with $u_r=-1.0\pm0.1\kms$; see Figure
\ref{fig_rho_v_data}. The infall extends inward to $0.03\pc$ as shown
by \citet{2011A&A...530A..53K}, who see inverse P Cygni profiles in CO
lines and infer $u_r=-2.5\kms$. \citet{1986ApJ...304..501H} provide
somewhat less robust evidence for infall at $r=0.5\pc$, finding
$v_T\approx6\kms$ and $-u_r\approx 5\kms$.

\citet{2011A&A...525A.151B} estimate $v_T=6\kms$ at
$r=0.04(D/6\kpc)\pc$, and a rotational velocity $v_c=2.1\kms$. Hence
in this core, the turbulent pressure dominates both the rotational
energy and the kinetic energy of the infall, by a factor of 9 or
more. Using the masses estimated below, it follows that the infall
velocity is well below the free-fall velocity, by a factor of about
two or three. This is evidence that something is slowing the infall. The fact
that the turbulent velocity is substantially larger than the infall
velocity suggests that turbulent pressure is responsible.

The bolometric luminosity of G10.6-0.4 is $L\approx10^6L_\odot$, and
the stellar mass is therefor at least $M_*\approx 200M_\odot$
\citep{2011ApJ...729..100L}; if the stellar mass function in the
proto-cluster is like the global initial mass function, the stellar
mass could be a thousand solar masses. The gas mass on the scale of
$1.5''\approx 0.04\pc$ is $M_g\sim150M_\odot$
\citep{2009ApJ...703.1308K}. The rough equality of stellar and gas
masses suggests that the observations are probing sizes around
$r_*$. This may explain the increase in the velocity inferred from
$-1\kms$ at $r\approx0.3\pc$ to $-2.5\kms$ at $0.03\pc$, while at the
same time $|u_r|<\sqrt{2GM/r}$. The lines in Figure
\ref{fig_rho_v_data} come from the same model show in Figure
\ref{fig_rho_v_vs_r_080}, but at a later time, specifically, when
$M_*\approx100M_\odot$.

It is worth remarking that the innermost two data points in Figure
\ref{fig_rho_v_data} illustrate a similar phenomenon to that found in
\citet{1997ApJ...476..730P}; the density increases inward, while the
turbulent velocity is flat or possibly slightly increasing inward.

We note that this observation of infall over a decade in radius, from
$r\approx 0.03\pc$ to $0.3\pc$, and possibly to $0.5\pc$, is
consistent with the solutions found in this paper. This observation
does not appear to be consistent with the predictions of the turbulent
core model of \citet{2003ApJ...585..850M}, where the fiducial
hydrostatic core radius $R_{\rm core}=0.05\pc$, and the collapse
proceeds from the inside out, i.e., the infall should be seen only for
$r\lesssim R_{\rm core}$.

In addition, the apparent increase of $|u_r|$ as $r$ increases beyond
$0.3\pc$ suggested by the observations of \citet{1986ApJ...304..501H}
is inconsistent with collapse models.

\subsection{The Star Formation Efficiency Per Free Fall Time}

We have shown that the star formation rate in self-gravitating
turbulent flows increases linearly in time in the absence of any feedback
physics. This rapid growth in the star formation rate cannot continue
for a dynamical time in real GMCs, since the fraction of gas converted
into stars is seen to be less than $10-20\%$ in the most extreme
cases, and lower in many GMCs.

We suggest that the low observed star formation efficiency is due to
two types of effects. The first is the effect of early
(pre-supernovae) feedback from the newly formed stars, e.g.,
\citet{2010ApJ...709..191M}.

The second is the effect of initial conditions;
GMCs with large virial parameters will not form stars efficiently,
since they will disperse in less than a GMC free-fall time. This
dispersal cuts off the flow of gas to any collapsing regions in the
cloud, starving the nascent stars. Evidence for this is presented in
paper I, in particular in Figure 14.

Recent papers on the star formation rate in 3D simulations have
generally presented the result in terms of a constant value of $\eff$
\citep[e.g.][]{2011ApJ...730...40P,2012MNRAS.419.3115B,2012ApJ...754...71K}. 
The constant $\eff$ that is reported in these papers is found by
fitting a straight line to $M_*(t)$ at late times, typically a
substantial fraction of a free-fall time after the first stars have
formed. However, it is clear from examining the figures in these
papers that $M_*(t)$ is increasing more rapidly than $\sim t$.  
\phil{Workers that have examined the early time behavior of the star formation rate in their simulations such as 
\citet{Myer+14} have found that $\eff$ is not constant with time.}

\subsection{Long-lived Density Structures} 
The theory of adiabatically heated turbulent collapse predicts the
existence of long-lived (compared to the local free-fall time) density
structures. These structures are static in the sense that their shape
does not change; however, they are dynamic in that fluid flows through
them from large (multi-parsec) to small (milli-parsec) scales. For a
significant fraction, of order one quarter, of the large scale
dynamical time, these structures contain no newly formed stars. Seen
in projection, or via extinction, they will have power-law surface
density probability distribution functions. In the early phase, before
the time we have denoted by $t_*$, the gas mass contained in the
structure above a given column will increase with time; after $t_*$ it
will be constant.

Direct observational measurements of $k_\rho$ are difficult, and
exhibit a large range $1.0\lesssim k_\rho\lesssim 2.0$, e.g.,
\citep{2000ApJ...537..283V,2014ApJ...785...42P}, making it difficult
to distinguish between different theoretical models.

Regions in which $\rho(r)\sim r^{-k_\rho}$ will naturally produce
extinction maps with powerlaw tails to the probability distribution of
extinction \citep{2011ApJ...727L..20K}.  There are only a handful of
observational measurements of the surface density PDF, which should
scale as $\Sigma\sim r^{-2/(k_\rho -1)}$. Power law tails have been
seen in nearby star forming regions, e.g.,
\citet{1999A&A...345..965C,2009ApJ...703...52L,2009A&A...508L..35K}. \citet{2010A&A...512A..67L}
attribute this the effects of gravity. \citet{2011ApJ...727L..20K}
perform numerical simulations of self-gravitating turbulence and find
power-law tails in both the volume density and surface density
distributions. They attribute these tails to collapsing regions. They
also note that if the power law tail of the density $PDF
\sim\rho^{-m}$ is produced by a run of density around density maxima
$\rho(r)\sim r^{-k_\rho}$, then $m=3/k_\rho$.

We agree with this interpretation, with the refinement that the outer
layers of the star forming regions, before self-gravity becomes
dominant, will have a run of density with $k_\rho=1.6-1.8$, while the
gravity-dominated inner region has $k_\rho=1.5$. It is the turbulent
pressure, not the self-gravity, that produces $k_\rho>1.5$, and hence
density PDF tails that have $m<2$.

\subsection{Driving by Protostellar Jets}
Low mass star forming regions show $p=0.5$, similar to that in the ISM
as a whole. Many such regions show prominent proto-stellar jets
\citet{2001ARA&A..39..403R}, which are often suggested to be drivers
of turbulence, e.g., \citet{2005ApJ...632..941Q}. In his analytic
models of turbulence driven by proto-stellar jets,
\citet{2007ApJ...659.1394M} finds $\sigma(r)\sim r^{1/2}$, i.e.,
$p=0.5$ for $r<l$, where $l\approx0.38\pc$ is the characteristic size
of a region subject to protostellar jets with fiducial
properties. Simulations of jet driven turbulence
\citep{2009ApJ...695.1376C} find {\em steeper} spectra,
$p\approx0.87$. The theory does not predict $p\approx0.2$ over an extended
range of radii, whereas the observations show $p\lesssim0.2$ for
$0.04\pc\lesssim r\lesssim 1.6\pc$, e.g., \citep{1995ApJ...446..665C}
or \citep{1997ApJ...476..730P}.

\citet{2004ApJ...615L..45H} note that $p\approx0.5$ over a large range
of scales and for low mass star forming regions, and argue that either
jets produce a feature in $\sigma(r)$ around $r=l$, or the driving due
to outflows is not important even at small scales. In fact, many low
mass star forming regions, e.g., those studied by
\citet{1983ApJ...270..105M}, fall on the original Larson relation,
i.e, with the appropriate magnitude, indicating that neither jets nor
conversion of infall energy to turbulence enhance the turbulence
substantially over that expected from a simple extension of the
behavior on larger scales.

The flat run of turbulent velocity seen in massive star forming
regions, with no indication of a peak on scales of order a parsec,
combined with the considerations above, suggest that the enhanced
turbulence is not due to driving by proto-stellar jets.

\subsection{Local Versus Global Collapse}
The notion that there are coherent collapsing substructures in
turbulent flows in which gravity is important has been suggested in
the recent literature, e.g., \citet{2011MNRAS.411...65B}, and in
particular \citet{2012ApJ...750...13C}. The latter authors stress the
existence of self-similar collapsing spheres at high density. Like
\citet{2011ApJ...727L..20K}, they measured density PDFs $P(\rho)\sim
\rho^m$ with $m\approx1.6-1.8$, and inferred $k_\rho=3/m\approx
1.67-1.88$. It is these spheres, and their dynamics, that we model in
this paper.

\citet{2011MNRAS.411...65B} argue that
  ``...the observational data are consistent with molecular clouds in
  a state of hierarchical and chaotic gravitational collapse,
  i.e. developing local centres of collapse throughout the whole cloud
  while the cloud itself is collapsing, and making equilibrium
  unnecessary at all stages prior to the formation of actual stars.''

We agree that local centers of collapse occur throughout the
cloud. There is direct evidence for local collapse, on parsec scales,
in the form of inverse P Cygni profiles seen toward known star forming
clumps
\citep{2001ApJ...562..770D,2008ApJ...684.1273K,2011A&A...530A..53K,2011A&A...525A.151B,2013A&A...554A..83V}.

However, we disagree that GMCs are collapsing globally. We are not aware of any
detections of P Cygni profiles toward GMCs on large scales. We
would also point out that the Larson's Law index $p$ measured on large
scales is $\sim 0.4-0.5$
\citep{1981MNRAS.194..809L,2011ApJ...740..120R}, substantially larger
than that measured on small scales, typically $p\approx0.1-0.2$
\citep{1995ApJ...446..665C,1997ApJ...476..730P}. If small values of
$p$ are taken as evidence for collapse around star forming regions,
the large values of $p$ seen on large scales must be taken as evidence
against large scale collapse in GMCs. Given enough time, and in the
absence of star formation, GMCs might collapse globally, but in the
Milky Way this does not appear to be happening.

\citet{2007ApJ...657..870V} ran simulations in which two streams of
gas were directed at each other; the subsequent collision generates
turbulence. This is followed, after a dynamical time or two, by a
global gravitational collapse. This collapse drives yet higher
turbulent velocity, with the result that the global kinetic energy is
thereafter similar in magnitude to the global gravitational
energy. This is analogous to our result, but we are examining a local
collapse, whereas their's is a global result in a simulation
containing a fairly large number of locally collapsing regions
embedded in a globally collapsing cloud.

\citet{2008MNRAS.390..769V} noted that in their simulations, regions
around density maxima showed convergent flows, a result recovered in
the simulations in paper I. They note in their section 6.1 that their
diagnostic tools were not sufficient to decide whether the inflow was
the cause or the consequence of gravitational contraction. We have
shown that it is both---initially convergence leads to an
over-density, and eventually to star formation, whose gravity then
speeds the infall (increases the convergence) on small scales (inside
$r_*$).

\section{CONCLUSIONS}
\label{sec:concl}
We have shown that turbulent self-gravitating flows with virial
parameters near unity have the following properties:

The run of density asymptotes to 
\be
\rho(r,t)=
\begin{dcases}
\rho(r_0)\left({r\over r_0}\right)^{-3/2}, & r<r_*\\
\rho(R,t)\left({r\over R}\right)^{-k_\rho}, \ k_\rho\approx1.6-1.8 & r>r_*.
\end{dcases}
\ee
Since $k_\rho$ is fairly close to $1.5$ at all radii, taking $r_0=R$
is a fair approximation. The prediction that
$\rho(r,t)\to\rho(r)$ for $r<r_*$ is in agreement with the MHD and
HD simulations in Paper I.

This run of density leads to a density PDF of the form $P(\rho)\sim
\rho^{1.6-1.8}$ at high densities, and the related power law for the
surface density PDF. This is in accord both with high resolution three
dimensional MHD and hydro simulations (e.g. \citealt{2011ApJ...727L..20K}; \citealt{2012ApJ...750...13C}; \phil{Paper I}), and with
observational maps of visual extinction \phil{\citep[e.g.][]{2009ApJ...703...52L,2009A&A...508L..35K}}.

The infall velocity 
%
\be
u_r(r,t)=
\begin{dcases}
-\Gamma\sqrt{GM_*(t)\over r}, \sim r^{-1/2} & r<r_*\\
-\Gamma\sqrt{GM(r,t)\over r} \sim r^{0.2} & r>r_*,
\end{dcases}
\ee
%
where \phil{\sout{$\Gamma\approx 0.5$} $\Gamma \approx 0.7$} at small radii, and $\Gamma\approx 1.0$ at
large radii.

In contrast to most earlier models, the star forming clump is never in
hydrostatic equlibrium. A clear test to decide between the two types
of model is to measure the infall velocity over a large range of
radii; both the expanding collapse wave model and the turbulent core
model predict zero infall velocity at moderate radii (outside the
collapse wave). We noted in \S \ref{sec: slow infall} that there are some hints
that the infall extends to $r\sim0.3\pc$. Models that select solutions with finite infall
velocities at large distance, e.g., \citet{2004ApJ...615..813F}
predict lower values for the infall velocity at large radii, for a given mass
accretion rate and run of density than the model presented here. In
particular, they predict that $|u_r|$ decreases monotonically as $r$
increases, whereas the adiabatically heated models predict that
$|u_r|$ {\em increases} as $r$ increases, for $r>r_*$.

The turbulent velocity 
%
\be
v_T(r,t)=
\begin{dcases}
{1\over 2\etaeff}\Gamma\sqrt{GM_*(t)\over r}, \sim r^{-1/2} & r<r_*\\
{1.2\over \etaeff}\Gamma\sqrt{GM(r,t)\over r} \sim r^{0.2} & r>r_*,
\end{dcases}
\ee
%
The result $p<0$ for $r<r_*$ explains the observation of
\citet{1997ApJ...476..730P} that the linewidth ($v_T$) increases with
increasing density, a result that they noted was puzzling in the
context of collapse models, since the latter assumed $p>0$ at all radii.

The result that $p\approx0.2$ for $r>r_*$ is the first explanation, to
our knowledge, for the deviation from Larson's law in massive
clumps. In fact, both the larger magnitude of $v_T$ at a given small
distance, and the different slope $p$ are in good agreement with
observations. Both are the direct result of the conversion of infall
kinetic energy and gravitational potential energy to turbulent motions
as the accreting gas falls inward.

The stellar mass increases quadratically with time
%
\be  %$
M_*(t)=\phi M_{\rm cl}\left({t-t_*\over \tff}\right)^2.
\ee  %$
%

The mass accretion rate 
%
\be
\dot M(r,t)=
\begin{dcases}
4\pi R^2\rho(R)u_r(r,t), \sim t\,r^{0} & r<r_*\\
4\pi R^2\rho(R)u_r(r,t) \sim t^0\,r^{0.2} & r>r_*.
\end{dcases}
\ee
%
It may be possible, with ALMA, to measure both the density (using
different tracers) and the infall velocity over a range of radii in
massive star forming clumps.

%% We have argued that the star formation rate in gravitationally bound
%% objects is controlled by gravity. Examples of such objects include the
%% most massive giant molecular clouds, clumps, and cores. Under the
%% assumption that turbulence provides pressure support, the expanding
%% collapse wave model predicts that the star formation rate is an
%% increasing function of time and that a substantial fraction of the
%% cloud mass will turn into stars on a free fall time. From our result of 
%% super-linear growth of star formation rate, most of this star
%% formation will in fact happen over the last one or two tenths of a
%% free fall time. We find that the radial density profile steepens over 
%% time around density peaks at which stars form and a power-law 
%% tail at the high density end builds up on top of lognormal density PDF. 
%% Combined with the fact that the exclusion of self-gravity in our simulation 
%% produces lognormal density PDF shows self-gravity to dynamically alter the 
%% density PDF. Furthermore, both velocity profile and the velocity power 
%% spctrum show the turbulent velocity to be enhanced on small 
%% scales near local density peaks suggesting self-gravity to be affecting 
%% not only the density structure but the properties of turbulence as well. 
%% Unlike the high SFE seen in our simulations of virialized cloud, super-virial 
%% clouds show much slower growth in stellar mass; we have argued that this is 
%% because there is not enough time for gas to collapse before turbulence 
%% completely alters the density structure. Our results show that self-gravity 
%% acting on scales larger than 0.1 pc or more have a direct consequence 
%% on clouds' density structure, turbulence, and the rate of star formation. 
%% From our simulation results as well as the agreement with recent observations, 
%% we conclude that star formation rate is a dynamic, time-varying property, not 
%% a constant as previously thought.

%\acknowledgment

We would like to thank P. Klaassen for helpful conversations, and for
sharing unpublished results. NM is supported by the Canada Research
Chair program and by NSERC of Canada. \phil{PC acknowledges support from the NASA ATP
program through NASA grant NNX13AH43G, and NSF grant AST-1255469}.  This work was supported in part
by the National Science Foundation under Grant No. PHYS-1066293 and
the hospitality of the Aspen Center for Physics. Some of the
computations were performed on the gpc supercomputer at the SciNet HPC
Consortium \citep{2010JPhCS.256a2026L}. SciNet is funded by: the
Canada Foundation for Innovation under the auspices of Compute Canada;
the Government of Ontario; Ontario Research Fund - Research
Excellence; and the University of Toronto. The authors acknowledge the
Texas Advanced Computing Center (TACC) at The University of Texas at
Austin for providing HPC resources that have contributed to the
research results reported within this paper. URL:
\url{http://www.tacc.utexas.edu}

\appendix
\label{sec: appendix}
\section{Estimating the Index $p$ in Larson's Law}
We estimate the power law index $p$ in Larson's Law, for finite
values of $r$, first for $r<r_*$ and then for $r>r_*$.

\subsection{Estimating $p$ Inside the Sphere of Influence ($r<r_*$)}
We have shown in \S \ref{sec: inside} that $\lim_{r\to0}p=-1/2$. Here
we extend the derivation to finite values, but still for $r<r_*$.

We start from the momentum Equation (\ref{eqn: momentum}). Using
Equation (\ref{eqn: tau u_r}) and retaining the time derivative,
%
\be  %$
{\tdyn\over \tvr}+p'+\left({v_T\over
  u_r}\right)^2(2p-k_\rho)+{GM_*(t)/r\over u_r^2}=0.
\ee  %$
%
The last term must vary slowly with $r$, so we find $p'=-1/2$, even at
finite $r$. Since the continuity equation enforces $\partial
\rho/\partial t=0$, Equation (\ref{eqn: rho u}) then ensures that
$k_\rho=3/2$ at finite $r$. However, the time derivative of the infall
velocity is not zero, so $p(r)$ will depend on $r$.

Solving for $p(r)$,
%
\be  %$
p(r) = {1\over2}k_\rho -{1\over2}\left({u_r\over  v_T}\right)^2
\left[{\tdyn\over tvr}+p'+\Gamma^{-2}(0)\right],
\ee  %$
%
where $\Gamma(0)$ is given by Equation (\ref{eqn: Gamma}). Some algebra
yields
%
\be \label{eqn: p small r almost} %$
p(r)=p(0)-{1\over 2}\left({u_r\over v_T}\right)^2{\tdyn\over \tvr}.
\ee  %$
%

To evaluate this, we need an expression for $\tdyn/\tvr$ at small but finite $r$. We start
from Equation (\ref{eqn: u_r small r}), and estimate
% 
\be  \label{eqn: deriv}%$
{\partial u_r\over\partial t}\approx{1\over 2}u_r {1\over
  M_*(t)}{dM_*(t)\over dt}.
\ee  %$
%
Recall that
% 
\be  \label{eqn: mdot appendix} %$
{dM_*(t)\over dt}=-\lim_{r\to0}4\pi r^2\rho(r,t)u_r(r,t),
\ee  %$
%
while from Equation (\ref{eqn: mass}) we have
%% % 
%% \be  %$
%% M(r,t)={4\pi\over  3-k_\rho}\rho(r_*,t)r_*^3
%% \left({r\over r_*(t)}\right)^{3-k_\rho}
%% \ee  %$
%% %
%% for $r<r_*$. Then
% 
\be  \label{eqn: mass density} %$
4\pi r^2\rho(r,t)=
%%{3-k_\rho\over r} {4\pi\over
%%  3-k_\rho}\rho(r,t)r^3 =
{3-k_\rho\over r}M_g(r,t).
\ee  %$
%
Using eqns. (\ref{eqn: mdot appendix}) and (\ref{eqn: mass density}) in Equation (\ref{eqn: deriv}),
% 
\be  %$
{\partial u_r\over\partial t}\approx 
-\left({3-k_\rho\over 2}\right){u_r^2\over r}{M_g(r,t)\over M_*(t)}.
\ee  %$
%
Thus
% 
\be  \label{eqn: tdyn over tvr}%$
{\tdyn\over \tvr}\approx
 -\left({3-k_\rho\over 2}\right){M_g(r,t)\over M_*(t)}\approx
-{3\over 4}\left({r\over r_*(t)}\right)^{3/2}
\ee  %$
%
for $r<r_*$.

Using this in Equation (\ref{eqn: p small r almost}),
%
\be \label{eqn: p small r} %$
p(r)\approx -{1\over 2} +{3\over 2}\etaeff^2\left({r\over r_*}\right)^{3/2}
\ee  %$
%
This prediction is shown in Figure (\ref{fig_p_vs_r_101}) as the dashed line.

\subsection{Estimating $p$ Outside the Sphere of Influence ($r>r_*$)}
The calculation of $p$ for $r<r_*$ was greatly simplified by the fact
that $\partial\rho/\partial t=0$ inside the sphere of influence of the
stars. Outside the sphere of influence, however, we cannot neglect the
time variation of $\rho$, so the continuity equation becomes
%
\be  \label{eqn: final continuity equation}%$
{\tdyn\over\trho}+2+p'-k_\rho=0,
\ee  %$
%

Combining eqns. (\ref{eqn: final continuity equation}), (\ref{eqn:
  Gprime}), and (\ref{eqn: large r momentum}), we find
% 
\be  \label{eqn: p for r big}%$
p=\Bigg[
{1\over2}-{1\over (2\Gamma')^2}
-{1\over4}\left({u_r\over v_T}\right)^2
\left(
\left|{\tdyn\over\trho}\right|
+{\tdyn\over\tvt}
\right)
\Bigg ]
\left(1-{1\over2}{u_r^2\over v_T^2}\right)^{-1}.
\ee  %$
%

%% Given a value for $p$, the value of $k_\rho$ follows from
%% Equation (\ref{eqn: large r momentum}), allowing us to solve for $p'$ from
%% Equation (\ref{eqn: final continuity equation}).

To use Equation (\ref{eqn: p for r big}) to evaluate $p$, we need to
estimate the time scale ratios, which depend on the outer boundary
conditions. A simple estimate for the time scale over which the infall
velocity changes is
%
\be  %$
{\tdyn\over\tvr}\approx 1.
\ee  %$
%
The estimate for the density time scale has to take account of the
outer boundary condition, since in our favored scenario, stars form in
converging flows. At large scales in such flows
$\rho(R)\approx\bar\rho$, where $\bar\rho$ is independent of
time. Thus
% 
\be  %$
\lim_{r\to R}{\tdyn\over\trho}=0.
\ee  %$
%
At smaller radii, but still at radii large compared to $r_*$,
we expect 
$|\tdyn/\rho|\leq 1$.
% 
Since the solution is scale invariant, we estimate
%
\be  %$
\left|{\tdyn\over \trho}\right|\approx
\left|{\ln r/r_*\over\ln R/r_*} - 1 \right|.
\ee  %$
%

From Equation (\ref{eqn: eliminate}), and anticipating that $p<<1$, we have
%
\be  %$
\left({v_T\over u_r}\right)^2\approx \etaeff^2.
\ee  %$
%
For our fiducial $\etaeff=2/3$ and $\Gamma'=1.6$, the denominator in Equation (\ref{eqn:
  p for r big}) is roughly $3/4$; as a crude estimate we take
$|\tdyn/\tau_\rho|=1$, so 
%
\be  %$
p\approx4/3\left[{1\over 2}-{1\over10}-{1\over4}\cdot{1\over 2}\cdot2\right]\approx0.2
\ee  %$
%
as a very rough estimate.

%
\bibliography{Turbulent}

\end{document}

%%   Left over text:
%%
%%

It is useful to define
%
\be  \label{eqn: p}%$
p\equiv {r\over v_T}{\partial v_T\over \partial r},
\ p'\equiv {r\over u_r}{\partial u_r\over \partial r},\ {\rm and}\ 
k_\rho \equiv -{r\over \rho}{\partial\rho\over \partial r},
\ee  %$
%
and
% 
\be \label{eqn: tau} %$
\tvt\equiv v_T\Big/{\partial v_T\over\partial t},\,
\tvr\equiv u_r\Big/{\partial u_r\over\partial t},\ {\rm and}\ 
\trho\equiv\rho\Big/{\partial\rho\over\partial t}.
\ee  %$
%
We also define the dynamical time $\tdyn=r/u_r(r,t)$, and note
that this is a negative number for infall.

While $p=p(r,t)$ is a function of both $r$ and $t$, and
similarly for $p'$ and $k_\rho$,  we will show that $v_T$,
$u_r$ and $\rho$ follow broken power laws in $r$, so that $p$,
$p'$, and $k_\rho$ will, after an initial transient period, vary only
slowly with both $r$ and $t$.

With these approximations, the continuity equation, momentum equation,
and the energy closure equation become
%
\be \label{eqn: second continuity} %$
{\partial\rho\over\partial t}+
{\rho u_r\over r}\left[2+p'-k_\rho\right]=0,
\ee  %$
%
%
\be  \label{eqn: second momentum}%$
{\partial u_r\over\partial t}
+{u_r^2\over r}\left[
p'+\left({v_T\over u_r}\right)^2(2p-k_\rho)
+{GM(r,t)/r\over u_r^2}\right]
=0,
\ee  %$
and
%
\be  \label{eqn: second energy}%$
{\partial v_T\over \partial t}
+{v_Tu_r\over r}\left[p+1+\etaeff{v_T\over u_r}\right]
=0.
\ee  %$
%

%

Using eqns. (\ref{eqn: tau}), these equations can also be
written as 
%
\be  \label{eqn: final continuity equation}%$
{\tdyn\over\trho}+2+p'-k_\rho=0,
\ee  %$
%
%
\be  \label{eqn: final momentum equation}%$
{\tdyn\over\tvr}+p'+\left({v_T\over
  u_r}\right)^2(2p-k_\rho)+{GM/r\over u_r^2}=0,
\ee  %$
%
and
%
\be  %$
{\tdyn\over\tvt}+p+1+\etaeff{v_T\over u_r}=0.
\ee  %$
%

In the previous subsection we showed that $u_r(r,t)$ must vary with time,
although rather slowly compared to the local dynamical time at small
$r$. Equation (\ref{eqn: eliminate}) shows that $v_T(r,t)$ varies with
time as well. Thus the dynamics has a fixed (in time) attractor for
the density distribution at small $r$, but both the turbulent and
infall velocity increase with time at a fixed small radius, as long as $r<r_*$.

Equation (\ref{eqn: small r momentum})  motivates the definition 
%
\be  \label{eqn: Gamma}%$
\Gamma(r,t) = 
\left[
\left({v_T\over u_r}\right)^2(k_\rho-2p) - \left(p'+{\tdyn\over\tvr}\right)
\right]^{-1/2}.
\ee  %$
%

As noted above the three terms inside the square brackets in
(\ref{eqn: puzzle}) so arrange themselves so as to nearly, but not
quite, cancel off, with the residual having an $r$ dependence proportional to
$r^{3/2}$. Because of this near-cancellation, the time scale
$\tvr>>\tdyn$ at small r, but it is not zero.

### 2014-10-28 ms.tex
Added pdf_frame_132
added disks
Discussion of plots and PDFs

@@
 The simulated volume has a large number of refined regions.
-We show four of them in Figure \ref{fig:snapshot} as slices perpendicular to the angular momentum vector.
+We show three of them in Figure \ref{fig:snapshots} as slices perpendicular to the angular momentum vector.
 We determine the angular momentum vector by taking a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
+We also so a zoom out picture of the same region, which is the same for all three stars.
+While this is not unexpected as star formation is clustered, the large scale dynamics are determined by the larger region.
+
+A second family of two star particles were found in our simulation, show im Figure \ref{fig:snapshots 2}.
+Here again we show a zoom in and zoom out picture, comparing detail between the two particles around an axis of the angular momentum.
+Again the zoom in plots show two distinct star forming regions, while the zoom-out plot show that the larger region and hence large scale dynamics are the same.
 
 \begin{figure*}
-\plottwo{movie_disk_frame_0259.png}{movie_disk_frame_0257.png}
+\plottwo{movie_disk_frame_0259.png}{Disk_0132_0000.png}
 \plottwo{movie_disk_frame_0256.png}{movie_disk_frame_0272.png}
-\caption{Slices along the angular momentum axis of four zoomed in simulation.
+\caption{{\bf Change plots to show a zoom out for 132 and zoom in on the three regions that are the same} Slices along the angular momentum axis of four zoomed in simulation.
 To determine the angular momentum axis, we take a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
@@
 
-These figures suggest that the accretion can occur by streams or by spherical flows.
+\begin{figure*}
+\plottwo{movie_disk_frame_0259.png}{movie_disk_frame_0257.png}
+\plottwo{movie_disk_frame_0256.png}{movie_disk_frame_0272.png}
+\caption{{\bf Do same for other plots in 132} Slices along the angular momentum axis of four zoomed in simulation.
+To determine the angular momentum axis, we take a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
+We then take a slice in the plane normal to the angular momentum to plot the regions around these dense points.
+\label{fig:snapshots 2}}
+\end{figure*}
+
+The zoom in plots in Figures \ref{fig:snapshots} and \ref{fig:snapshots 2} suggest that the accretion can occur by streams or by spherical flows.
 However, the degree by which these modes of accretion dominate, is not clear.
@@
 As with the left panel, however, the turbulent velocity is significantly less volatile than the radial velocity and varies only slightly with $r$.
+These expectations are in line with the analytic results of Murray and Chang (2014), albeit with a somewhat larger scale that they suggest.
 
@@
 Here it is apparent that the regions that can be treated a point mass corresponds to regions in where the density follows a $r^{-3/2}$ power law and the radial and turbulent velocities follow a $r^{-1/2}$ power law.
-This is due to mass concentration at smaller radii as evident by the sudden drop in $M(r)$ at $r \approx 10^{-2}$ pc.
+This is due to mass concentration at smaller radii as evident by the sudden drop in $M(r)$ at $r \approx 10^{-2} - 1$ pc.
+The flattening of $M(r)$ on larger scales is interesting as this is due to the entire mass of the forming star cluster as opposed to the individual stars that MC14 originally envisioned.
+Hence the entire mass of the star cluster contributes to the large scale dynamics of turbulent collapse.
 
@@
 \caption{WRITE CAPTION: Mass, $MSun$, as a function of radial position from a local maximum density point.     \label{fig:angular_momentum_example}}
+\end{figure*}
+
+\begin{figure*}
+\plottwo{pdf_frame_0256.png}{pdf_frame_0132.png}
+\caption{Probability density function, $pdf$ as a function of density $rho/rho_0$.  \bf Verify range of slope calculation \label{fig:pdf_example}}
+\end{figure*}
+
+\begin{figure*}
+\plottwo{Disk_0132_0000.png}{Disk_0132_0004.png}
+\caption{Visualization of density profiles for two distinct particles in tracking1b. \label{fig:rho_example_132}}
 \end{figure*}

### 2014-11-11 ms.tex
improve density visualizations for 132
added more frames to folder, and to tex

@@
 \begin{figure*}
-\plottwo{movie_disk_frame_0259.png}{Disk_0132_0000.png}
-\plottwo{movie_disk_frame_0256.png}{movie_disk_frame_0272.png}
+\plottwo{movie_disk_0132_0001.png}{movie_disk_0132_0003.png}
+\plottwo{movie_disk_0132_0000.png}{Disk_0132_0000.png}
 \caption{{\bf Change plots to show a zoom out for 132 and zoom in on the three regions that are the same} Slices along the angular momentum axis of four zoomed in simulation.
@@
 \begin{figure*}
-\plottwo{movie_disk_frame_0259.png}{movie_disk_frame_0257.png}
-\plottwo{movie_disk_frame_0256.png}{movie_disk_frame_0272.png}
+\includegraphics[width=0.48\textwidth]{movie_disk_0132_0002.png}
+\includegraphics[width=0.48\textwidth]{movie_disk_0132_0004.png}
+\centering{\includegraphics[width=0.48\textwidth]{Disk_0132_0002.png}}
 \caption{{\bf Do same for other plots in 132} Slices along the angular momentum axis of four zoomed in simulation.
@@
 \label{fig:snapshots 2}}
+\end{figure*}
+
+\begin{figure*}
+\plottwo{Track_0132_000_rho.png}{pdf_frame_0132_0000.png}
+\plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
+\caption{Tracking plot for density and associated pdf graphs for two regions.
+The density graphs have slopes of -3/2, -2, and -5/2 shown for reference.
+The upper density frame on the left shows a consistent slope matching the -3/2 reference line very nicely.
+The pdf graphs show three volumes, one at 1pc, one at 2pc, and one at 5pc.
+\label{fig:Track_and_pdf_000}}
+\end{figure*}
+
+\begin{figure*}
+\plottwo{Track_0132_002_rho.png}{pdf_frame_0132_0002.png}
+\plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
+\caption{Tracking plot for density and associated pdf graphs for two regions.
+The density graphs have slopes of -3/2, -2, and -5/2 shown for reference.
+The upper density frame on the left shows a consistent slope matching the -3/2 reference line very nicely.
+The pdf graphs show three volumes, one at 1pc, one at 2pc, and one at 5pc.
+\label{fig:Track_and_pdf_002}}
 \end{figure*}

### 2014-11-13 ms.tex
Added captions to rho, vr, etc graphs for 132

@@
 \plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
-\caption{Tracking plot for density and associated pdf graphs for two regions.
-The density graphs have slopes of -3/2, -2, and -5/2 shown for reference.
-The upper density frame on the left shows a consistent slope matching the -3/2 reference line very nicely.
-The pdf graphs show three volumes, one at 1pc, one at 2pc, and one at 5pc.
-\label{fig:Track_and_pdf_000}}
+\caption{Tracking 132 000  These plots show various data for one of the particles formed in the simulation.
+The upper left graph shows the density profile, with reference lines indicating slopes of -3/2, -2, and -5/2.
+The slope between $0.02$pc and $0.4$pc approximately matches the -5/2 reference line.
+The plot on the upper right shows the pdf function related to volumes of 1pc, 2pc, and 5pc.
+The bottom left shows a plot of the radial velocity and turbulent velocity associated with the particle.
+The lower right graph shows the radial density profile of the particle.
+\label{fig:132 000 graphs}}
 \end{figure*}
@@
 \plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
-\caption{Tracking plot for density and associated pdf graphs for two regions.
-The density graphs have slopes of -3/2, -2, and -5/2 shown for reference.
-The upper density frame on the left shows a consistent slope matching the -3/2 reference line very nicely.
-The pdf graphs show three volumes, one at 1pc, one at 2pc, and one at 5pc.
-\label{fig:Track_and_pdf_002}}
+\caption{Tracking 132 002  These plots show the data for a different particle formed in the simulation at the same time as the previous plots.
+The upper left graph of the density in the region neatly matches the -3/2 slope.
+The pdf graph in the upper right again shows the plots for volumes related to $1$pc, $2$pc, and $5$pc.
+The lower left plot demonstrates the loose connection between the radial velocity and the turbulent velocity.
+The graph on the lower right again shows the density profile of this second particle.
+\label{fig:132_002_graphs}}
 \end{figure*}

### 2014-11-13 ms.tex
rearranged plots

@@
 \begin{figure*}
+\plottwo{Disk_0132_0000.png}{Disk_0132_0002.png}
+\caption{Cut of the two star forming regions.}
+\end{figure*}
+
+\begin{figure*}
+\plottwo{Disk_0132_0000.png}{Track_0132_000_rho.png}
+\plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
+\plottwo{Track_0132_000_mass.png}{pdf_frame_0132_0000.png}
+\caption{Tracking 132 000  These plots show various data for one of the particles formed in the simulation.
+The upper left graph shows the density profile, with reference lines indicating slopes of -3/2, -2, and -5/2.
+The slope between $0.02$pc and $0.4$pc approximately matches the -5/2 reference line.
+The plot on the upper right shows the pdf function related to volumes of 1pc, 2pc, and 5pc.
+The bottom left shows a plot of the radial velocity and turbulent velocity associated with the particle.
+The lower right graph shows the radial density profile of the particle.
+\label{fig:132 000 graphs}}
+\end{figure*}
+
+\begin{figure*}
+\plottwo{Disk_0132_0002.png}{Track_0132_002_rho.png}
+\plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
+\plottwo{Track_0132_002_mass.png}{pdf_frame_0132_0002.png}
+\caption{Tracking 132 002  These plots show the data for a different particle formed in the simulation at the same time as the previous plots.
+The upper left graph of the density in the region neatly matches the -3/2 slope.
+The pdf graph in the upper right again shows the plots for volumes related to $1$pc, $2$pc, and $5$pc.
+The lower left plot demonstrates the loose connection between the radial velocity and the turbulent velocity.
+The graph on the lower right again shows the density profile of this second particle.
+\label{fig:132_002_graphs}}
+\end{figure*}
+
+\begin{figure*}
 \plottwo{movie_disk_0132_0001.png}{movie_disk_0132_0003.png}
@@
 \label{fig:snapshots 2}}
-\end{figure*}
-
-\begin{figure*}
-\plottwo{Track_0132_000_rho.png}{pdf_frame_0132_0000.png}
-\plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
-\caption{Tracking 132 000  These plots show various data for one of the particles formed in the simulation.
-The upper left graph shows the density profile, with reference lines indicating slopes of -3/2, -2, and -5/2.
-The slope between $0.02$pc and $0.4$pc approximately matches the -5/2 reference line.
-The plot on the upper right shows the pdf function related to volumes of 1pc, 2pc, and 5pc.
-The bottom left shows a plot of the radial velocity and turbulent velocity associated with the particle.
-The lower right graph shows the radial density profile of the particle.
-\label{fig:132 000 graphs}}
-\end{figure*}
-
-\begin{figure*}
-\plottwo{Track_0132_002_rho.png}{pdf_frame_0132_0002.png}
-\plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
-\caption{Tracking 132 002  These plots show the data for a different particle formed in the simulation at the same time as the previous plots.
-The upper left graph of the density in the region neatly matches the -3/2 slope.
-The pdf graph in the upper right again shows the plots for volumes related to $1$pc, $2$pc, and $5$pc.
-The lower left plot demonstrates the loose connection between the radial velocity and the turbulent velocity.
-The graph on the lower right again shows the density profile of this second particle.
-\label{fig:132_002_graphs}}
 \end{figure*}
@@
 Here the radial density profile follows a $r^{-2}$ power law.
-\begin{figure*}
-\plottwo{Track_0256_000_rho.png}{Track_0132_000_rho.png}
-\caption{Density as a function of radial position from maximum density point.
+%\begin{figure*}
+%\plottwo{Track_0256_000_rho.png}{Track_0132_000_rho.png}
+%\caption{Density as a function of radial position from maximum density point.
 The density reaches a maximum of $10^6$ times mean density at $r \approx 10^{-3}$ pc and declines to approximately mean density at $r\approx 3$ pc.
@@
 \label{fig:good_rho_example}
-{bf\ This discusses left panel only, should we move the discussion outside the caption?}}
-\end{figure*}
+%{bf\ This discusses left panel only, should we move the discussion outside the caption?}}
+%\end{figure*}
 
@@
 
-\begin{figure*}
-\plottwo{Track_0256_000_vr.png}{Track_0132_000_vr.png}
-\caption{Radial velocity, $v_r$ (blue line), and turbulent velocity, $v_{\rm rms}$ (green line) as a function of radial position from maximum density point.
+%\begin{figure*}
+%\plottwo{Track_0256_000_vr.png}{Track_0132_000_vr.png}
+%\caption{Radial velocity, $v_r$ (blue line), and turbulent velocity, $v_{\rm rms}$ (green line) as a function of radial position from maximum density point.
 A $r^{-1/2}$ power law is shown as a red line.
 \label{fig:vr_example}}
-\end{figure*}
+%\end{figure*}
 
@@
 
-\begin{figure*}
-\plottwo{Track_0256_000_mass.png}{Track_0132_000_mass.png}
-\caption{WRITE CAPTION: Mass, $MSun$, as a function of radial position from a local maximum density point.     \label{fig:mass_example}}
-\end{figure*}
+%\begin{figure*}
+%\plottwo{Track_0256_000_mass.png}{Track_0132_000_mass.png}
+%\caption{WRITE CAPTION: Mass, $MSun$, as a function of radial position from a local maximum density point.     \label{fig:mass_example}}
+%\end{figure*}
 
@@
 \plottwo{Track_0256_000_l.png}{Track_0132_000_l.png}
-\caption{WRITE CAPTION: Mass, $MSun$, as a function of radial position from a local maximum density point.     \label{fig:angular_momentum_example}}
+\caption{WRITE CAPTION: Angular momentum as a function of distance from maximum density point.     \label{fig:angular_momentum_example}}
 \end{figure*}
 
-\begin{figure*}
-\plottwo{pdf_frame_0256.png}{pdf_frame_0132.png}
-\caption{Probability density function, $pdf$ as a function of density $rho/rho_0$.  \bf Verify range of slope calculation \label{fig:pdf_example}}
-\end{figure*}
+%\begin{figure*}
+%\plottwo{pdf_frame_0256.png}{pdf_frame_0132.png}
+%\caption{Probability density function, $pdf$ as a function of density $rho/rho_0$.  \bf Verify range of slope calculation \label{fig:pdf_example}}
+%\end{figure*}
 
-\begin{figure*}
-\plottwo{Disk_0132_0000.png}{Disk_0132_0004.png}
-\caption{Visualization of density profiles for two distinct particles in tracking1b. \label{fig:rho_example_132}}
-\end{figure*}
+%\begin{figure*}
+%\plottwo{Disk_0132_0000.png}{Disk_0132_0004.png}
+%\caption{Visualization of density profiles for two distinct particles in tracking1b. \label{fig:rho_example_132}}
+%\end{figure*}
 

### 2014-12-02 ms.tex
rearranged plots and update text

@@
 \end{equation}
-which when met causes the local grid to be refined by a factor of 2 provide that the maximum refinement level is not reached.
+which when met causes the local grid to be refined by a factor of 2 provided that the maximum refinement level is not reached.
 
@@
 2014.
+These clumpy regions have the highest densities and, hence, are prone to fulfill the criterion for star particle formation.
 
-The simulated volume has a large number of refined regions.
-We show three of them in Figure \ref{fig:snapshots} as slices perpendicular to the angular momentum vector.
-We determine the angular momentum vector by taking a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
-We also so a zoom out picture of the same region, which is the same for all three stars.
-While this is not unexpected as star formation is clustered, the large scale dynamics are determined by the larger region.
-
-A second family of two star particles were found in our simulation, show im Figure \ref{fig:snapshots 2}.
-Here again we show a zoom in and zoom out picture, comparing detail between the two particles around an axis of the angular momentum.
-Again the zoom in plots show two distinct star forming regions, while the zoom-out plot show that the larger region and hence large scale dynamics are the same.
+At the instance in time shown in Figure \ref{fig:entire projection}, a total of five star particles have been formed.
+However, these star particles are not formed in isolation.
+Rather they congregate in two star forming regions shown in Figure \ref{fig:star forming regions}.
+In these regions, the sink particles that are formed lay within 0.1 pc of one another, which reflects that the highest density regions are concentrated toward the center.
+Visually, the character of accretion at the this outer scale appear to flow along filaments, thought previous work (Lee et al.
+2014) has shown that the majority of accretion is spherical.
+We discuss this more quantitatively below {\bf INCLUDE A STUDY OF LARGE SCALE ACCRETION}
 
@@
 \plottwo{Disk_0132_0000.png}{Disk_0132_0002.png}
-\caption{Cut of the two star forming regions.}
+\caption{Cut of the two star forming regions.\label{fig:star forming regions}}
 \end{figure*}
 
-\begin{figure*}
-\plottwo{Disk_0132_0000.png}{Track_0132_000_rho.png}
-\plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
-\plottwo{Track_0132_000_mass.png}{pdf_frame_0132_0000.png}
-\caption{Tracking 132 000  These plots show various data for one of the particles formed in the simulation.
-The upper left graph shows the density profile, with reference lines indicating slopes of -3/2, -2, and -5/2.
-The slope between $0.02$pc and $0.4$pc approximately matches the -5/2 reference line.
-The plot on the upper right shows the pdf function related to volumes of 1pc, 2pc, and 5pc.
-The bottom left shows a plot of the radial velocity and turbulent velocity associated with the particle.
-The lower right graph shows the radial density profile of the particle.
-\label{fig:132 000 graphs}}
-\end{figure*}
+The first family of three star particles is shown in Figure \ref{fig:snapshots} and the second family of two star particles is shown in Figure \ref{fig:snapshots 2}.
+We plot these particles and their surrounding $4\times 10^{-2}$ pc region as slices perpendicular to the local angular momentum vector.
+We determine the local angular momentum vector by taking a $0.01$ pc region about the star particle to calculate the angular momentum vector.
+It is clear from these plots that the clumpy filamentary nature of the local star forming environment extends down to much smaller scales than we have been able to probe in previous work (Lee et al.
+2014).
 
-\begin{figure*}
-\plottwo{Disk_0132_0002.png}{Track_0132_002_rho.png}
-\plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
-\plottwo{Track_0132_002_mass.png}{pdf_frame_0132_0002.png}
-\caption{Tracking 132 002  These plots show the data for a different particle formed in the simulation at the same time as the previous plots.
-The upper left graph of the density in the region neatly matches the -3/2 slope.
-The pdf graph in the upper right again shows the plots for volumes related to $1$pc, $2$pc, and $5$pc.
-The lower left plot demonstrates the loose connection between the radial velocity and the turbulent velocity.
-The graph on the lower right again shows the density profile of this second particle.
-\label{fig:132_002_graphs}}
-\end{figure*}
+{\bf INCLUDE ACCRETION TOPOLOGY STUDY HERE.}
+
+At small radii (similar to $10^{-2}$ pc), a protostellar disk appears in all five snapshots.
+The radial size of these disks varies.
+These disks also show a coherent sense of rotation about a common axis, e.g., no counter-rotating disks.
+In addition, some of the snapshots show multiple high density regions.
+This could be a result of accretion of high density clumps or fragmentation in the disk or streams.
+
+We also include a zoom-out picture of the same region, which is the same for all three stars (lower right plot in Figure \ref{fig:snapshots}, bottom plot in Figure \ref{fig:snapshots 2}).
+While this is not unexpected as star formation is clustered, the large scale dynamics are determined by the larger region.
+We should note that these star particles are within about 0.1 pc of each other in their respective families.
+
+{\bf ARE THESE DISK GRAVITATIONALLY STABLE?}
 
@@
 
-The zoom in plots in Figures \ref{fig:snapshots} and \ref{fig:snapshots 2} suggest that the accretion can occur by streams or by spherical flows.
-However, the degree by which these modes of accretion dominate, is not clear.
-At small radii (similar to $10^{-2}$ pc), a protostellar disk appears in all four snapshots.
-The radial size of these disks varies.
-These disks also show a coherent sense of rotation about a common axis, e.g., no counter-rotating disks.
-In addition, some of the snapshots show multiple high density regions.
-This could be a result of accretion of high density clumps or fragmentation in the disk or streams.
-
-\subsection{Examples of Collapse}
+\subsection{Radial Profile of Collapse}
 
@@
 \begin{figure*}
+\plottwo{Disk_0132_0000.png}{Track_0132_000_rho.png}
+\plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
+\plottwo{Track_0132_000_mass.png}{pdf_frame_0132_0000.png}
+\caption{Tracking 132 000  These plots show various data for one of the particles formed in the simulation.
+The upper left graph shows the density profile, with reference lines indicating slopes of -3/2, -2, and -5/2.
+The slope between $0.02$pc and $0.4$pc approximately matches the -5/2 reference line.
+The plot on the upper right shows the pdf function related to volumes of 1pc, 2pc, and 5pc.
+The bottom left shows a plot of the radial velocity and turbulent velocity associated with the particle.
+The lower right graph shows the radial density profile of the particle.
+\label{fig:132 000 graphs}}
+\end{figure*}
+
+\begin{figure*}
+\plottwo{Disk_0132_0002.png}{Track_0132_002_rho.png}
+\plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
+\plottwo{Track_0132_002_mass.png}{pdf_frame_0132_0002.png}
+\caption{Tracking 132 002  These plots show the data for a different particle formed in the simulation at the same time as the previous plots.
+The upper left graph of the density in the region neatly matches the -3/2 slope.
+The pdf graph in the upper right again shows the plots for volumes related to $1$pc, $2$pc, and $5$pc.
+The lower left plot demonstrates the loose connection between the radial velocity and the turbulent velocity.
+The graph on the lower right again shows the density profile of this second particle.
+\label{fig:132_002_graphs}}
+\end{figure*}
+
+\begin{figure*}
 \plottwo{Track_0256_000_l.png}{Track_0132_000_l.png}

### 2014-12-09 ms.tex
Updated captions for fig:snapshots and snapshots2

@@
 \plottwo{movie_disk_0132_0000.png}{Disk_0132_0000.png}
-\caption{{\bf Change plots to show a zoom out for 132 and zoom in on the three regions that are the same} Slices along the angular momentum axis of four zoomed in simulation.
+\caption{The upper frames, and the bottom left frame show slices along the angular momentum axis of three local particles in one of the regions demonstrating significant star formation.
+Each of the zoomed-in slices have a radius of $0.02$ pc.
 To determine the angular momentum axis, we take a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
-We then take a slice in the plane normal to the angular momentum to plot the regions around these dense points.
+We then take a slice in the plane normal to the angular momentum vector to plot the regions around these dense points.
+The bottom right frame shows the zoomed-out region of formation, with a radius of $1.5$ pc.
 \label{fig:snapshots}}
@@
 \centering{\includegraphics[width=0.48\textwidth]{Disk_0132_0002.png}}
-\caption{{\bf Do same for other plots in 132} Slices along the angular momentum axis of four zoomed in simulation.
-To determine the angular momentum axis, we take a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
-We then take a slice in the plane normal to the angular momentum to plot the regions around these dense points.
+\caption{The upper frames show the zoomed-in slices along the angular momentum axis of the local particles in the second region of significant star production.
+Again, each of the zoomed-in slices have a radius of $0.02$ pc.
+The bottom frame shows the zoomed out region, with a radius of $1.5$ pc.
 \label{fig:snapshots 2}}

### 2014-12-23 ms.tex
Added Average.pdf files, caption, and discussion

@@
 In these regions, the sink particles that are formed lay within 0.1 pc of one another, which reflects that the highest density regions are concentrated toward the center.
-Visually, the character of accretion at the this outer scale appear to flow along filaments, thought previous work (Lee et al.
+Visually, the character of accretion at the this outer scale appear to flow along filaments, though previous work (Lee et al.
 2014) has shown that the majority of accretion is spherical.
@@
 
+\begin{figure*}
+\plottwo{Average_5_013.pdf}{Average_5_24.pdf}
+\plottwo{Average_02_013.pdf}{Average_02_24.pdf}
+\caption{Here we show the mass accretion rates as a function of density.
+The upper figures are displaying rates corresponding to a region of $0.5$ pc around the particle, while the lower figures are showing a radius of $0.02$ pc.
+The two plots on the left show the region with three particles forming, and the plots on the right show the region with two particles forming.
+The more massive region with three particles indicates the mass is coming in at a density value roughly three times the average density, and the curves are similar between the smaller and larger radii.
+The graph for the less massive region on the right shows a distinctly separate profile between the larger and the smaller radius.
+\label{fig:accretion average rho}}
+\end{figure*}
+
+In prior work, we examined the geometry of accretion, i.e., does the accretion proceed along filament or over 4$\pi$ steradian, by plotting the cumulative accretion rate as a function of density.
+In Figure \ref{fig:accretion average rho}, we plot the cumulative accretion rate as a function of density, normalized to the average density in shells of radii 0.5 (upper plots) and 0.02 pc (lower plots).
+Plots on the left are associated with the more massive star forming region, where 3 star particles reside and plots on the right are associated with the less massive star forming region, where two star particles reside.
+The density where 50\% of the accretion is accounted for occurs at $\rho/\bar{\rho} \approx 3$ for the more massive star forming region and appears to be independent of whether the shell has a radius of 0.5 or 0.02 pc.
+On the other hand, the less massive star forming region has half of the accretion account for at $\rho/\bar{\rho} \approx 10$ at r=0.5 pc.
+This suggests that the accretion is directed along filaments.
+This filamentary nature of accretion does not maintain itself to lower radii, where at 0.02 pc, the 50\% point in cumulatiove accertion occurs at $\rho/\bar{\rho} \approx 1$, which suggest accretion along $4\pi$ steradian.
 {\bf INCLUDE ACCRETION TOPOLOGY STUDY HERE.}

### 2015-01-13 ms.tex
Brought geometry.pdf files to paper- need captions
Add Mdot, & disc. more mass. region & start less

@@
 
-\begin{figure*}
-\plottwo{Average_5_013.pdf}{Average_5_24.pdf}
-\plottwo{Average_02_013.pdf}{Average_02_24.pdf}
-\caption{Here we show the mass accretion rates as a function of density.
-The upper figures are displaying rates corresponding to a region of $0.5$ pc around the particle, while the lower figures are showing a radius of $0.02$ pc.
-The two plots on the left show the region with three particles forming, and the plots on the right show the region with two particles forming.
-The more massive region with three particles indicates the mass is coming in at a density value roughly three times the average density, and the curves are similar between the smaller and larger radii.
-The graph for the less massive region on the right shows a distinctly separate profile between the larger and the smaller radius.
-\label{fig:accretion average rho}}
-\end{figure*}
-
-In prior work, we examined the geometry of accretion, i.e., does the accretion proceed along filament or over 4$\pi$ steradian, by plotting the cumulative accretion rate as a function of density.
-In Figure \ref{fig:accretion average rho}, we plot the cumulative accretion rate as a function of density, normalized to the average density in shells of radii 0.5 (upper plots) and 0.02 pc (lower plots).
-Plots on the left are associated with the more massive star forming region, where 3 star particles reside and plots on the right are associated with the less massive star forming region, where two star particles reside.
-The density where 50\% of the accretion is accounted for occurs at $\rho/\bar{\rho} \approx 3$ for the more massive star forming region and appears to be independent of whether the shell has a radius of 0.5 or 0.02 pc.
-On the other hand, the less massive star forming region has half of the accretion account for at $\rho/\bar{\rho} \approx 10$ at r=0.5 pc.
-This suggests that the accretion is directed along filaments.
-This filamentary nature of accretion does not maintain itself to lower radii, where at 0.02 pc, the 50\% point in cumulatiove accertion occurs at $\rho/\bar{\rho} \approx 1$, which suggest accretion along $4\pi$ steradian.
-{\bf INCLUDE ACCRETION TOPOLOGY STUDY HERE.}
-
 At small radii (similar to $10^{-2}$ pc), a protostellar disk appears in all five snapshots.
@@
 We should note that these star particles are within about 0.1 pc of each other in their respective families.
-
-{\bf ARE THESE DISK GRAVITATIONALLY STABLE?}
 
@@
 
+{\bf ARE THESE DISK GRAVITATIONALLY STABLE?}
+
+\begin{figure*}
+\plottwo{Average_5_013.pdf}{Average_5_24.pdf}
+\plottwo{Average_02_013.pdf}{Average_02_24.pdf}
+\caption{Here we show the mass accretion rates as a function of density.
+The upper figures are displaying rates corresponding to a region of $0.5$ pc around the particle, while the lower figures are showing a radius of $0.02$ pc.
+The two plots on the left show the region with three particles forming, and the plots on the right show the region with two particles forming.
+The more massive region with three particles indicates the mass is coming in at a density value roughly three times the average density, and the curves are similar between the smaller and larger radii.
+The graph for the less massive region on the right shows a distinctly separate profile between the larger and the smaller radius.
+\label{fig:accretion average rho}}
+\end{figure*}
+
+In prior work, we examined the geometry of accretion, i.e., does the accretion proceed along filament or over 4$\pi$ steradian, by plotting the cumulative accretion rate as a function of density.
+In Figure \ref{fig:accretion average rho}, we plot the cumulative accretion rate as a function of density, normalized to the average density in shells of radii 0.5 (upper plots) and 0.02 pc (lower plots).
+Plots on the left are associated with the more massive star forming region, where 3 star particles reside and plots on the right are associated with the less massive star forming region, where two star particles reside.
+The density where 50\% of the accretion is accounted for occurs at $\rho/\bar{\rho} \approx 3$ for the more massive star forming region and appears to be independent of whether the shell has a radius of 0.5 or 0.02 pc.
+This suggests that the accretion is proceeding along filaments.
+Similarly, the less massive star forming region has half of the accretion account for at $\rho/\bar{\rho} \approx 10$ at r=0.5 pc, again suggesting filamentary accretion.
+However, this filamentary nature of accretion does not maintain itself to lower radii, where at 0.02 pc, the 50\% point in cumulatiove accertion occurs at $\rho/\bar{\rho} \approx 1$, which suggest accretion is spherical.
+
+Figures \ref{fig:geometry} and \ref{fig:geometry 2} show the moments of inertia for the three star particles in the more massive and less massive star forming regions, respectively.
+For the more massive star forming region (Fig.
+\ref{fig:geometry}), note that the moments of inertia are identical for radii greater than $2e-2$, due the fact that these three star particles sit in the same massive star forming region.
+This is similarly true for the less massive region for $r\gtrsim 0.1$ pc.
+At small radii, the distribution of gas appears to be flattened or rod-like, which is implied by the relative values of the three principle axes.
+Around $r\approx 0.02$ and $r \approx 0.1$ pc, respectively, the geometry of the gas is relatively spherical.
+Outside this region, the gas become flatter again.
+However, the flattened distributions seen at small and larger radii may also reflect the effect of streams of gas or gravitational instabilities that are associated with the accretion.
+
+\begin{figure}
+\plotone{geometry_4_cum.pdf}
+\caption{Plots of the moments of inertia for the three star particles in the more massive star forming region.
+Note that the moments of inertia are identical for radii greater than $2e-2$, due the fact that these three star particles sit in the same massive star forming region.
+Inside this region, we find that the two largest moments of inertia are close in value to each other, suggesting that the geometry of the gas resembles a flattened distribution.
+Around $r\approx 0.02$ pc, the geometry of the gas becomes relatively spherical.
+Outside this region, the gas become flatter again.
+However, the flattened distributions seen at small and larger radii may also reflect the effect of streams of gas that might drive the accretion.
+\label{fig:geometry}}
+\end{figure}
+
+\begin{figure}
+\plotone{geometry_4_cum.pdf}
+\caption{$Geometry_2$ and $geometry_4$ Plots of the moments of inertia for the two star particles in the less massive star forming region.
+The moments of inertia for these particles are identical down to a radius of $r\approx 0.1$ pc
+\label{fig:geometry 2}}
+\end{figure}
+
 \subsection{Radial Profile of Collapse}
@@
 These simulations display much more complicated dynamics than is captured by the simple analytic theory of Murray and Chang 2014.
-In spite of this, we compare radially averaged profiles around these high density points to their analytic theory.
-In the left panel of Figure \ref{fig:good_rho_example}, we show the radial density profile around one such high density point.
+In spite of this, we proceed to compare radially averaged profiles around these high density points to their analytic theory.
+Toward this end, we construct radial profiles around the most massive star particle in each of the star forming regions.
+
+In Figure \ref{fig:132 000 graphs}, we show the radial distribution of $\rho$, $v_r$, $v_t$, $M(r)$, $\dot{M}$, and the density distribution function around the three star particles.
+In the top left plot, we show a slice of the region perpendicular to the z-axis.
+The top right plot shows $\rho(r)$ alongside lines of $r^{-3/2}$, $r^{-2}$ and $r^{-5/2}$.
+Here we find that at small radii, $\rho(r)\sim r^{-2}$ at small radii and $\rho(r) \sim r^{-3/2}$ between 0.1 - 1 pc.
+At other radii, the density profile is surprisingly steep, steeper than even $r^{-5/2}$.
+
+The velocity profiles show similar change in behavior in the transition increasing from small radii, around $r\approx 0.02$ pc.
+The radial velocity tends to increase to $0.02$ pc, and then decreases following the slope noted in the red line of 0.5.
+This demarcation from the smaller radii to larger is due to the disk-like shape of the star particle at the smaller radii.
+The turbulent velocity, green line, decreases inward.
+
+The middle right graph shows the mass as a function of radius.
+The mass monotonically increases.
+However there is a flattening around $0.1$ pc before it steepens again.
+This is important because this region behaves similarly to a point mass.
+Which is also reflected in the density profile and the radial velocity profile where $\rho \sim r^-1.5$ and $v_r \sim r^{-1/2}$ as shown by MC15 around point masses.
+
+The bottom left graph shows the mass accretion rates for all three particles in the more massive region.
+At small radii, the drop off in $\dot M$ is due to the presence of a disk, the drop off at larger radii is due to the step density profile.
+
+The bottom right graph depicts the density PDF for three different radii.
+Here it roughly corresponds to the density profile shown in the top right plot.
+For instance, the PDF inside a 1 pc sphere has a slope of $\rho ^-1.3$ which corresponds to a radial profile of $\rho \sim r^-2.5$, whereas for a 2 pc sphere, the slope is now $\rho ^-1.5$ corresponding to radial profile of $\rho \sim r^-2$.
+
+The comparative plots for the less massive region are shown in \ref{fig:132_002_graphs}.
+The upper left graph shows a slice perpendicular to the z axis in the plane of peak density.
+The graph shows a single filamentary strand of accretion towards the inner disk-like structure.
+
+Here again the top right plot shows $\rho(r)$ alongside lines of $r^{-3/2}$, $r^{-2}$ and $r^{-5/2}$.
+In this instance, however, the slope of $\rho(r)$ is approximately $r^-2$ for $r \gtrsim 0.1$ pc as well as for $r \lesssim 5e-3$ pc.
+Between radii $r \eqsim 0.01$ pc and $r \eqsim 0.1$ pc, $\rho$ closely matches the slope of $r^{-3/2}$.
+There is also a sharp spike in density at $r = 6e-3$ pc that is notable in the mass plot as well.
+
+%In the left panel of Figure \ref{fig:good_rho_example}, we show the radial density profile around one such high density point.
 As this figure shows, the radial density profile follows a $r^{-3/2}$ power law in agreement with analytic expections.
@@
 \plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
-\plottwo{Track_0132_000_mass.png}{pdf_frame_0132_0000.png}
-\caption{Tracking 132 000  These plots show various data for one of the particles formed in the simulation.
-The upper left graph shows the density profile, with reference lines indicating slopes of -3/2, -2, and -5/2.
-The slope between $0.02$pc and $0.4$pc approximately matches the -5/2 reference line.
-The plot on the upper right shows the pdf function related to volumes of 1pc, 2pc, and 5pc.
-The bottom left shows a plot of the radial velocity and turbulent velocity associated with the particle.
-The lower right graph shows the radial density profile of the particle.
+\plottwo{Track_0132_003_mdot.png}{pdf_frame_0132_0000.png}
+\caption{Tracking 132 000  These plots show various data for one of the particles formed in the more massive region in the simulation.
+The upper left plot shows the zoomed out region with radius of $1.5$ pc.
+The upper right graph shows the density profile for one of the particles in the region with reference lines indicating slopes of $-5/2$, $-2$, and $-3/2$.
+The slope between $0.02$pc and $0.4$pc approximately matches the $-5/2$ reference line.
+The middle left shows a plot of the radial velocity and turbulent velocity associated with the particle.
+The middle right graph shows the radial density profile of the particle.
+The plot on the lower left compares the mass accretion rates between the three particles in the more massive region.
+The accretion rates are very similar at radii greater than $r \approx 10^{-2}$ pc.
+The plot on the lower right shows the pdf function related to volumes of 1pc, 2pc, and 5pc.
 \label{fig:132 000 graphs}}
@@
 \plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
-\plottwo{Track_0132_002_mass.png}{pdf_frame_0132_0002.png}
+\plottwo{Track_0132_004_mdot.png}{pdf_frame_0132_0002.png}
 \caption{Tracking 132 002  These plots show the data for a different particle formed in the simulation at the same time as the previous plots.

### 2015-01-13 ms.tex
Work on intro

@@
 
-\section{Analytic Expectations}
+The rapid formation of stars remains an unsolved problem in astrophysics.
 
-Recently, Murray \& Chang (2014) considered the spherical collapse of isothermal turbulently supported gas accounting for adiabatic heating from compression of the turbulence.
-In particular,  {\bf SHOW EQUATIONS}
-Their basic results are that the run of density asymptotes to 
+Here we will focus on the physics of rapid star formation and accretion.  An early model of Shu (1977) estimated the accretion rate onto stars by 
+assuming that stars form from hydrostatic cores
+supported by thermal gas pressure. The accretion
+rate in his model was independent of time, given
+by $M = m0c_s^3/G$, where $c_s$  is the
+sound speed in molecular gas, and $m_0 = 0.975$.
+Shu (1977) predicted a maximum accretion rate was $yr^{-1}$, which is too small to explain
+the origin of massive stars.
+
+Later workers overcame the difficulty with the accretion rate by adopting the turbulence speed in lieu of the sound speed (Myers \& Fuller 1992, McLaughlin \& Pudritz 1997, McKee \& Tan 2003).
+In doing so they were able to replace the slower signal speed of sound with the faster turbulent speed.
+However, they rely on assuming that a hydrostatic core supported by turbulent pressure is the initial condition and that turbulence is static and unaffected by the collapse.
+
+Recently Murray \& Chang (2015) developed a 1-D model of spherical collapse that treat turbulence as a dynamical variable and does not assume that the initial condition is a hydrostatically supported region.
+They were able to accomplished this by taking advantage of the recent results of Robertson \& Goldreich (2012) on compressible turbulence.
+In particular
+
+{\bf need equation}
+
+Hence, this allowed MC15 to include an energy equation for the turbulent velocity that alongside the equations for mass continuity and momentum gives a set of equations that can be solved in spherical symmetry numerically.
+In addition, they were able to analytically show that the results of their calculations gave density and velocity profiles that appear to be in line with both recent numerical calculations (Lee et al 2015) and observations.
+
+In particular, their basic results are that the run of density asymptotes to 
 \be
@@
 %
+
+These results compared favorable with the results of numerical simulations (Lee et al. 2015).  However, it is unclear if the results of
 
@@
 The comparative plots for the less massive region are shown in \ref{fig:132_002_graphs}.
-The upper left graph shows a slice perpendicular to the z axis in the plane of peak density.
+The upper left graph shows a slice perpendicular to the z axis in the plane of peak density around this region of star formation.
 The graph shows a single filamentary strand of accretion towards the inner disk-like structure.
@@
 Here again the top right plot shows $\rho(r)$ alongside lines of $r^{-3/2}$, $r^{-2}$ and $r^{-5/2}$.
-In this instance, however, the slope of $\rho(r)$ is approximately $r^-2$ for $r \gtrsim 0.1$ pc as well as for $r \lesssim 5e-3$ pc.
-Between radii $r \eqsim 0.01$ pc and $r \eqsim 0.1$ pc, $\rho$ closely matches the slope of $r^{-3/2}$.
-There is also a sharp spike in density at $r = 6e-3$ pc that is notable in the mass plot as well.
+Between $0.01\,{\rm pc} \lesssim r \lesssim 0.1$ pc, $\rho$ closely matches the slope of $r^{-3/2}$.
+This same region shows a flattening in the mass profile (middle right plot) and a radial velocity profile that is similar to $r^{-1/2}$ (middle left plot).
+This region also shows flat $\dot{M}$ profile (lower left plot).
+For larger radii, the density profile shows a break toward a steeper power law in $r$ and the velocity profiles shows a switch to rising with $r$ rather than falling with $r$.
+The mass profile shows a continued monotonic increase with $r$, indicating that the mass is dominated by the gas and not by a central mass.
+
+In addition, a sharp spike in density appears at $r = 6e-3$ pc that is notable in the velocity, mass, and $\dot M$, plots as well.
+The plot of moments of inertia also indicate that this point is where the two particles begin to diverge most significantly.
+
+The middle left plot shows the velocity profiles, which also indicate an anomoly at $r \eqsim 6e-3$ pc.
+However, the general trend of both turbulent and radial velocity is increasing inward from $r \eqsim 0.1$ pc, and also increasing outward from $r \eqsim 0.1$ pc.
+
+The mass distribution, the middle right plot, again increases monotonically.  The flattening in this less massive region occurs at the notable radius of $r \eqsim 6e-3$ pc, before steepening again.
 
@@
 %{bf\ This discusses left panel only, should we move the discussion outside the caption?}}
-%\end{figure*}
-
-However, to make a comparison of the average behavior of high density star forming regions, we average over a number
-
-What plots should we show?
--velocities vs r
--angular momentum vs r
--mass vs r
--mdot vs r
-
-In Figure \ref{fig:vr_example}, we plot $v_r$ (blue line) and $v_{\rm rms}$ (green line) as a function of $r$.
-In the left panel, the radial velocity curve is sporadic, but the general trend is that $v_r$ declines in toward $0.1$ pc, but then rises inward of this point.
-Plotted in red is a $r^{-1/2}$ power law and this inward rise appears to follow this general trend.
-The turbulent velocity is smoother, but it follows the general trend of the $v_r$ curve in that it falls slowly inward toward $]\approx 0.1 $ pc before rising inward of that, roughly tracking the $r^{-1/2}$ power law.
-These general trends are in rough agreement with our analytic expectations.
-In the right panel, the radial velocity curve is smoother and increases inwards towards $0.01$ pc, tracking the $r^{-1/2}$ power law fairly closely.
-This radial velocity begins to fall inward of $0.01$ pc, possibly due to rotational support from the angular momentum of a disk.
-The turbulent velocity in the right panel seems to roughly follow an inverse relationship with the radial velocity.
-As with the left panel, however, the turbulent velocity is significantly less volatile than the radial velocity and varies only slightly with $r$.
-These expectations are in line with the analytic results of Murray and Chang (2014), albeit with a somewhat larger scale that they suggest.
-
-%\begin{figure*}
-%\plottwo{Track_0256_000_vr.png}{Track_0132_000_vr.png}
-%\caption{Radial velocity, $v_r$ (blue line), and turbulent velocity, $v_{\rm rms}$ (green line) as a function of radial position from maximum density point.
-A $r^{-1/2}$ power law is shown as a red line.
-\label{fig:vr_example}}
-%\end{figure*}
-
-A hint of why these different behaviors emerge from looking at the $M(r)$.
-In Figure \ref{fig:mass_example}, we plot the mass as a function of radius.
-Regions where $M(r)$ flattens implies that the mass in this region can be treated using a point mass approximation.
-Here it is apparent that the regions that can be treated a point mass corresponds to regions in where the density follows a $r^{-3/2}$ power law and the radial and turbulent velocities follow a $r^{-1/2}$ power law.
-This is due to mass concentration at smaller radii as evident by the sudden drop in $M(r)$ at $r \approx 10^{-2} - 1$ pc.
-The flattening of $M(r)$ on larger scales is interesting as this is due to the entire mass of the forming star cluster as opposed to the individual stars that MC14 originally envisioned.
-Hence the entire mass of the star cluster contributes to the large scale dynamics of turbulent collapse.
-
-%\begin{figure*}
-%\plottwo{Track_0256_000_mass.png}{Track_0132_000_mass.png}
-%\caption{WRITE CAPTION: Mass, $MSun$, as a function of radial position from a local maximum density point.     \label{fig:mass_example}}
 %\end{figure*}
@@
 \end{document}
+
+However, to make a comparison of the average behavior of high density star forming regions, we average over a number
+
+What plots should we show?
+-velocities vs r
+-angular momentum vs r
+-mass vs r
+-mdot vs r
+
+In Figure \ref{fig:vr_example}, we plot $v_r$ (blue line) and $v_{\rm rms}$ (green line) as a function of $r$.
+In the left panel, the radial velocity curve is sporadic, but the general trend is that $v_r$ declines in toward $0.1$ pc, but then rises inward of this point.
+Plotted in red is a $r^{-1/2}$ power law and this inward rise appears to follow this general trend.
+The turbulent velocity is smoother, but it follows the general trend of the $v_r$ curve in that it falls slowly inward toward $]\approx 0.1 $ pc before rising inward of that, roughly tracking the $r^{-1/2}$ power law.
+These general trends are in rough agreement with our analytic expectations.
+In the right panel, the radial velocity curve is smoother and increases inwards towards $0.01$ pc, tracking the $r^{-1/2}$ power law fairly closely.
+This radial velocity begins to fall inward of $0.01$ pc, possibly due to rotational support from the angular momentum of a disk.
+The turbulent velocity in the right panel seems to roughly follow an inverse relationship with the radial velocity.
+As with the left panel, however, the turbulent velocity is significantly less volatile than the radial velocity and varies only slightly with $r$.
+These expectations are in line with the analytic results of Murray and Chang (2014), albeit with a somewhat larger scale that they suggest.
+
+%\begin{figure*}
+%\plottwo{Track_0256_000_vr.png}{Track_0132_000_vr.png}
+%\caption{Radial velocity, $v_r$ (blue line), and turbulent velocity, $v_{\rm rms}$ (green line) as a function of radial position from maximum density point.
+A $r^{-1/2}$ power law is shown as a red line.
+\label{fig:vr_example}}
+%\end{figure*}
+
+A hint of why these different behaviors emerge from looking at the $M(r)$.
+In Figure \ref{fig:mass_example}, we plot the mass as a function of radius.
+Regions where $M(r)$ flattens implies that the mass in this region can be treated using a point mass approximation.
+Here it is apparent that the regions that can be treated a point mass corresponds to regions in where the density follows a $r^{-3/2}$ power law and the radial and turbulent velocities follow a $r^{-1/2}$ power law.
+This is due to mass concentration at smaller radii as evident by the sudden drop in $M(r)$ at $r \approx 10^{-2} - 1$ pc.
+The flattening of $M(r)$ on larger scales is interesting as this is due to the entire mass of the forming star cluster as opposed to the individual stars that MC14 originally envisioned.
+Hence the entire mass of the star cluster contributes to the large scale dynamics of turbulent collapse.
+
+%\begin{figure*}
+%\plottwo{Track_0256_000_mass.png}{Track_0132_000_mass.png}
+%\caption{WRITE CAPTION: Mass, $MSun$, as a function of radial position from a local maximum density point.     \label{fig:mass_example}}
+%\end{figure*}

### 2015-05-28 ms.tex
minor changes to the ms.tex and included a new file for the whole simulation named movieframe_quad2_0498.png

@@
 rate in his model was independent of time, given
-by $M = m0c_s^3/G$, where $c_s$  is the
+by $M = m_0c_s^3/G$, where $c_s$  is the
 sound speed in molecular gas, and $m_0 = 0.975$.
@@
 
-Recently Murray \& Chang (2015) developed a 1-D model of spherical collapse that treat turbulence as a dynamical variable and does not assume that the initial condition is a hydrostatically supported region.
+Recently Murray \& Chang (2015), hereafter MC15, developed a 1-D model of spherical collapse that treat turbulence as a dynamical variable and does not assume that the initial condition is a hydrostatically supported region.
 They were able to accomplished this by taking advantage of the recent results of Robertson \& Goldreich (2012) on compressible turbulence.
@@
 ) to model isothermal, self-gravitating, hydrodynamic turbulence on isothermal gas with three-dimensional (3D),
-periodic grids and 10 levels of refinement on a root grid of $128^3$, giving an effective resolution of $128K^3$.  
+periodic grids and 8 levels of refinement on a root grid of $128^3$, giving an effective resolution of $32K^3$.  
 Self-gravity is computed with a multi-grid Poisson solver (see Ricker
 2008), coupled with a fast-Fourier transform solution on the root grid.
+Similar to (Lee et.
+al Feb.
+2015) our FLASH runs use the Harten-Lax-van Leer-Contact Riemann solver and an unsplit solver.
+({\bf Lee et.
+al use Lee \& Deane (2009) do we?})
 
-To initialize our simulations, we drive turbulence by applying a large scale ($1 \le k \le 2$) solenoidal 
-acceleration field as a momentum and energy source term.  We apply this field in the absence of gravity and sink particle formation for 3 dynamical times until a statistical steady state is reached.
+We start with a box with the physical length set to $L = 16$pc using periodic boundary condtions.
+The intial mass density is $\rho = 3x10^{-22}$ and the isothermal, ambient gas is set to a sound speed of $c_s = 2.64x10^{4}$ {\bf we also set $Mach_ambient = 5.0$ include?}.
+To initialize our simulations, we drive turbulence by applying a large scale ($1 \le kL \le 2$) solenoidal acceleration field as a momentum and energy source term.
+We apply this field in the absence of gravity and sink particle formation for 3 dynamical times until a statistical steady state is reached.
 {\bf WHY JUST SOLENOIDAL}
@@
 As gas collapse under self gravity, certain regions rapidly increase in density.
-These regions are refine when the Truelove criterion ($\lambda \le 4 \Delta x$) is met.
+These regions are refined when the Truelove criterion ($\lambda \le 4 \Delta x$) is met.
 This corresponds to a condition on the density
 \begin{equation}
-refinement criteria
+%\frac{\rho}{\rho_0} = 45 \text{x} 4^{l} (\frac{N_{root}}{128})^2 (\frac{16\text{pc}}{L})^2 (\frac{c_s}{2.65x10^4 \text{cm s})^2 (\frac{3x10^{-22}\text{g cm^-3}}{\rho_0})
+\label{eq:refinement_criteria}
 \end{equation}
@@
 (2014), our sink particle creation prescription is different from the prescription of Federrath et al (???) where additional checks need to be performed.
+Our base grid's resolution of $128^3$ gives a cell length of $1.2 \text{x} 10^{-1}$pc which is sufficient to resolve the Jean's length
+\begin{equation}
+\lambda_J \equiv \sqrt{\frac{\pi c_{s}^2}{G \bar{\rho}}} \approx 3.5 \text{pc}
+\end{equation}
 
@@
 \begin{figure}
-\plotone{frame0241.png}
-\caption{Projection along the x-axis of the entire simulation volume.
-The root grid is $128^3$ with up to 10 levels of refinement, giving an effective resolution of $128\,{\rm K}^3$.
-This snapshot is taken at ??.
+\plotone{movieframe_quad2_0498.png}
+\caption{Projection of the density along the z-axis of the entire simulation volume.
+The root grid is $128^3$ with up to 8 levels of refinement, giving an effective resolution of $32\,{\rm K}^3$.
+This snapshot is taken $2.8$Myr after star formation is turned on.
 \label{fig:entire projection}}

### 2015-05-28 sphereAcc_v1.py
scripts for mass accretion rate

@@ -1,0 +1,164 @@
+import matplotlib
+matplotlib.use('Agg') #I tried a few other backends and got the same result
+from matplotlib import rc
+rc("text", usetex=True)
+from matplotlib import rcParams
+rcParams.update({'figure.autolayout': True})
+
+import matplotlib.pyplot as p 
+import numpy
+from math import pi
+import Image
+import os
+import argparse
+
+quad = os.getcwd()[-5:]
+seed = os.getcwd()[-13:-6]
+parser = argparse.ArgumentParser(description = "Radius")
+parser.add_argument('Radius', metavar='N1', type=float)
+args = parser.parse_args()
+sizeSphere = args.Radius
+
+r = 0.02*3e18
+def plotSphAccretion( filename, fileout, histout) : 
+	a = numpy.load(filename)
+	rhodotSph = a['rhodotSph']
+	rhoSph = a['rhoSph']
+	aSph = a['aSph']
+	print filename, rhodotSph.size
+	img = rhodotSph
+	file1 = "filename1_0_{0}.png".format(sizeSphere)
+	file2 = "filename2_0_{0}.png".format(sizeSphere)
+
+	p.figure(figsize=(15,10))
+	p.axes(projection='mollweide',aspect=2)
+	p.imshow(img,extent=(-pi,pi,-pi/2.,pi/2.),aspect=0.5)
+	p.colorbar(orientation="horizontal")
+	p.savefig(file1)
+
+	img = numpy.log10(rhoSph)
+	p.figure(figsize=(15,10))
+	p.axes(projection='mollweide',aspect=2)
+	p.imshow(img,extent=(-pi,pi,-pi/2.,pi/2.),aspect=0.5)
+	p.colorbar(orientation="horizontal")
+	p.savefig(file2)
+
+	newImage = Image.new("RGB", (1500,2000))
+	image1 = Image.open(file1) 
+	image2 = Image.open(file2) 
+
+	newImage.paste( image1, (0,0)) 
+	newImage.paste( image2, (0,1000)) 
+	newImage.save(fileout) 
+
+	p.clf()
+	num_bins = 100
+	rhodotSphFlat = -rhodotSph.flatten()
[…]

### 2015-05-28 reduce_scripts/all_profiles.py
adding reduction scripts

@@ -1,0 +1,710 @@
+import matplotlib
+matplotlib.use("Agg")
+
+from yt.mods import *
+import matplotlib.pyplot as p
+
+import math
+import numpy
+import random
+import time
+import datetime
+import os
+
+bins = 70
+meanDens = 3e-22
+i = 0
+Msun = 1.99e33
+G = 6.67e-8
+parsec = 3.09e18
+logRhoMin = -3.0
+
+def getRadialProfile_yt(pf, xc, yc, zc, vxc, vyc, vzc, fileout="rad_profile.out", radiusMin=1e-3, radiusSphere=3.0, particleMass = 0.0) : 
+	if (Sphere_Bulk):
+		sp = pf.h.sphere((xc, yc, zc), radiusSphere/pf['pc'])
+		sp2 = pf.h.sphere((xc, yc, zc), Bulk_sphere_radius/pf['pc'])
+		print Bulk_sphere_radius
+		bulk = sp2.quantities["BulkVelocity"]()
+	if (Particle_Bulk):
+		sp = pf.h.sphere((xc, yc, zc), radiusSphere/pf['pc'])
+		#sp2 = pf.h.sphere((xc, yc, zc), 0.01/pf['pc'])
+		bulk = numpy.array([vxc, vyc, vzc])
+	if (NoParticle_Bulk):
+		sp = pf.h.sphere((xc, yc, zc), 4.0/pf['pc'])
+		bulk = sp.quantities["BulkVelocity"]()
+	sp.set_field_parameter("bulk_velocity", bulk)
+	rp = BinnedProfile1D( sp, bins, "Radiuspc", 1e1**logRhoMin, radiusSphere, log_space=True)
+	rp.add_fields("Density", weight="CellVolume")
+	maxDens = rp["Density"][0]
+	
+	rp.add_fields("RadialVelocity")  
+	rp.add_fields("VelocityMagnitude")
+	
+	rp.add_fields("CellMassMsun", accumulation=True,weight=None)
+
+	rbin = rp["Radiuspc"]
+	vrbin = rp["RadialVelocity"]
+	mTbin = rp["CellMassMsun"] + particleMass
+	vKbin = numpy.sqrt(G*mTbin*Msun/(parsec*rbin))
+	rhobin = rp["Density"]
+	mdotbin = 4.0*3.141*rbin**2*vrbin*rhobin
+	rp.add_fields("TangentialVelocity")
[…]

### 2015-05-28 reduce_scripts/plot.py

@@ -1,0 +1,172 @@
+import matplotlib
+matplotlib.use("Agg")
+
+import matplotlib.pyplot as pl
+import numpy as np
+import matplotlib.backends.backend_pdf as mat_pdf
+import argparse
+import glob
+import sys
+import os
+
+parser = argparse.ArgumentParser(description = "start number to end number")
+
+parser.add_argument('start', metavar='N1', type=int)
+parser.add_argument('end', metavar='N2', type=int)
+parser.add_argument('step', metavar='N3', type=int)
+parser.add_argument('--bigsphere', action='store_true')
+parser.add_argument('--smallsphere', action='store_true')
+parser.add_argument('--particle', action='store_true')
+parser.add_argument('--noparticle', action='store_true')
+parser.add_argument('--shell', action='store_true')
+parser.add_argument('--shellsphere', action='store_true')
+
+args = parser.parse_args()
+
+withBigSphere = args.bigsphere
+withSmallSphere = args.smallsphere
+withParticle = args.particle
+withNoParticle = args.noparticle
+withShell = args.shell
+withShellSphere = args.shellsphere
+
+mp = 1.6e-24
+pi = np.pi
+parsec = 3e18
+sec_in_yr = np.pi* 1e7
+Msun = 2e33
+G = 6.67e-8
+
+if (withBigSphere):
+	compare_files = 'bigsphere_'
+	vel_Plot_label = '3.0 pc'
+
+if (withSmallSphere):
+	compare_files = 'smallsphere_'
+	vel_Plot_label = '0.01 pc'
+
+if (withParticle):
+	compare_files = 'part_'
+	vel_Plot_label = 'Bulk by Particle Vel'	
+
+if (withNoParticle):
+	compare_files = 'nopart'
+	vel_Plot_label = 'No Particles'	
+
+if (withShell):
+	compare_files = 'shell_'
+	vel_Plot_label = 'Bulk by Shell'
+
+if (withShellSphere):
+	compare_files = 'shellsphere_'
+	vel_Plot_label = 'Bulk by Sphere in Shell'
+
+file_prefix = 'rad_profile'
[…]

### 2015-06-10 reduce_scripts/vis_multi_setting.py
added script for easy visualization of simulation box

@@ -1,0 +1,260 @@
+import shutil
+import argparse
+import os
+from yt.mods import *
+import matplotlib.colorbar as cb
+from PIL import Image
+
+#def project( filein, xc, yc, zc, zoom_width, fileout = 'junk.png', withParticles=False) :
+#	if(withProjectionFull):
+#		print "Doing a projection plot of the Full Box."
+#		p = ProjectionPlot(pf, "z", "Density")
+#	if not (withProjectionFull):
+#		p = ProjectionPlot(pf, "z", "Density", center = (xc, yc, zc), width = (zoom_width, 'pc'))
+#        p.set_zlim("Density", 1e-3,1e0)
+#	if(withParticles) :
+#		p.annotate_particles(1.0/pf['unitary'], p_size=10.0)
+#        pid = os.getpid()
+#        p.save("junk{0:06d}".format(pid))
+#        shutil.move("junk{0:06d}_Projection_z_Density.png".format(pid),fileout)
+#
+def Proj_plotting(pf, xc, yc, zc, zoom_width, quad, i, j):
+        plot_out_prefix = 'movieframe'
+	if(withProjectionFull):
+		print "Doing a projection plot of the Full Box."
+		p = ProjectionPlot(pf, "z", "Density")
+		fileout="{0}_{1}_{2:04d}_fullprojection.png".format(plot_out_prefix, quad, i)
+		if(withParticles) :
+			p.annotate_particles(1.0/pf['unitary'], p_size=10.0)
+			fileout="{0}_{1}_{2:04d}_fullproject_particle.png".format(plot_out_prefix, quad, i)
+		pid = os.getpid()
+		p.save("junk{0:06d}".format(pid))
+		shutil.move("junk{0:06d}_Projection_{1}_Density.png".format(pid, plot_axis),fileout)
+	if not (withProjectionFull):
+		plot_field = plot_axis + '-velocity'
+		fileout="{0}_{1}_{2:04d}_{3}_{4}_{5:03d}_{6}pc.png".format(plot_out_prefix, quad, i, plot_axis, plot_field, j, zoom_width)
+		p_plot = ProjectionPlot(pf, plot_axis, plot_field, weight_field = 'Density', center = (xc, yc, zc), width = (zoom_width, 'pc'))
+		pid = os.getpid()
+		p_plot.save("junk{0:06d}".format(pid))
[…]

### 2015-06-10 ms.tex
expanded some details on the initialization and some of the numerical results

@@
 periodic grids and 8 levels of refinement on a root grid of $128^3$, giving an effective resolution of $32K^3$.  
-Self-gravity is computed with a multi-grid Poisson solver (see Ricker
-2008), coupled with a fast-Fourier transform solution on the root grid.
+Self-gravity is computed with a multi-grid Poisson solver (see Ricker 2008), coupled with a fast-Fourier transform solution on the root grid.
 Similar to (Lee et.
@@
 2015) our FLASH runs use the Harten-Lax-van Leer-Contact Riemann solver and an unsplit solver.
-({\bf Lee et.
-al use Lee \& Deane (2009) do we?})
+(Lee \& Deane (2009)
 
 We start with a box with the physical length set to $L = 16$pc using periodic boundary condtions.
-The intial mass density is $\rho = 3x10^{-22}$ and the isothermal, ambient gas is set to a sound speed of $c_s = 2.64x10^{4}$ {\bf we also set $Mach_ambient = 5.0$ include?}.
+The intial mass density is $\rho = 3x10^{-22}$ and the isothermal, ambient gas is set to a sound speed of $c_s = 2.64x10^{4}$.
 To initialize our simulations, we drive turbulence by applying a large scale ($1 \le kL \le 2$) solenoidal acceleration field as a momentum and energy source term.
@@
 
-This fully developed turbulent state is the initial condition to which we add self-gravity and sink particle formation for
-our star formation experiments. Sink particles are formed as described in Lee et al. (2014) and described briefly below.
+This fully developed turbulent state is the initial condition to which we add self-gravity and sink particle formation for our star formation experiments.
+Sink particles are formed as described in Lee et al.
+(2014) and described briefly below.
 
-The gravitational force is computed differently than the gas.
+The gravitational force is computed differently between sink particles than between sink particles and the gas.
 Sink particle-sink particle forces are computed via direct N-body calculation.
@@
 \begin{equation}
-%\frac{\rho}{\rho_0} = 45 \text{x} 4^{l} (\frac{N_{root}}{128})^2 (\frac{16\text{pc}}{L})^2 (\frac{c_s}{2.65x10^4 \text{cm s})^2 (\frac{3x10^{-22}\text{g cm^-3}}{\rho_0})
+\frac{\rho}{\rho_{0}} = 45 \cdot 4^{l} \left( \frac{N_{root}}{128} \right)^2 \left( \frac{16 \text{pc}}{L} \right)^2 \left( \frac{c_{s}}{2.65\text{x}10^4 \text{cm s$^{-1}$}} \right)^2 \left( \frac{3 \text{x}10^{-22} \text{g cm$^{-3}$}}{\rho_{0}} \right)
 \label{eq:refinement_criteria}
 \end{equation}
-which when met causes the local grid to be refined by a factor of 2 provided that the maximum refinement level is not reached.
-
-When the Truelove criterion is met at highest refinement level, the excess mass in a cell is transferred either to a newly created sink particle or to a sink particle whose accretion radius includes the cell.
+ where $l$ is the refinement level, with $l = 0$ corresponding to the root grid.
+When this density condition is met the local grid is refined by a factor of 2 provided that the maximum refinement level has not reached.
+We reach a maximum of $\rho / \rho_{0} \approx 3 \text{x} 10^{7}$ at $l$ = 8, the maximum refinement level. Thus we traverse nearly 8 orders of magnitude in the density in our AMR simulations.
+When the Truelove criterion is met at the highest refinement level, the excess mass in a cell is transferred either to a newly created sink particle or to a sink particle whose accretion radius includes the cell.
 This behavior is the same as in Lee et al.
@@
 After this equilibrium is established, we then apply self gravity, adaptive mesh refinement, and sink particle creation.
-In Figure \ref{fig:entire projection} we show a projection along the x-axis of the entire simulation volume that has up to 10 levels of refinement, giving an effective resolution of $128\,{\rm K}^3$.
+In Figure \ref{fig:entire projection} we show a projection along the z-axis of the entire simulation volume that has up to 8 levels of refinement, giving an effective resolution of $32\,{\rm K}^3$.
 Regions that are highly refined are the densest regions which are smoother than the low-density more pixelated regions.
@@
 
-At the instance in time shown in Figure \ref{fig:entire projection}, a total of five star particles have been formed.
-However, these star particles are not formed in isolation.
-Rather they congregate in two star forming regions shown in Figure \ref{fig:star forming regions}.
+At the instance in time shown in Figure \ref{fig:entire projection}, a total of 20 star particles have been formed.
+A couple of these star particles are formed in isolation though the majority formed in binaries and there is a ring of ten in one region.
+They congregated in several star forming regions shown in Figure \ref{fig:star forming regions}.
 In these regions, the sink particles that are formed lay within 0.1 pc of one another, which reflects that the highest density regions are concentrated toward the center.

### 2015-06-18 ms.tex
moved section on velocity graph to after section on plots and numerical results

@@
 
-\section{Detailed Simulations of Turbulent Collapse}
+\section{Detailed Simulations of Turbulent Collapse} \label{sec:simulation setup}
 
@@
 This corresponds to a condition on the density
-\begin{equation}
+\be
 \frac{\rho}{\rho_{0}} = 45 \cdot 4^{l} \left( \frac{N_{root}}{128} \right)^2 \left( \frac{16 \text{pc}}{L} \right)^2 \left( \frac{c_{s}}{2.65\text{x}10^4 \text{cm s$^{-1}$}} \right)^2 \left( \frac{3 \text{x}10^{-22} \text{g cm$^{-3}$}}{\rho_{0}} \right)
 \label{eq:refinement_criteria}
-\end{equation}
+\ee
  where $l$ is the refinement level, with $l = 0$ corresponding to the root grid.
@@
 Our base grid's resolution of $128^3$ gives a cell length of $1.2 \text{x} 10^{-1}$pc which is sufficient to resolve the Jean's length
-\begin{equation}
+\be
 \lambda_J \equiv \sqrt{\frac{\pi c_{s}^2}{G \bar{\rho}}} \approx 3.5 \text{pc}
-\end{equation}
+\ee
 
@@
 
-At the instance in time shown in Figure \ref{fig:entire projection}, a total of 20 star particles have been formed.
+At the instance in time shown in Figure \ref{fig:entire projection}, a total of 20 {\bf GET ACCURATE COUNT} star particles have been formed.
 A couple of these star particles are formed in isolation though the majority formed in binaries and there is a ring of ten in one region.
@@
 \begin{figure*}
-\plottwo{Disk_0132_0000.png}{Disk_0132_0002.png}
-\caption{Cut of the two star forming regions.\label{fig:star forming regions}}
+\plottwo{Disk_0132_0000.png}{movie_quad3_frame_0320_part_0000_05pc.png}
+\caption{Cut of the two star forming regions. The region on the right formed a single isolated particle, after two filaments {\bf INCLUDE5PC image? } had collided.\label{fig:star forming regions}}
 \end{figure*}
+
+% Start with quad3 gold standard
+
+The first particle formed in isolation and remained isolated for the remainder of the simulation run. It is $\approx 1.6x10^5\text{yrs}$ old and is $\approx 10 M_{\odot}$
+
+%% Include image of the velocity and disk
 
@@
 We should note that these star particles are within about 0.1 pc of each other in their respective families.
+
+%\begin{figure*}
+%\plottwo{movie_disk_0132_0001.png}{movie_disk_0132_0003.png}
+%\plottwo{movie_disk_0132_0000.png}{Disk_0132_0000.png}
+%\caption{The upper frames, and the bottom left frame show slices along the angular momentum axis of three local particles in one of the regions demonstrating significant star formation.
+Each of the zoomed-in slices have a radius of $0.02$ pc.
+To determine the angular momentum axis, we take a $0.01$ pc region about the densest local point and calculate the angular momentum vector.
+We then take a slice in the plane normal to the angular momentum vector to plot the regions around these dense points.
+The bottom right frame shows the zoomed-out region of formation, with a radius of $1.5$ pc.
+\label{fig:snapshots}}
+%\end{figure*}
 
@@
 \label{fig:geometry 2}}
+\end{figure}
+
+\subsection{Radial Profile of Velocity after Star Formation}
+
+These simulations display much more complicated dynamics than is captured by the simple analytic theory of Murray and Chang 2014.
+In spite of this, we proceed to compare radially averaged profiles around these high density points to their analytic theory.
+Toward this end, we construct radial profiles around the most massive star particle in each of the star forming regions.
+
+{\bf Include condition to place in which shell?}
+We start with the isolated case seen in the right image of figure \ref{fig:star forming regions}.
+In Figure \ref{fig:quad3_320_velocity}, we show the radial distribution of $v_r$, $v_\phi$, $V_{kepler}$, $v_{turbulent}$ in spherical shells around the star particle.
+In calculating these values, we use a spherical geometry, though we're working on subtracting the velocity in a  cylindrical geometrical case as well.
+FLASH provides us with $\vec{v}$ in each cell, which includes the bulk flow of the cloud.
+To remove the bulk flow velocity we calculate the momentum of the material inside of (including) the thin spherical shell we're interested in.
+We then remove that mass weighted velocity from the total velocity.
+This give us our total velocity line in figure \ref{fig:quad3_320_velocity} (the magenta "dot" line).
+To obtain the radial infall velocity (blue triangle line) we used $p_r[i] = \vec{v_r}[i] \cdot \vec{r}[i] * m[i]$ equation \ref{eq:obtain_vr} to obtain the radial momentum of the ith cell inside the spherical shell of interest.
+We then sum over all $p_{r}[i]$ that are in the shell, divide by the total mass in the shell and obtain the mass weighted radial velocity for that shell.
+To obtain $\vec{v_{\phi}}$ we start with equation \ref{eq:calc v_phi} as we need to know the directionality of $\vec{v}_{\phi}$ in order to properly subtract it to obtain $v_{turbulent}$.
+
+\be
+\vec{v}_{\phi} = \vec{\Omega} \times \vec{r}
+\label{eq:calc v_phi}
+\ee
+With \ref{eq:calc v_phi} in mind, we start with what we know from FLASH, $\vec{r}$ and $\vec{v}$ for each cell, and use that to obtain $\vec{L}$
+
+\be
+\vec{L}_{shell} = \sum{\vec{r} \times \vec{v} * m}
+\label{eq:calc L shell}
+\ee
+where the sum is over all cells in the shell and $\vec{L_{shell}}$ is the angular momentum vector of the shell. We also calculate the elements of the moment of Inertia tensor for each cell.
+
+\be
+I_{ii} = (x_{jj}^2 + x_{kk}^2) * m[n]
+I_{ij} = -x_{i}[n]*x_{j}[n] * m[n]
+\label{eq:moment of inertia}
+\ee
+for the nth cell in the shell. We sum over all cells in the shell to obtain the Inertia tensor for each shell.
+We use $\vec{L}$ and $I$ to obtain $\Omega$
+
+\be
+I^{-1}\vec{L} = \Omega
+\label{eq:Obtain Omega}
+\ee
+having obtained $\Omega$ we refer back to \ref{eq:calc v_phi} and plot the results in figure \ref{fig:quad3_320_velocity} where the black +'s are $v_{\phi}$.
+
+In calculating the turbulent velocity we used a slightly more intelligent approach then the brute force subtract in quadrature.
+\be
+\vec{v}_{t} = \sqrt{\vec{v}^2 + \vec{v}_{\phi}^2 + \vec{v}_{r}^2 - 2(\vec{v} \cdot \vec{v}_{\phi}) - 2(\vec{v} \cdot \vec{v}_{r}) + 2(\vec{v}_{\phi} \cdot \vec{v}_{r})}
+\label{eq:Vrms subtraction}
+\ee
+
+as a check that we were accounting for all of the velocity that we started with, we did a quick check using $\vec{v}_{sum} = \sqrt{\vec{v}^2 + \vec{v}_{\phi}^2 + \vec{v}_{r}^2 + \vec{v}_{t}}$ and it traces $\vec{v}$ extremely well.
+
+The dashed red line is the Keplerian velocity:
+\be
+\vec{v}_{K} = \sqrt{G m \over \vec{r}}
+\label{eq:Kepler velocity}
+\ee
+
+The authors want to touch upon the fact that the total velocity ($\vec{v}$) does not quite match up with $\vec{v}_{K}$.
+The reason is, as one can see from the plotted line of $c_s$ in figure (\ref{fig:quad3_320_velocity}) the flows are mostly super or trans-sonic.
+Thus, one should expect to see shocks, which would remove some energy from the flow in the form of thermal heating.
+These simulations, as mentioned in section \ref{sec:simulation setup} are isothermal.
+Thus, the heating that results from these shocks is immediately radiated away.
+
+%\begin{equation}
+%p_r[i] = \vec{v_r}[i] \cdot \vec{r}[i] * m[i]
+%vr_nomass[i] = (vx[i]*x[i] + vy[i]*y[i] + vz[i]*z[i])/r[i]
+%vr = vr_nomass[i] * cellMass[i]%
+%
+%\label{eq:obtain_vr}
+%\end{equation}
+
+\begin{figure}
+\plotone{velocity_quad3_0320_shellsphere_000.pdf}
+\caption{Plot of the run of velocities for the isolated star particle seen in the right image of figure \ref{fig:star forming regions}.
+The dotted line in magenta is the numerical total velocity.
+The dot-dash green line is the turbulent velocity, the blue triangle line is the radial infall while the black $+$ is $v_{\phi}$.
+A color version of this plot is available online.
+\label{fig:quad3_320_velocity}}
 \end{figure}

### 2015-06-18 ms.tex
added cs to velocity graph and made minor corrections

@@
 We start with the isolated case seen in the right image of figure \ref{fig:star forming regions}.
-In Figure \ref{fig:quad3_320_velocity}, we show the radial distribution of $v_r$, $v_\phi$, $V_{kepler}$, $v_{turbulent}$ in spherical shells around the star particle.
+In Figure \ref{fig:quad3_320_velocity}, we show the radial distribution of the radial infall velocity $\vec{v}_r$, the angular velocity $\vec{v}_{\phi}$, the keplerian velocity $\vec{v}_{K}$ and the turbulent velocity $\vec{v}_{t}$ in spherical shells around the star particle.
 In calculating these values, we use a spherical geometry, though we're working on subtracting the velocity in a  cylindrical geometrical case as well.
@@
 This give us our total velocity line in figure \ref{fig:quad3_320_velocity} (the magenta "dot" line).
-To obtain the radial infall velocity (blue triangle line) we used $p_r[i] = \vec{v_r}[i] \cdot \vec{r}[i] * m[i]$ equation \ref{eq:obtain_vr} to obtain the radial momentum of the ith cell inside the spherical shell of interest.
+To obtain the radial infall velocity (blue triangle line) we use $p_r[i] = \vec{v}_r[i] \cdot \vec{r}[i] * m[i]$ to obtain the radial momentum of the ith cell inside the spherical shell of interest.
 We then sum over all $p_{r}[i]$ that are in the shell, divide by the total mass in the shell and obtain the mass weighted radial velocity for that shell.
-To obtain $\vec{v_{\phi}}$ we start with equation \ref{eq:calc v_phi} as we need to know the directionality of $\vec{v}_{\phi}$ in order to properly subtract it to obtain $v_{turbulent}$.
+To obtain $\vec{v}_{\phi}$ we start with equation \ref{eq:calc v_phi} as we need to know the directionality of $\vec{v}_{\phi}$ in order to properly subtract it to obtain $\vec{v}_{t}$.
 
@@
 
-as a check that we were accounting for all of the velocity that we started with, we did a quick check using $\vec{v}_{sum} = \sqrt{\vec{v}^2 + \vec{v}_{\phi}^2 + \vec{v}_{r}^2 + \vec{v}_{t}}$ and it traces $\vec{v}$ extremely well.
+as a check that we were accounting for all of the velocity that we started with, we added the velocities in quadrature: $\vec{v}_{sum} = \sqrt{\vec{v}^2 + \vec{v}_{\phi}^2 + \vec{v}_{r}^2 + \vec{v}_{t}}$ and it traces $\vec{v}$ extremely well.
 

### 2015-06-18 reduce_scripts/all_profiles.py
added in update to all profiles and plot.py

@@ -10,6 +10,8 @@ import random
 import time
 import datetime
 import os
+from numpy.linalg import inv
+
 
 bins = 70
 meanDens = 3e-22
@@ -127,7 +129,6 @@ def getRadialProfile_py(pf, xc, yc, zc, fileout="rad_profile.out", radiusMin=1e-
 				vy_sphere_bulk_bin[index] = vy_sphere_bulk_bin[index] + (vy[i]*cellMass[i])
 				vz_sphere_bulk_bin[index] = vz_sphere_bulk_bin[index] + (vz[i]*cellMass[i])
 		menc[0] = mbin[0]
-		
 		for shell in range(1,bins):
 			menc[shell] = mbin[shell] + menc[shell-1]
 			vx_sphere_bulk_bin[shell] = vx_sphere_bulk_bin[shell] + vx_sphere_bulk_bin[shell-1] 
@@ -138,15 +139,6 @@ def getRadialProfile_py(pf, xc, yc, zc, fileout="rad_profile.out", radiusMin=1e-
 		vx_bulkvelocity_bin = vx_sphere_bulk_bin/menc
 		vy_bulkvelocity_bin = vy_sphere_bulk_bin/menc
 		vz_bulkvelocity_bin = vz_sphere_bulk_bin/menc
-
-		#vx_bulkvelocity_bin[:] = vx_bulkvelocity_bin[-1]
-		#vy_bulkvelocity_bin[:] = vy_bulkvelocity_bin[-1]
-		#vz_bulkvelocity_bin[:] = vz_bulkvelocity_bin[-1]
-
-#		print vx_bulkvelocity_bin[-1], vy_bulkvelocity_bin[-1], vz_bulkvelocity_bin[-1]
-#		sp2 = pf.h.sphere((xc, yc, zc), 3.0/pf['pc'])
-#		bulk = sp2.quantities["BulkVelocity"]()
-#		print bulk
 
 ###########################################################################
 ###########################################################################
@@ -240,6 +232,7 @@ def getRadialProfile_py(pf, xc, yc, zc, fileout="rad_profile.out", radiusMin=1e-
 	# Find the middle radius of the bin
 	lgrbin = lgradiusMin + (lgradiusSph-lgradiusMin)*numpy.arange(bins)/bins
 	rbin = 1e1**lgrbin
+	# put r in cm
 	rbinparsec = rbin * parsec
 
[…]

### 2015-06-18 reduce_scripts/plot.py

@@ -9,6 +9,104 @@ import glob
 import sys
 import os
 
+def plotting_routine(rbin, vrbin, vrmsbin, vrmsnbin, vKbin, vmagbin, vmagnbin, mTbin, rhobin, mdotbin, norm, angXbin, angYbin, angZbin, vphi_magbin):
+	velocity_plot(rbin, vrbin, vrmsbin, vrmsnbin, vKbin, vmagbin, vmagnbin, mTbin, vphi_magbin)
+	density_plot(rbin, mTbin, rhobin)
+	ang_moment_plot(rbin, mTbin, vphi_magbin)
+	total_mass_plot(rbin, mTbin)
+	Mdot_plot(rbin, vrbin, rhobin, mdotbin)
+	pdf_close(pdf_timestep)
+
+def velocity_plot(rbin, vrbin, vrmsbin, vrmsnbin, vKbin, vmagbin, vmagnbin, mTbin, vphi_magbin):
+	pl.clf()
+	pl.loglog( rbin, -vrbin/1e5, 'bv-', label='Infall Velocity')
+	pl.loglog( rbin, vrmsbin/1e5, 'g.-', label='Turbulent Velocity')  
+	pl.loglog( rbin, vmagbin/1e5, 'm.', label='Total Velocity')  
+	pl.loglog( rbin, vKbin/1e5, 'r--', label='Keplerian Velocity')  
+	pl.loglog( rbin, vphi_magbin/1e5, 'k+', label='Angular Velocity')
+	pl.loglog( rbin, test_all_together/1e5, 'c', label='Sum Velocity')  
+	pl.axhline( c_s/1e5, color = 'c',label='cs')  
+
+	pl.legend(loc=0, fontsize='medium', frameon=False)
+	pl.ylabel('Velocity ($km/s$)')
+	pl.xlabel('Radius ($pc$)')
+	pl.ylim(1e-2, 10e0)
+	pl.xlim(1e-3, 3e0)
+	pl.title('{0} framestep {1:04d}'.format(vel_Plot_label, framestep))
+	pdf_timestep.savefig()
+	
+def density_plot(rbin, mTbin, rhobin):
+	pl.clf()
+	# rhobin should be in g/cm**3
+	pl.loglog(rbin, rhobin/mp, label='Number density') #plotting number density
+	pl.loglog(rbin, 2.0/(rbin**3), label='-3')
+	pl.loglog(rbin, 80./(rbin**2.5), label='-2.5')
[…]

### 2015-07-02 ms.tex
forgot to update the reduction scripts in previous commit

@@
 
-The first particle formed in isolation and remained isolated for the remainder of the simulation run. It is $\approx 1.6x10^5\text{yrs}$ old and is $\approx 10 M_{\odot}$
+The first particle formed in isolation and remained isolated for the remainder of the simulation run. It is $\approx 1.6x10^5\text{yrs}$ old and is $7 M_{\odot}$
 
@@
 
-\begin{figure*}
+\begin{figure}
 \plottwo{Average_5_013.pdf}{Average_5_24.pdf}
@@
 \label{fig:accretion average rho}}
-\end{figure*}
+\end{figure}
 
@@
 
-\subsection{Radial Profile of Velocity after Star Formation}
-
-These simulations display much more complicated dynamics than is captured by the simple analytic theory of Murray and Chang 2014.
-In spite of this, we proceed to compare radially averaged profiles around these high density points to their analytic theory.
-Toward this end, we construct radial profiles around the most massive star particle in each of the star forming regions.
+\subsection{Derivation of Radial Profile of Velocity}
 
@@
 We start with the isolated case seen in the right image of figure \ref{fig:star forming regions}.
-In Figure \ref{fig:quad3_320_velocity}, we show the radial distribution of the radial infall velocity $\vec{v}_r$, the angular velocity $\vec{v}_{\phi}$, the keplerian velocity $\vec{v}_{K}$ and the turbulent velocity $\vec{v}_{t}$ in spherical shells around the star particle.
+In Figure \ref{fig:quad3_320_velocity_example}, we show the radial distribution of the radial infall velocity $\vec{v}_r$, the angular velocity $\vec{v}_{\phi}$, the keplerian velocity $\vec{v}_{K}$ and the turbulent velocity $\vec{v}_{t}$ in spherical shells around the star particle.
 In calculating these values, we use a spherical geometry, though we're working on subtracting the velocity in a  cylindrical geometrical case as well.
+
+\begin{figure}
+\plotone{velocity_320_example.pdf}
+\caption{Plot of the run of velocities for the isolated star particle seen in the right image of figure \ref{fig:star forming regions}.
+The dotted line in magenta is the numerical total velocity.
+The dot-dash green line is the turbulent velocity, the blue triangle line is the radial infall while the black $+$ is $v_{\phi}$.
+A color version of this plot is available online.
+\label{fig:quad3_320_velocity_example}}
+\end{figure}
+
 FLASH provides us with $\vec{v}$ in each cell, which includes the bulk flow of the cloud.
@@
 We then remove that mass weighted velocity from the total velocity.
-This give us our total velocity line in figure \ref{fig:quad3_320_velocity} (the magenta "dot" line).
+This give us our total velocity line in figure \ref{fig:quad3_320_velocity_example} (the magenta "dot" line).
 To obtain the radial infall velocity (blue triangle line) we use $p_r[i] = \vec{v}_r[i] \cdot \vec{r}[i] * m[i]$ to obtain the radial momentum of the ith cell inside the spherical shell of interest.
@@
 
-\be
-I_{ii} = (x_{jj}^2 + x_{kk}^2) * m[n]
+\begin{gather*}
+I_{ii} = (x_{jj}^2 + x_{kk}^2) * m[n]\\
 I_{ij} = -x_{i}[n]*x_{j}[n] * m[n]
 \label{eq:moment of inertia}
-\ee
+\end{gather*}
 for the nth cell in the shell. We sum over all cells in the shell to obtain the Inertia tensor for each shell.
@@
 \ee
-having obtained $\Omega$ we refer back to \ref{eq:calc v_phi} and plot the results in figure \ref{fig:quad3_320_velocity} where the black +'s are $v_{\phi}$.
+having obtained $\Omega$ we refer back to \ref{eq:calc v_phi} and plot the results in figure \ref{fig:quad3_320_velocity_example} where the black +'s are $v_{\phi}$.
 
@@
 as a check that we were accounting for all of the velocity that we started with, we added the velocities in quadrature: $\vec{v}_{sum} = \sqrt{\vec{v}^2 + \vec{v}_{\phi}^2 + \vec{v}_{r}^2 + \vec{v}_{t}}$ and it traces $\vec{v}$ extremely well.
+This check is included
 
@@
 The authors want to touch upon the fact that the total velocity ($\vec{v}$) does not quite match up with $\vec{v}_{K}$.
-The reason is, as one can see from the plotted line of $c_s$ in figure (\ref{fig:quad3_320_velocity}) the flows are mostly super or trans-sonic.
+The reason is, as one can see from the plotted line of $c_s$ in figure (\ref{fig:quad3_320_velocity_example}) the flows are mostly super or trans-sonic.
 Thus, one should expect to see shocks, which would remove some energy from the flow in the form of thermal heating.
@@
 Thus, the heating that results from these shocks is immediately radiated away.
-
-%\begin{equation}
-%p_r[i] = \vec{v_r}[i] \cdot \vec{r}[i] * m[i]
-%vr_nomass[i] = (vx[i]*x[i] + vy[i]*y[i] + vz[i]*z[i])/r[i]
-%vr = vr_nomass[i] * cellMass[i]%
-%
-%\label{eq:obtain_vr}
-%\end{equation}
+\subsection{Evolution of the Radial Profile before Star Formation}
 
 \begin{figure}
-\plotone{velocity_quad3_0320_shellsphere_000.pdf}
-\caption{Plot of the run of velocities for the isolated star particle seen in the right image of figure \ref{fig:star forming regions}.
-The dotted line in magenta is the numerical total velocity.
-The dot-dash green line is the turbulent velocity, the blue triangle line is the radial infall while the black $+$ is $v_{\phi}$.
-A color version of this plot is available online.
-\label{fig:quad3_320_velocity}}
+\plottwo{velocity_quad3_0320_shellsphere_000.png}{density_quad3_0320_shellsphere_000.png}
+\plottwo{mass_quad3_0320_shellsphere_000.png}{mdot_quad3_0320_shellsphere_000.png}
+\caption{This set is a place holder for the set of four before star formation. For Quad2
+\label{fig:quad3_000_four}}
 \end{figure}
+
+\begin{figure}
+\plottwo{velocity_quad3_0320_shellsphere_000.png}{density_quad3_0320_shellsphere_000.png}
+\plottwo{mass_quad3_0320_shellsphere_000.png}{mdot_quad3_0320_shellsphere_000.png}
+\caption{This set is a place holder for the set of four before star formation. For Quad3
+\label{fig:quad_000_four}}
+\end{figure}
+
+\subsection{Evolution of the Radial Profile after Star Formation}
+
+Before the star even forms its apparent that
+
+\begin{figure}
+\plotone{velocity_quad3_0200_shellsphere_000.png}
+\caption{Plot of the velocity 40,000 years after formation. Note the turnover from turbulent  pressure support to angular support and the formation of an accrection disk is already well underway.
+\label{fig:quad3_200_velocity}}
+\end{figure}
+
+\begin{figure*}
+\plottwo{velocity_quad3_0320_shellsphere_000.png}{density_quad3_0320_shellsphere_000.png}
+\plottwo{mass_quad3_0320_shellsphere_000.png}{mdot_quad3_0320_shellsphere_000.png}
+\caption{Upper right is the run of density.
+Note the slope at large radius $\approx r^{-1.8}$ increases to $\approx r^{-3}$ at radii inside of a couple hundreths of a parsec.Lower left is a plot of the mass as a function of radius for the isolated star particle seen in the right image of figure \ref{fig:star forming regions}.
+Note that the stellar mass dominates out until about a tenth of a parsec.
+\label{fig:quad3_320_four}}
+\end{figure*}
+
+A key point in this plot and throughout the time evolution plots is that the turbulent velocity has become the more dominant velocity over the radial infall.
+What this means is that Kinetic energy from the falling gas is being transfered into the turbulent motions of the gas.
 
@@
 
-\begin{figure*}
-\plottwo{Disk_0132_0000.png}{Track_0132_000_rho.png}
-\plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
-\plottwo{Track_0132_003_mdot.png}{pdf_frame_0132_0000.png}
-\caption{Tracking 132 000  These plots show various data for one of the particles formed in the more massive region in the simulation.
+%\begin{figure*}
+%\plottwo{Disk_0132_0000.png}{Track_0132_000_rho.png}
+%\plottwo{Track_0132_000_vr.png}{Track_0132_000_mass.png}
+%\plottwo{Track_0132_003_mdot.png}{pdf_frame_0132_0000.png}
+%\caption{Tracking 132 000  These plots show various data for one of the particles formed in the more massive region in the simulation.
 The upper left plot shows the zoomed out region with radius of $1.5$ pc.
@@
 \label{fig:132 000 graphs}}
-\end{figure*}
+%\end{figure*}
 
-\begin{figure*}
-\plottwo{Disk_0132_0002.png}{Track_0132_002_rho.png}
-\plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
-\plottwo{Track_0132_004_mdot.png}{pdf_frame_0132_0002.png}
-\caption{Tracking 132 002  These plots show the data for a different particle formed in the simulation at the same time as the previous plots.
+%\begin{figure*}
+%\plottwo{Disk_0132_0002.png}{Track_0132_002_rho.png}
+%\plottwo{Track_0132_002_vr.png} {Track_0132_002_mass.png}
+%\plottwo{Track_0132_004_mdot.png}{pdf_frame_0132_0002.png}
+%\caption{Tracking 132 002  These plots show the data for a different particle formed in the simulation at the same time as the previous plots.
 The upper left graph of the density in the region neatly matches the -3/2 slope.
@@
 \label{fig:132_002_graphs}}
-\end{figure*}
+%\end{figure*}
 
-\begin{figure*}
-\plottwo{Track_0256_000_l.png}{Track_0132_000_l.png}
-\caption{WRITE CAPTION: Angular momentum as a function of distance from maximum density point.     \label{fig:angular_momentum_example}}
-\end{figure*}
+%\begin{figure*}
+%\plottwo{Track_0256_000_l.png}{Track_0132_000_l.png}
+%\caption{WRITE CAPTION: Angular momentum as a function of distance from maximum density point.     \label{fig:angular_momentum_example}}
+%\end{figure*}
 

### 2015-07-02 reduce_scripts/all_profiles.py

@@ -749,7 +749,7 @@ for i in range(args.start,args.end,args.step) :
 			
 			vxc = vxp[j]
 			vyc = vyp[j]
-			vzc = vzp[j]	 
+			vzc = vzp[j]
 			if (withParticle) or (withSmallSphere) or (withBigSphere):
 				print 'Using yt'
 				print 'On particle:', j + 1, 'of:', xp.size

### 2015-07-02 reduce_scripts/plot.py

@@ -11,20 +11,25 @@ import os
 
 def plotting_routine(rbin, vrbin, vrmsbin, vrmsnbin, vKbin, vmagbin, vmagnbin, mTbin, rhobin, mdotbin, norm, angXbin, angYbin, angZbin, vphi_magbin):
 	velocity_plot(rbin, vrbin, vrmsbin, vrmsnbin, vKbin, vmagbin, vmagnbin, mTbin, vphi_magbin)
+	pdf_timestep.savefig()
 	density_plot(rbin, mTbin, rhobin)
+	pdf_timestep.savefig()
 	ang_moment_plot(rbin, mTbin, vphi_magbin)
+	pdf_timestep.savefig()
 	total_mass_plot(rbin, mTbin)
+	pdf_timestep.savefig()
 	Mdot_plot(rbin, vrbin, rhobin, mdotbin)
+	pdf_timestep.savefig()
 	pdf_close(pdf_timestep)
 
 def velocity_plot(rbin, vrbin, vrmsbin, vrmsnbin, vKbin, vmagbin, vmagnbin, mTbin, vphi_magbin):
 	pl.clf()
 	pl.loglog( rbin, -vrbin/1e5, 'bv-', label='Infall Velocity')
 	pl.loglog( rbin, vrmsbin/1e5, 'g.-', label='Turbulent Velocity')  
-	pl.loglog( rbin, vmagbin/1e5, 'm.', label='Total Velocity')  
+#	pl.loglog( rbin, vmagbin/1e5, 'm.', label='Total Velocity')  
 	pl.loglog( rbin, vKbin/1e5, 'r--', label='Keplerian Velocity')  
 	pl.loglog( rbin, vphi_magbin/1e5, 'k+', label='Angular Velocity')
-	pl.loglog( rbin, test_all_together/1e5, 'c', label='Sum Velocity')  
+#	pl.loglog( rbin, test_all_together/1e5, 'c', label='Sum Velocity')  
 	pl.axhline( c_s/1e5, color = 'c',label='cs')  
 
 	pl.legend(loc=0, fontsize='medium', frameon=False)
@@ -33,79 +38,100 @@ def velocity_plot(rbin, vrbin, vrmsbin, vrmsnbin, vKbin, vmagbin, vmagnbin, mTbi
 	pl.ylim(1e-2, 10e0)
 	pl.xlim(1e-3, 3e0)
[…]

### 2015-07-18 ms.tex
added some sections on the run of velocity

@@
 \shorttitle{Neutron Star Redshift Measurements} 
-\shortauthors{Bildsten, Chang and Paerels}
+\shortauthors{Murray, Change and Murray}
 
@@
 We then remove that mass weighted velocity from the total velocity.
-This give us our total velocity line in figure \ref{fig:quad3_320_velocity_example} (the magenta "dot" line).
+This gives us our total velocity line in figure \ref{fig:quad3_320_velocity_example} (the magenta "dot" line).
 To obtain the radial infall velocity (blue triangle line) we use $p_r[i] = \vec{v}_r[i] \cdot \vec{r}[i] * m[i]$ to obtain the radial momentum of the ith cell inside the spherical shell of interest.
@@
 
+Write things here.
+
 \begin{figure}
@@
 
-\subsection{Evolution of the Radial Profile after Star Formation}
+\subsection{Evolution of the Run of Density and Velocity after Star Formation}
 
-Before the star even forms its apparent that
+Use the 498 plot first, then use the quad3 plot. bc quad3 is starved of fuel.
+The Cross over point of the edge of the disk is the switch btwn the v phi and v turbulent.
+the second cross over is the radius of the sphere of influence of the star.
+R_{*} that matches up with the plot of the mass.
+go to the same radius and show that matches up with the 2nd cross over point.
+
+Show the quad2 498 velocity and mass. make the point that the mass about doubles at .1 pc and that matches with that second cross over point. That indicates the sphere of influence of the star
+
+After star formation we see that the angular velocity overtakes the turbulent velocity as the dominant velocity and moves outward to $10^{-2}$pc.
+This cross-over point is where the gas runs into the accretion disk of the star.
+The authors note that the disks are larger than what they would be observationally due to our neglecting magnetic fields.
+This omission is currently being resolved in current simulations and will be addressed in a later paper.
+This turnover point can also be seen in figure (\ref{fig:quad3_slice_placeholder}).
+As the star grows the infall velocity at large radii is eventually overtaken by the turbulent velocity.
+This implies that part of the kinetic energy from the infalling gas gets converted into turbulence.
+The same conclusion can be had from the potential energy, that it is converting into turbulent energy.
+\bf(Explain how you get this conclusion.
+Why is it consistent with Robertson and Goldreich)
+It is worth noting again that we start star formation with with a substantial turbulent state driven by a solenoidal acceleration field. \bf(Klessen et al)
+did a simulation that started with spherical collapse with small amt of turblulence.
+What they found was that while turbulence does grow most energy goes to $\vec{v}_r$.
+In other words, the growth time for the turbulence is comparable to or longer than the infall time.
+By the time the globule collapses the growth in $\vec{v}_{t}$ is negligible.
+Given these results, it appears that the calculation of the growth rate of turbulent motion from laminar flow is not applicable to a turbulent collapse.
 
